---
title: "ROTA Metagenomics: Alpha diversity"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: true
    theme: lumen
---

# Summary
This document contains the scripts used to calculate and analyze alpha diversity metrics on the Rota metagenome dataset.

# Setup
In this section are the chunks that call libraries into the environment and define some functions for plots that are used repeatedly. Just to make the code a little bit cleaner.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.width=10, fig.height=8)
```

```{r libraries}
## override library function to silence all package messages.
silence = function(x) suppressPackageStartupMessages(suppressWarnings({invisible({capture.output(x)})}))
library = function(...) silence({base::library(...)})
library('data.table')
library('tidyverse')

load(file="initialized-dataset.RData")

set.seed(42)
```

# Alpha Diversity.
Here, wealculate alpha diversity metrics.
```{r AlphaDiv}
library('vegan')
library('picante')
library('VGAM')
library('ggpubr')

## Effective species richness
emr = goodcounts %>%
    abundance(0.0025) %>%
    apply(2, function(x) sum(x > 0))

## Transpose.
ogurarecounts = rarecounts %>% t()

## Run alpha diversity metrics
metrics = c('shannon', 'simpson')
divs = metrics %>% lapply(function(m) vegan::diversity(ogurarecounts, m)) %>% do.call('cbind', .) %>% as.data.table(keep.rownames=TRUE) %>% setNames(c('rn', metrics)) %>% setkey(rn)

## Run faith's PD.
pd = picante::pd(ogurarecounts, rwoltree) %>%
    as.data.table(keep.rownames=TRUE) %>%
    setkey(rn)

## Bind dataframes.
setkey(metadata, sample_name)
alphadivs = merge(divs, pd) %>% setkey(rn)
alphadivdata = metadata[alphadivs,]
alphadivdata = alphadivdata[,EMR := emr[sample_name]]

## Add some vars
alphadivdata[,log_post_rota_sero_iga := log(post_rota_sero_iga)]
alphadivdata[,log_pre_rota_sero_iga := log(pre_rota_sero_iga)]

## Drop individuals missing a timepoint.
alphadivdata = alphadivdata[sample_name %in% colnames(all_timepoint_rarecounts),]

## Version of the dataset with selected columns and NAs dropped.
alphadivnona = alphadivdata[, .(SR, EMR, PD, shannon, simpson
                              , enrollment_id, country, visit
                              , formula_bm, sex, age, feeding_milk, feeding_form
                              , abx_status, gest_age, inf_birth_weight, delivery
                              , Seroconverted, Secretor, rota_sero_iga_fc
                              , log_pre_rota_sero_iga, log_post_rota_sero_iga)
                            ] %>% na.omit()

## After dropping NAs, drop individuals missing a timepoint one more time.
alphadivnona = alphadivnona[enrollment_id %in%
               alphadivnona[,length(visit) %>% unique()
                            , by = enrollment_id][V1 > 2 ,enrollment_id],]

```

```{r PlottingFunctions}
library('ggplot2')
library('cowplot')

## Jitterplot.
ggplot_jitterplot = function(dataset, x, y, color, group, facet_formula)
{
    facet_formula = as.formula(facet_formula)
    ggplot(dataset) +
    geom_path(aes_string(x = x, y = y, color = color, group = group)
            , height = 0, alpha = 0.3, size = 1) +
    geom_jitter(aes_string(x = x, y = y, color = color)
              , height = 0, alpha = 0.5, size = 2) +
    facet_wrap(facet_formula) +
    theme_bw()
}

## Boxplot.
ggplot_boxplot = function(dataset, x, y, color, facet_formula)
{
    facet_formula = as.formula(facet_formula)
    ggplot(dataset) +
        geom_boxplot(aes_string(x = x, y = y, color = color)) +
        facet_wrap(facet_formula) +
        theme_bw()
}

## loess curves.
ggplot_loess = function(dataset, x, y, color, facet_formula)
{
    facet_formula = as.formula(facet_formula)
    ggplot(dataset) +
        stat_smooth(aes_string(x = x, y = y, color = color)
                  , method = 'loess', formula = 'y ~ x') +
        geom_jitter(aes_string(x = x, y = y, color = color)
                  , height = 0, alpha = 0.5, size = 2) +
        facet_wrap(facet_formula) +
        theme_bw()
}

all_plots = function(dataset, y)
{
    ggplot_jitterplot(dataset = dataset, x = 'age', y = y
                      , color = 'country', group = 'enrollment_id'
                      , facet_formula = country ~ delivery) %>% plot()

    ggplot_boxplot(dataset = dataset, x = 'visit', y = y
                       , color = 'country'
                       , facet_formula = ~delivery) %>% plot()

    ggplot_loess(dataset = dataset, x = 'age', y = y
                 , color = 'country'
                 , facet_formula = delivery ~ country) %>% plot()
}
        

```

Univariate correlations show that IgA foldchange is near-significant with respect to species richness, however we don't have the statistical power necessary for a significant result slicing the dataset up like this, and so we'll just move on.
```{r Univariate}
library('broom')

metrics = c('log(SR)','log(PD)','shannon','logit(simpson)', 'log(EMR)')
visits = metadata[,visit] %>% na.omit() %>% unique()

## Correlate fc to NB + 2mo + 4mo
rota_fc_univariate = lapply(visits,
function(v)
{
    metrics %>%
    lapply(function(metric)
    {
        cor.test(alphadivdata[visit == v, eval(parse(text=metric))]
               , alphadivdata[visit == v, log(rota_sero_iga_fc)]
               , method = 'spearman') %>%
            broom::tidy() %>% 
            cbind(metric, dv = 'IgA foldchange', .)
    }) %>% do.call('rbind',.) %>% cbind(visit = v, .)
}) %>% do.call('rbind',.)

rota_fc_univariate$p.adj = p.adjust(rota_fc_univariate$p.value)

## Correlate pre to the NB + 2mo
rota_pre_univariate = lapply(visits[-3],
function(v)
{
    metrics %>%
    lapply(function(metric)
    {
        cor.test(alphadivdata[visit == v, eval(parse(text=metric))]
               , alphadivdata[visit == v, log_pre_rota_sero_iga]
               , method = 'spearman') %>%
            broom::tidy() %>% 
            cbind(metric, dv = 'Pre-vaccine IgA', .)
    }) %>% do.call('rbind',.) %>% cbind(visit = v, .)
}) %>% do.call('rbind',.)

rota_pre_univariate$p.adj = p.adjust(rota_pre_univariate$p.value)

rota_pre_univariate %>% arrange(p.adj)
rota_fc_univariate %>% arrange(p.adj)
```
Here, we plot the generic alpha div metrics.
```{r PlotAlphaDivs, fig.width=6, fig.height = 5}
all_plots(alphadivnona, y = 'log(SR)')
all_plots(alphadivnona, y = 'log(PD)')
all_plots(alphadivnona, y = 'shannon')
all_plots(alphadivnona, y = 'logit(simpson)')
```

# Linear mixed effect models
In this section, we use a set of linear mixed models to determine the factors that significantly influence alpha diversity among the cohort. 

First, we plot the response vars to make sure that they are roughly normally distributed. Simpson is not so, however applying a logit transform fixes the issue. We also seem to get some improvement by log-transforming OGU Richness and Faith's PD. Shannon is better (although not perfect) sans transformation.
```{r ModelFuns}
library('lme4')
library('afex')
library('car')
library('performance')

## Quick enclose function, puts a term in parens.
enclose = function(x) paste0('(',x,')')

fixedeff = c('formula_bm', 'visit', 'abx_status', 'country', 'sex'
             , 'inf_birth_weight', 'gest_age', 'delivery'
             , 'Seroconverted', 'Secretor')
randeff = c('enrollment_id')

norm_plot = function(x, transform = 'no', label)
{
    no = function(x) x
    x_trans = get(transform)(x)
    pd = data.table(x, x_trans)
    list(
        ggplot(pd, aes(x=x)) +
            geom_density(alpha = 0.5, color = 'black', fill = 'gray') +
            theme_bw() +
            ggtitle(label)
        
      , ggplot(pd, aes(x=x_trans)) +
            geom_density(alpha = 0.5, color = 'black', fill = 'gray') +
            theme_bw() +
            ggtitle(paste0(label, '(', transform, '-transform)'))
      
      , ggplot(pd,aes(sample=x)) +
            stat_qq() +
            stat_qq_line() +
            theme_bw() +
            ggtitle(paste0('QQ plot: ',label))
        
      , ggplot(pd,aes(sample=x_trans)) +
            stat_qq() +
            stat_qq_line() +
            theme_bw() +
            ggtitle(paste0('QQ plot: ',label, '(', transform, '-transform)'))
    ) %>% plot_grid(plotlist = .)
}

lmm_modl = function(data, result_term, fixedeff, randeff, label)
{
    random_int = lapply(randeff, function(x) paste0('1 | ', x)) %>% lapply(enclose)
    
    fixed_effects_terms = paste0(fixedeff, collapse = ' + ')
    
    rand_int_terms = random_int %>% 
                        unlist() %>% 
                        unique() %>% 
                        paste0(collapse = ' + ')
    
    effect_terms = paste0(fixed_effects_terms, ' + ', rand_int_terms)
    
    ## We're going to figure out what the best model for the random effect is first.
    nullform = as.formula(paste0(result_term, ' ~ 1 + ', rand_int_terms))
    nullmodl = lmer(nullform, data = data, REML = FALSE, na.action = 'na.fail')
    icc(nullmodl)
    
    ## Build the full formula & model
    fullform = as.formula(paste0(result_term, ' ~ 1 + ', effect_terms))
    fullmodl = lmer(fullform, data = data, REML = FALSE, na.action = 'na.fail')
    
    ## Get vifs & new icc
    vif(fullmodl)
    icc(fullmodl)
    
    ## Make sure that our model is informative.
    modlaov = anova(nullmodl, fullmodl, refit=TRUE)
    print(modlaov)
    
    fit = data.table(predicted = predict(fullmodl), residual = residuals(fullmodl))
    
    list(
        ggplot(fit,aes(x=residual)) +
            geom_histogram() +
            theme_bw() +
            ggtitle(paste0("Residuals / OGU Richness", label))
      , ggplot(fit,aes(x=residual, y = predicted)) +
          geom_point() +
          theme_bw() +
          ggtitle(paste0("Residuals vs. predicted / ", label))
      , ggplot(fit, aes(sample=residual)) +
          stat_qq() +
          stat_qq_line() +
          theme_bw() +
          ggtitle(paste0("Residual QQ-plot / ", label))
    ) %>% plot_grid(plotlist = .)
    
    aovmixed = mixed(fullform
                     , data=data
                     , check_contrasts = TRUE
                     , test_intercept = TRUE
                     , method="KR")
    print(aovmixed)
    print(summary(fullmodl))
    
    signifdt = aovmixed$anova_table %>% as.data.table(keep.rownames=TRUE)
    return(signifdt)
}
```

Build linear model for species richness.
```{r BuildModels}
norm_plot(alphadivnona[,SR], transform = 'log', label = 'OGU Richness')  
SR_signifdt = lmm_modl(alphadivnona, 'log(SR)', fixedeff, randeff, label = 'OGU Richness')

norm_plot(alphadivnona[,EMR], transform = 'log', label = 'EM Richness')  
EMR_signifdt = lmm_modl(alphadivnona, 'EMR', fixedeff, randeff, label = 'OGU Richness')

norm_plot(alphadivnona[,PD], transform = 'log', label = "Faith's PD")  
PD_signifdt = lmm_modl(alphadivnona, 'log(PD)', fixedeff, randeff, label = "Faith's PD")

norm_plot(alphadivnona[,simpson], transform = 'logit', label = "Simpson")  
simpson_signifdt = lmm_modl(alphadivnona, 'logit(simpson)', fixedeff, randeff, label = 'Simpson')

norm_plot(alphadivnona[,shannon], transform = 'log', label = "Shannon")  
shannon_signifdt = lmm_modl(alphadivnona, 'shannon', fixedeff, randeff, label = 'Shannon')
```
```{r FollowUp}
fixedeff = c('formula_bm', 'visit', 'abx_status', 'country', 'sex'
             , 'inf_birth_weight', 'gest_age', 'delivery'
             , 'log_pre_rota_sero_iga', 'Secretor')
randeff = c('enrollment_id')

SR_pre_signifdt = lmm_modl(alphadivnona, 'log(SR)', fixedeff, randeff, label = 'OGU Richness')
PD_pre_signifdt = lmm_modl(alphadivnona, 'log(PD)', fixedeff, randeff, label = 'Faith\'s PD')
EMR_pre_signifdt = lmm_modl(alphadivnona, 'EMR', fixedeff, randeff, label = 'OGU EMR')
simpson_pre_signifdt = lmm_modl(alphadivnona, 'logit(simpson)', fixedeff, randeff, label = 'Simpson')
shannon_pre_signifdt = lmm_modl(alphadivnona, 'shannon', fixedeff, randeff, label = 'Shannon')

fixedeff = c('formula_bm', 'visit', 'abx_status', 'country', 'sex'
             , 'inf_birth_weight', 'gest_age', 'delivery'
             , 'log(rota_sero_iga_fc)', 'Secretor')
randeff = c('enrollment_id')

```

```{r PlotSignifs}

signifdt = list(
      SR_signifdt[,.(variable = rn, Fval = `F`, pval = `Pr(>F)`, metric = 'Richness')]
    , EMR_signifdt[,.(variable = rn, Fval = `F`, pval = `Pr(>F)`, metric = 'EMR')]
    , PD_signifdt[,.(variable = rn, Fval = `F`, pval = `Pr(>F)`, metric = 'PD')]
    , shannon_signifdt[,.(variable = rn, Fval = `F`, pval = `Pr(>F)`, metric = 'shannon')]
    , simpson_signifdt[,.(variable = rn, Fval = `F`, pval = `Pr(>F)`, metric = 'simpson')]
) %>% do.call('rbind', .)

signifdt = signifdt[variable != '(Intercept)',]
signifdt[, significant := pval < 0.05]

ggplot(signifdt, aes(x = metric, y = variable)) +
    geom_tile(data = signifdt[is.na(significant),]
            , color = 'gray', size = 1, fill = 'gray') +
    geom_tile(data = signifdt[pval >= 0.1 ,]
            , color = 'gray', size = 1, fill = NA) +
    geom_tile(data = signifdt[pval < 0.1,]
            , color = 'darkgray', size = 2, fill = NA) +
    geom_tile(data = signifdt[pval < 0.05,]
            , color = 'black', size = 2, fill = NA) +
    geom_point(data = signifdt[!is.na(significant),]
             , aes(size = Fval, fill = Fval)
             , shape = 21) +
    scale_x_discrete(expand = c(0,0)) +
    scale_y_discrete(expand = c(0,0)) +
    scale_size_continuous(range = c(0,15), limits = c(0, 20), breaks = c(0:4)*5) +
    coord_equal(ratio = 1) +
    theme_pubclean() +
    theme(
        axis.title.x = element_blank()
      , axis.title.y = element_blank()
    ) +
    ggtitle('LMM with KR test: Alpha diversity')

```

# Exit
```{r Save}
save(list = ls(all.names = TRUE), file = 'alpha-div.RData')
sessionInfo()
```
