---
title: "ROTA Metagenomics: Build figures"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: true
    theme: lumen
---
# Summary
This document contains the scripts used to build nice-looking figures and tables for the paper. For the code used to analyze the data, please refer to the markdown files mentioned in each section.

# Setup
In this section are the chunks that call libraries into the environment.
```{r}
## override library function to silence all package messages.
silence = function(x) suppressPackageStartupMessages(suppressWarnings({invisible({capture.output(x)})}))
library = function(...) silence({base::library(...)})

library('data.table')
library('magrittr')
library('svglite')
library('DescTools')

## load in theme elements.
source('scripts/load-palettes.R')

if (!dir.exists('figures')) dir.create('figures');

## Labels for axis reassignments.
relabely = setNames(c('M1','M2','M4') %>% factor(levels = c('M1','M2','M4'))
                  , c('NB','2mo','4mo'))

## abbreviate response categories
abbreviate.response = function(x) c('Responder' = 'R', 'Non-responder' = 'NR', 'Non.responder' = 'NR')[x]

set.seed(42)
```

# Figure 1: Study overview & RV1 responses

Build panels for extended data figure 1, some of which are used in figure 1.
```{r, fig.width = 6.5, fig.height = 7.5}
library('cowplot')
library('ggsignif')
library('ggplot2')
library('ggnewscale')
library('ggpubr')
library('svglite')
library('ggbeeswarm')
library('car')
library('emmeans')
library('lme4')


load(file="initialized-dataset.RData")

neutral_gray = '#444444'
neutral_gray_light = hex(mixcolor(alpha = 0.5, color1 = hex2RGB(neutral_gray), color2 = hex2RGB('#FFFFFF')))
neutral_gray_pale = hex(mixcolor(alpha = 0.5, color1 = hex2RGB(neutral_gray_light), color2 = hex2RGB('#FFFFFF')))
```

```{r}
load(file='alpha-div.RData')
load(file='beta-div.RData')


fm = metadata[!(age > 30 & visit == 'NB') & !is.na(Seroconverted),] ## exclude >30day nb visits
fm[!is.na(Seroconverted),table(country, visit)]

cs_unfiltered_md = fm[
     !duplicated(enrollment_id)
   & !is.na(post_rota_sero_iga)
   & !is.na(pre_rota_sero_iga)
   
    , .(enrollment_id, country
      , Seroconverted
      , seropositive_pre = pre_rota_sero_iga > 20
      , seropositive_post = post_rota_sero_iga > 20
      , rota_stool_pcr
      , rota_sero_iga_fc
      , log_rota_sero_iga_fc = log(rota_sero_iga_fc)
      , post_rota_sero_iga
      , log_post_rota_sero_iga = log(post_rota_sero_iga)
      , pre_rota_sero_iga
      , log_pre_rota_sero_iga = log(pre_rota_sero_iga)
      , delivery
      , gest_age
      , Secretor
      , inf_hispanic
      , inf_birth_weight
      , sex
      )] %>% setkey(enrollment_id)

```

## combined pre/post jitter/boxplots
```{r, fig.width = 5, fig.height = 3.5}

## Run stats
kw_pre = kruskal.test(
  cs_unfiltered_md[,pre_rota_sero_iga] ~ cs_unfiltered_md[,country]
)

kw_pre %>% print()

dunn_pre = DunnTest(
  cs_unfiltered_md[,pre_rota_sero_iga] ~ cs_unfiltered_md[,country],
  method = 'BH'
)

dunn_pre %>% print()

kw_post = kruskal.test(
  cs_unfiltered_md[,post_rota_sero_iga] ~ cs_unfiltered_md[,country]
)

kw_post %>% print()

dunn_post = DunnTest(
  cs_unfiltered_md[,post_rota_sero_iga] ~ cs_unfiltered_md[,country],
  method = 'BH'
)

dunn_post %>% print()


## pre post boxplots
cs_data_prepost = rbind(
     cs_unfiltered_md[, .(country, Seroconverted, val = log2(exp(log_pre_rota_sero_iga)), prepost = '1.pre')]
   , cs_unfiltered_md[, .(country, Seroconverted, val = log2(exp(log_post_rota_sero_iga)), prepost = '2.post')]
)

## Custom axis defs
vs = cs_data_prepost[,prepost] %>% unique()
vl = length(vs)

axl  = rep(factor(c('1.pre','2.post')), 2) %>% sort()

## identify breaks by finding duped labels, from both directions (removing ends)
idx = axl %>% rev() %>% duplicated() %>% not() %>% rev() %>% which()
brk = axl %>% duplicated() %>% not() %>% which()

## grab intersect and subtract half
sep = data.frame(sep = intersect(idx + 1, brk) - 0.5)


p3 = ggplot(cs_data_prepost) +
    
    geom_violin(aes(y = val
                   , x = interaction(Seroconverted, prepost)  %>% factor() %>% as.numeric()
                   , group = interaction(Seroconverted, prepost) 
                   , fill = Seroconverted
                   , color = Seroconverted)
                 # , outlier.shape = NA
                  ) +

    geom_beeswarm(aes(y = val
                  , x = interaction(Seroconverted, prepost) %>% factor() %>% as.numeric()
                  , color = factor(Seroconverted)
                  )
                , size = 0.75
                , dodge.width = 0.5
                ) +
    
    scale_color_sero_outer +
    scale_fill_sero_pale +

    #geom_hline(yintercept = log2(20), linetype = 'dashed', color = neutral_gray) +
    annotate(geom = 'segment', x = 2.5, xend = 2.5, y = -1.5, yend = 6.5, color = neutral_gray) +
    annotate(geom = 'segment', x = 2.5, xend = 2.5, y = -4, yend = 9, linetype = 'dashed', color = neutral_gray) +
    xlab('') +
    scale_x_continuous(labels = setNames(c('M2','M6'), c('1.pre','2.post'))[vs], breaks = (idx + brk)/2) +
    ylim(c(-5.039786, 15)) +
    ggtitle('Pre- and post-vaccine\nanti-rotavirus serum IgA') +
    
    facet_grid(~country, scales = 'free_x') +
    ylab('Anti-rotavirus\nserum IgA (log2(AU))') +
    theme_bw() + 
    theme(plot.title = element_text(size=11)) +
    theme(panel.spacing = unit(0, "lines")) +
    theme(legend.position = 'bottom'
        , legend.title.position = 'top'
        , panel.grid = element_blank()
        , axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

## Change facet label colors
pdf(file = NULL)
g = ggplot_gtable(ggplot_build(p3))
dev.off()

stripr <- which(grepl('strip-t', g$layout$name))
fills <- col_pal_country_light

k = 1
for (i in stripr) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k = k+1
}

pre_post_sero_boxplots = ggplotify::as.ggplot(g)

plot(pre_post_sero_boxplots)
```

## foldchange boxplots2
```{r, fig.width = 3, fig.height = 3.5}

kw_fc = kruskal.test(
  cs_unfiltered_md[,log_rota_sero_iga_fc],
  cs_unfiltered_md[,factor(country)])

dunn_fc = DunnTest(
  cs_unfiltered_md[,log_rota_sero_iga_fc],
  cs_unfiltered_md[,factor(country)],
  method = 'BH')

kw_fc %>% print()
dunn_fc %>% print()

pfc = ggplot(cs_unfiltered_md) +
    geom_violin(aes(y = log2(exp(log_rota_sero_iga_fc))
                   , x = country
                   , color = Seroconverted
                   , fill = Seroconverted)) +
        
    #scale_color_country_outer +
    #scale_fill_country_pale +
    
    #new_scale_color() +
    #new_scale_fill() +
    
    geom_hline(yintercept = log2(3), linetype = 'dashed', color = neutral_gray) +

    geom_beeswarm(aes(y = log2(exp(log_rota_sero_iga_fc))
                  , x = country
                  , color = Seroconverted
                  , fill = Seroconverted
                  )
                , size = 0.75
                , dodge.width = 1
                ) +
    scale_color_sero_outer +
    scale_fill_sero_inner +
    ylim(c(-12, 12)) +
    
    xlab('') +
    ylab('Anti-rotavirus serum IgA\nlog2-foldchange') +
    ggtitle('Anti-rotavirus serum IgA foldchange') +
    theme_bw() + 
    theme(plot.title = element_text(size=11)) + theme(legend.position = 'none'
                     , legend.title.position = 'top'
                     , panel.grid = element_blank()) +
    
    theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5)
        , axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

fc_boxplots = pfc
plot(fc_boxplots)
```

## Tabulation of N's per country / seroconversion status.
```{r, fig.width = 4, fig.height = 2}
## Tabulate
source('table_header_bg.R') ## Functions for modification of table headers
xt = fm[!is.na(Seroconverted), table(country, visit)] %>%
    cbind('') %>%
    cbind(cs_unfiltered_md[,table(country, Seroconverted)])

xtabs = xt
colnames(xtabs)[5:6] = c('Non','Sero')
colnames(xtabs)[1:3] = relabely[colnames(xtabs)[1:3]] %>% as.character()

cs_unfiltered_md[,table(country, Seroconverted)] %>% chisq.test()
metadata[,table(country, visit)] %>% chisq.test()

tbody.style = tbody_style(color = "black",
   fill = c("white"), hjust=1, x=0.9)

plot_xtab = ggtexttable(xtabs,
           theme = ttheme(
             tbody.style = tbody.style
           )) %>%
    table_cell_bg(row = 2, column = (1:ncol(xtabs))+1
                  , fill= col_pal_country_pale[rownames(xtabs)[1]], color = 'white') %>%
    table_cell_bg(row = 3, column = (1:ncol(xtabs))+1
                  , fill= col_pal_country_pale[rownames(xtabs)[2]], color = 'white') %>%
    table_cell_bg(row = 4, column = (1:ncol(xtabs))+1
                  , fill= col_pal_country_pale[rownames(xtabs)[3]], color = 'white') %>%
    table_cell_bg(row = 2:4, column = 4+1
                  , fill= 'white', color = 'white') %>%
    
    table_cell_bg(row = (1:nrow(xtabs))+1, column = 6
              , fill =  col_pal_sero_pale[colnames(xt)[5]], color = 'white') %>%
    table_cell_bg(row = (1:nrow(xtabs))+1, column = 7
              , fill =  col_pal_sero_pale[colnames(xt)[6]], color = 'white') %>%

    table_header_bg(column = 5, fill = 'white', color = 'white') %>% 
    table_header_bg(column = 6, fill = col_pal_sero_dark[colnames(xt)[5]], color = 'white') %>%
    table_header_bg(column = 7, fill = col_pal_sero_dark[colnames(xt)[6]], color = 'white')

print(plot_xtab)

## stats
cs_unfiltered_md[,table(Seroconverted, country)] %>% chisq.test()
cs_unfiltered_md[,table(seropositive_pre, country)] %>% chisq.test()
cs_unfiltered_md[,table(rota_stool_pcr, country)] %>% chisq.test()
cs_unfiltered_md[,table(seropositive_post, country)] %>% chisq.test()
```

```{r, fig.width = 4, fig.height = 2}
## Tabulate
xt2 = cbind(cs_unfiltered_md[,table(country, seropositive_pre)]) %>%
    cbind('') %>%
    cbind(cs_unfiltered_md[,table(country, seropositive_post)]) %>%
    cbind('') %>%
    cbind(cs_unfiltered_md[,table(country, rota_stool_pcr)])

xtabs = xt2
colnames(xtabs)[1:2] = c('-','+')
colnames(xtabs)[4:5] = c('-','+')
colnames(xtabs)[7:8] = c('-','+')

tbody.style = tbody_style(color = "black", size = 10, hjust=1, x=0.9)

plot_xtab_addn = ggtexttable(xtabs,
           theme = ttheme(
             tbody.style = tbody.style
           )) %>%
    attr("ggtexttableGrob") %>% 
    gtable_add_rows(heights = unit(2, "lines"), pos = 0) %>%
    
    add_spanner_label(2,3,label = 'M2\nSeropos', gp = gpar(fontsize = 10, fontface = "bold")) %>%
    add_spanner_label(5,6,label = 'M6\nSeropos', gp = gpar(fontsize = 10, fontface = "bold")) %>%
    add_spanner_label(8,9,label = 'M4\nShedding', gp = gpar(fontsize = 10, fontface = "bold")) %>%
    
    table_header_bg(column = 4, fill = 'white', color = 'white') %>% 
    table_header_bg(column = 7, fill = 'white', color = 'white')

plot(plot_xtab_addn)
```

## Pre and post IgA scatterplots + rho
```{r, fig.width = 6.5, fig.height = 4}
## Pre vs post per country
## Calc spearman's corr for pre vs post
pre_vs_post_rhos = cs_unfiltered_md[order(country),] %>%
    split(by = 'country') %>%
    lapply(function(x) cor.test(x[,log_pre_rota_sero_iga], x[,log_post_rota_sero_iga], method = 'spearman', exact = FALSE) %>% broom::tidy()) %>%
    do.call('rbind',.)
pre_vs_post_rhos$country = cs_unfiltered_md[order(country), country] %>% unique()

## Exponent notation
pre_vs_post_rhos = pre_vs_post_rhos %>%
    mutate(pval = ifelse(p.value < 0.01
                         , '<0.01'
                         , paste0('= ', as.character(signif(p.value, digits=3))))) %>%
    mutate(labs =  paste0('  P ', pval, '\n','  rho = ', round(estimate, 2)))

p2 = ggplot(data = cs_unfiltered_md, aes(x = log2(exp(log_pre_rota_sero_iga))
                                       , y = log2(exp(log_post_rota_sero_iga)))) +
    geom_smooth(method = "loess", level = 0.95, aes(color = Seroconverted, fill = Seroconverted), family = 'gaussian', size= 0.75) +
    #geom_smooth(aes(group = interaction(country, Seroconverted))
    #        , method = "lm", se = FALSE, size = 0.75, color = neutral_gray) + 
    geom_point(aes(color = Seroconverted), size = 0.75) +
    scale_color_sero_outer +
    scale_fill_sero_pale +
    geom_text(data = pre_vs_post_rhos
              , aes(label = labs)
              , x = log2(exp(-4))
              , y = log2(exp(7.5))
              , hjust = 0, vjust = 1
              , lineheight = 0.95
              , color = neutral_gray) +
    facet_grid(~country) +
    coord_equal(ratio = 1) +
    xlab('Anti-RV serum IgA at M2 (log2(AU))') +
    ylab('Anti-RV serum IgA\nat M6 (log2(AU))') +
    geom_hline(yintercept = log2(20), linetype = 'dashed', color = neutral_gray) +
    geom_vline(xintercept = log2(20), linetype = 'dashed', color = neutral_gray) +
    theme_bw() + 
    theme(plot.title = element_text(size=11)) + 
    theme(panel.spacing = unit(0, "lines")) +
    theme(legend.position = 'left', legend.title.position = 'top', panel.grid = element_blank()) 

## Change facet label colors
pdf(file = NULL)
    g = ggplot_gtable(ggplot_build(p2))
dev.off()

stripr <- which(grepl('strip-t', g$layout$name))
fills <- col_pal_country_light

k = 1
for (i in stripr) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k = k+1
}

pre_post_iga_scatters = ggplotify::as.ggplot(g)

plot(pre_post_iga_scatters)

fwrite(cs_unfiltered_md[,.(country
                                 , M2_iga = log2(exp(log_pre_rota_sero_iga))
                                 , M6_iga = log2(exp(log_post_rota_sero_iga)))], 'source-data/ed-fig1/pre-vs-post-iga.txt.xls', sep='\t')

```
```{r, fig.width = 6.5, fig.height = 4}
## Pre vs post per country
## Calc spearman's corr for pre vs post
pre_vs_post_rhos = cs_unfiltered_md[order(country),] %>%
    (function(x) cor.test(x[,log_pre_rota_sero_iga], x[,log_post_rota_sero_iga], method = 'spearman', exact = FALSE) %>% broom::tidy())

## Exponent notation
pre_vs_post_rhos = pre_vs_post_rhos %>%
    mutate(pval = ifelse(p.value < 0.01
                         , '<0.01'
                         , paste0('= ', as.character(signif(p.value, digits=3))))) %>%
    mutate(labs =  paste0('  P ', pval, '\n','  rho = ', round(estimate, 2)))

p2 = ggplot(data = cs_unfiltered_md, aes(
    x = log2(exp(log_pre_rota_sero_iga))
  , y = log2(exp(log_post_rota_sero_iga)))) +
    geom_smooth(method = "loess", level = 0.95, aes(color = Seroconverted, fill = Seroconverted), family = 'gaussian', size= 0.75) +
    #geom_smooth(aes(group = interaction(country, Seroconverted))
    #        , method = "lm", se = FALSE, size = 0.75, color = neutral_gray) + 
    geom_point(aes(color = Seroconverted), size = 0.75) +
    scale_color_sero_outer +
    scale_fill_sero_pale +
    geom_text(data = pre_vs_post_rhos
              , aes(label = labs)
              , x = log2(exp(-4))
              , y = log2(exp(7.5))
              , hjust = 0, vjust = 1
              , lineheight = 0.95
              , color = neutral_gray) +
    coord_equal(ratio = 1) +
    xlab('Anti-RV serum IgA at M2 (log2(AU))') +
    ylab('Anti-RV serum IgA\nat M6 (log2(AU))') +
    geom_hline(yintercept = log2(20), linetype = 'dashed', color = neutral_gray) +
    geom_vline(xintercept = log2(20), linetype = 'dashed', color = neutral_gray) +
    theme_bw() + 
    theme(plot.title = element_text(size=11)) + 
    theme(panel.spacing = unit(0, "lines")) +
    theme(legend.position = 'right', legend.title.position = 'top', panel.grid = element_blank()) 

pre_post_iga_scatters_total = p2

plot(pre_post_iga_scatters_total)

```


## Microbiome sampling events, box/jitter
```{r, fig.width = 5, fig.height = 2.5}
## Build event table
event_tabs = filtered_metadata[,.(enrollment_id, Seroconverted, event = 'sgmg', age, country, visit)] %>% na.omit()

## Additional violins
p3 = ggplot() +
    geom_violin(data = event_tabs[event == 'sgmg' & visit == 'NB',]
               , aes(x = age, y = Seroconverted, fill = visit, color = visit)
               , outlier.shape = NA) +
    
    geom_violin(data = event_tabs[event == 'sgmg' & visit == '2mo',]
               , aes(x = age, y = Seroconverted, fill = visit, color = visit)
               , outlier.shape = NA) +
    
    geom_violin(data = event_tabs[event == 'sgmg' & visit == '4mo',]
               , aes(x = age, y = Seroconverted, fill = visit, color = visit)
               , outlier.shape = NA) +
        
    geom_beeswarm(data = event_tabs[event == 'sgmg',]
               , aes(x = age, y = Seroconverted, color = visit)
               , size = 0.8) +
    
    facet_grid(country~., scales = 'free', space = 'free') +
    theme_bw() + 
    theme(plot.title = element_text(size=11)) +
    xlab('Age at stool sample collection') +
    scale_x_continuous(limits = c(0, 180), expand = c(0,0)) +
    scale_y_discrete(labels = abbreviate.response(names(col_pal_sero_dark))) +
    scale_color_visit_outer +
    scale_fill_visit_pale +
    geom_vline(xintercept = c(30, 100), color = neutral_gray, linetype = 'dashed') +
    theme(
        , legend.position = 'none'
        , panel.grid = element_blank()
        #, plot.margin = margin(0, 0, 0, 0, "cm")
        ) +
    theme(panel.spacing = unit(0, "lines"))

## Change facet label colors
pdf(file = NULL)
    g = ggplot_gtable(ggplot_build(p3))
dev.off()

stripr <- which(grepl('strip-r', g$layout$name))
fills <- col_pal_country_light

k = 1
for (i in stripr) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k = k+1
}

sampling_event_boxplots = ggplotify::as.ggplot(g)

plot(sampling_event_boxplots)
```


## Render figure 1
```{r, fig.width = 6.5, fig.height = 3.5}
## rows of plots
r1 = plot_grid('', plot_xtab, labels = c('A','B'))

r2 = plot_grid(pre_post_sero_boxplots, fc_boxplots, ncol = 2, labels = c('C','D'), rel_widths = c(1, 1))

#r31 = plot_grid(plot_allv_ords,'',rel_heights = c(1, 0.01), nrow = 2)

#r3 = plot_grid(SR_emm_plot2, r31, rel_widths = c(1.25,2), ncol = 2)

#pg = plot_grid(r1, r2, r3, nrow = 3, rel_heights = c(0.5, 1, 1))


#svglite('figures/figure-1.svg', width = 6.5, height = 7.5, fix_text_size =  FALSE)
#    plot(pg2)
#dev.off()


svglite('figures/figure-1AB.svg', width = 6.5, height = 2.5, fix_text_size =  FALSE)
    plot(r1)
dev.off()


svglite('figures/figure-1CD.svg', width = 6.5, height = 3.5, fix_text_size =  FALSE)
    plot(r2)
dev.off()


#svglite('figures/figure-1EF.svg', width = 7.5, height = 3, fix_text_size =  FALSE)
#    plot(r3)
#dev.off()


fwrite(cs_unfiltered_md[,.(enrollment_id, country, Seroconverted, log_pre_rota_sero_iga, log_post_rota_sero_iga, rota_sero_iga_fc, rota_stool_pcr)], 'source-data/fig1/rv1-responses.txt.xls', sep='\t')

plot(r1)
plot(r2)
#plot(r3)
```


## Render extended data figure 1
```{r, fig.height = 5.5, fig.width = 6.5}
pd1 = plot_grid(pre_post_iga_scatters_total, pre_post_iga_scatters, labels = c('A','B'), ncol = 1, rel_heights = c(1.5, 1.25))
plot(pd1)

svglite('figures/figure-extended-data-additionalsero.svg', width = 6.5, height = 5.5, fix_text_size = FALSE)
    plot(pd1)
dev.off()

```


# Figure 2: Community diversity analysis overview wrt seroconversion

## Alpha diversity boxplots
```{r, fig.width = 3, fig.height = 3}

pab = ggplot(alphadivnona) +
    geom_violin(aes(y = log10(SR)
                   , x = relabely[visit]
                   #, fill = visit
                   #, color = visit
                   )
                 , fill = neutral_gray_pale
                 , color = neutral_gray
                 , outlier.shape = NA) +
    
    geom_beeswarm(aes(y = log10(SR)
                  , x = relabely[visit]
                  #, color = factor(visit)
                  )
                  , color = neutral_gray
                , size = 0.75
                , dodge.width = 0.5
                ) +
    #scale_color_visit_outer +
    #scale_fill_visit_pale +
    facet_grid(~country) +

    xlab('') +
    ylab('log10(OGU Richness)') +
    theme_bw() + 
    theme(plot.title = element_text(size=11)) + theme(legend.position = 'bottom'
                     , legend.title.position = 'top'
                     , panel.grid = element_blank()) +
    theme(panel.spacing = unit(0, "lines")) +
    theme(axis.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5)
        , axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

## Change facet label colors
pdf(file = NULL)
g = ggplot_gtable(ggplot_build(pab))
dev.off()
stripr <- which(grepl('strip-t', g$layout$name))
fills <- col_pal_country_light

k = 1
for (i in stripr) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k = k+1
}

alpha_boxplots_country = g

plot(alpha_boxplots_country)
```

## Emmeans for ogu richness
```{r, fig.width = 3, fig.height = 2.5}
lmm_modl_fit = function(data, result_term, fixedeff, randeff, label)
{
    random_int = lapply(randeff, function(x) paste0('1 | ', x)) %>% lapply(enclose)
    
    fixed_effects_terms = paste0(fixedeff, collapse = ' + ')
    
    rand_int_terms = random_int %>% 
                        unlist() %>% 
                        unique() %>% 
                        paste0(collapse = ' + ')
    
    effect_terms = paste0(fixed_effects_terms, ' + ', rand_int_terms)
    fullform = as.formula(paste0(result_term, ' ~ 1 + ', effect_terms))
    fullmodl = lmer(fullform, data = data, REML = FALSE, na.action = 'na.fail')

    return(fullmodl)
}

fixedeff = c('formula_bm', 'visit', 'abx_status', 'country', 'sex'
             , 'inf_birth_weight', 'gest_age', 'delivery'
             , 'Seroconverted', 'Secretor')

SR_modl = lmm_modl_fit(alphadivnona, 'log(SR)', fixedeff, randeff, label = 'OGU Richness')

SR_emms = emmeans(SR_modl, ~ visit + country)
SR_emm_df = as.data.frame(SR_emms)

# Plot EMMs with confidence intervals using ggplot2
SR_emm_plot = ggplot(SR_emm_df, aes(x = relabely[visit], y = emmean, group = relabely[visit])) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
            width = 0.15,
            position = position_dodge(width = 1)) +
  geom_line(aes(group = country), position = position_dodge(width = 1)) +
  geom_point(aes(color = visit), position = position_dodge(width = 1), size = 2) +
  labs(x = "Country", y = "log10(SR) emmeans",
       color = "Visit") +
  scale_color_visit_outer +
  facet_grid(~country) +
    
  xlab('') +
    ylab('log-OGU Richness emmean') +
    theme_bw() +
    theme(plot.title = element_text(size=11)) +
    theme(legend.position = 'none'
                 , legend.title.position = 'top'
                 , panel.grid = element_blank()) +
    theme(panel.spacing = unit(0, 'lines')) +
    theme(axis.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5)
    , axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

## Change facet label colors
pdf(file = NULL)
g = ggplot_gtable(ggplot_build(SR_emm_plot))
dev.off()
stripr <- which(grepl('strip-t', g$layout$name))
fills <- col_pal_country_light

k = 1
for (i in stripr) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k = k+1
}

SR_emm_plot2 = g

plot(SR_emm_plot2)

```

## Additional alpha divs plots
### SR
```{r, fig.width = 3, fig.height = 2.5}

pab2 = ggplot(alphadivnona) +
    geom_violin(aes(y = log10(SR)
                   , x = relabely[visit]
                   , fill = visit
                   , color = visit
                   )
                 #, fill = neutral_gray_pale
                 #, color = neutral_gray
                 , outlier.shape = NA) +
    
    geom_beeswarm(aes(y = log10(SR)
                  , x = relabely[visit]
                  , color = factor(visit)
                  )
                , size = 0.75
                , dodge.width = 0.5
                ) +
    scale_color_visit_outer +
    scale_fill_visit_pale +
    facet_grid(~country) +

    xlab('') +
    ylab('log10(OGU Richness)') +
    theme_bw() +
    theme(plot.title = element_text(size=11)) +
    theme(legend.position = 'none'
                     , legend.title.position = 'top'
                     , panel.grid = element_blank()) +
    theme(panel.spacing = unit(0, 'lines')) +
    theme(axis.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5)
        , axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

## Change facet label colors
pdf(file = NULL)
g = ggplot_gtable(ggplot_build(pab2))
dev.off()
stripr <- which(grepl('strip-t', g$layout$name))
fills <- col_pal_country_light

k = 1
for (i in stripr) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k = k+1
}

alpha_violins_country1 = g

plot(alpha_violins_country1)
```

### Emmeans for ogu richness
```{r, fig.width = 3, fig.height = 2.5}

SR_modl = lmm_modl_fit(alphadivnona, 'log(SR)', fixedeff, randeff, label = 'OGU Richness')

SR_emms = emmeans(SR_modl, ~ visit + country)
SR_emm_df = as.data.frame(SR_emms)

# Plot EMMs with confidence intervals using ggplot2
SR_emm_plot = ggplot(SR_emm_df, aes(x = relabely[visit], y = emmean, group = relabely[visit])) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
            width = 0.15,
            position = position_dodge(width = 1)) +
  geom_line(aes(group = country), position = position_dodge(width = 1)) +
  geom_point(aes(color = visit), position = position_dodge(width = 1), size = 2) +
  labs(x = "Country", y = "log10(SR) emmeans",
       color = "Visit") +
  scale_color_visit_outer +
  facet_grid(~country) +
    
  xlab('') +
    ylab('log-OGU Richness emmean') +
    theme_bw() +
    theme(plot.title = element_text(size=11)) +
    theme(legend.position = 'none'
                 , legend.title.position = 'top'
                 , panel.grid = element_blank()) +
    theme(panel.spacing = unit(0, 'lines')) +
    theme(axis.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5)
    , axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

## Change facet label colors
pdf(file = NULL)
g = ggplot_gtable(ggplot_build(SR_emm_plot))
dev.off()
stripr <- which(grepl('strip-t', g$layout$name))
fills <- col_pal_country_light

k = 1
for (i in stripr) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k = k+1
}

SR_emm_plot2 = g

plot(SR_emm_plot2)

```

### Simpson
```{r, fig.width = 3, fig.height = 2.5}

pab2 = ggplot(alphadivnona) +
    geom_violin(aes(y = logit(simpson)
                   , x = relabely[visit]
                   , fill = visit
                   , color = visit
                   )
                 #, fill = neutral_gray_pale
                 #, color = neutral_gray
                 , outlier.shape = NA) +
    
    geom_beeswarm(aes(y = logit(simpson)
                  , x = relabely[visit]
                  , color = factor(visit)
                  )
                , size = 0.75
                , dodge.width = 0.5
                ) +
    scale_color_visit_outer +
    scale_fill_visit_pale +
    facet_grid(~country) +

    xlab('') +
    ylab('logit(Simpson)') +
    theme_bw() +
    theme(plot.title = element_text(size=11)) +
    theme(legend.position = 'none'
                     , legend.title.position = 'top'
                     , panel.grid = element_blank()) +
    theme(panel.spacing = unit(0, 'lines')) +
    theme(axis.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5)
        , axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

## Change facet label colors
pdf(file = NULL)
g = ggplot_gtable(ggplot_build(pab2))
dev.off()
stripr <- which(grepl('strip-t', g$layout$name))
fills <- col_pal_country_light

k = 1
for (i in stripr) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k = k+1
}

alpha_violins_country2 = g

plot(alpha_violins_country2)
```


### Emmeans for simpson
```{r, fig.width = 3, fig.height = 2.5}
fixedeff = c('formula_bm', 'visit', 'abx_status', 'country', 'sex'
             , 'inf_birth_weight', 'gest_age', 'delivery'
             , 'Seroconverted', 'Secretor')

simpson_modl = lmm_modl_fit(alphadivnona, 'logit(simpson)', fixedeff, randeff)

simpson_emms = emmeans(simpson_modl, ~ visit + country)
simpson_emm_df = as.data.frame(simpson_emms)

# Plot EMMs with confidence intervals using ggplot2
simpson_emm_plot = ggplot(simpson_emm_df, aes(x = relabely[visit], y = emmean, group = relabely[visit])) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
            width = 0.15,
            position = position_dodge(width = 1)) +
  geom_line(aes(group = country), position = position_dodge(width = 1)) +
  geom_point(aes(color = visit), position = position_dodge(width = 1), size = 2) +
  labs(x = "Country", y = "logit(simpson) emmeans",
       color = "Visit") +
  scale_color_visit_outer +
  facet_grid(~country) +
    
  xlab('') +
    ylab('logit-simpson emmean') +
    theme_bw() +
    theme(plot.title = element_text(size=11)) +
    theme(legend.position = 'none'
                 , legend.title.position = 'top'
                 , panel.grid = element_blank()) +
    theme(panel.spacing = unit(0, 'lines')) +
    theme(axis.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5)
    , axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

## Change facet label colors
pdf(file = NULL)
g = ggplot_gtable(ggplot_build(simpson_emm_plot))
dev.off()
stripr <- which(grepl('strip-t', g$layout$name))
fills <- col_pal_country_light

k = 1
for (i in stripr) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k = k+1
}

simpson_emm_plot2 = g

plot(simpson_emm_plot2)

```

### PD
```{r, fig.width = 3, fig.height = 2.5}

pab3 = ggplot(alphadivnona) +
    geom_violin(aes(y = log(PD)
                   , x = relabely[visit]
                   , fill = visit
                   , color = visit)
                 , outlier.shape = NA) +
    
    geom_beeswarm(aes(y = log(PD)
                  , x = relabely[visit]
                  , color = factor(visit)
                  )
                , size = 0.75
                , dodge.width = 0.5
                ) +
    scale_color_visit_outer +
    scale_fill_visit_pale +
    facet_grid(~country) +

    xlab('') +
    ylab('log(PD)') +
    theme_bw() +
    theme(plot.title = element_text(size=11)) +
    theme(legend.position = 'none'
                     , legend.title.position = 'top'
                     , panel.grid = element_blank()) +
    theme(panel.spacing = unit(0, 'lines')) +
    theme(axis.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5)
        , axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

## Change facet label colors
pdf(file = NULL)
g = ggplot_gtable(ggplot_build(pab3))
dev.off()
stripr <- which(grepl('strip-t', g$layout$name))
fills <- col_pal_country_light

k = 1
for (i in stripr) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k = k+1
}

alpha_violins_country3 = g

plot(alpha_violins_country3)
```

### Emmeans for PD
```{r, fig.width = 3, fig.height = 2.5}
fixedeff = c('formula_bm', 'visit', 'abx_status', 'country', 'sex'
             , 'inf_birth_weight', 'gest_age', 'delivery'
             , 'Seroconverted', 'Secretor')

PD_modl = lmm_modl_fit(alphadivnona, 'log(PD)', fixedeff, randeff)

PD_emms = emmeans(PD_modl, ~ visit + country)
PD_emm_df = as.data.frame(PD_emms)

# Plot EMMs with confidence intervals using ggplot2
PD_emm_plot = ggplot(PD_emm_df, aes(x = relabely[visit], y = emmean, group = relabely[visit])) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
            width = 0.15,
            position = position_dodge(width = 1)) +
  geom_line(aes(group = country), position = position_dodge(width = 1)) +
  geom_point(aes(color = visit), position = position_dodge(width = 1), size = 2) +
  labs(x = "Country", y = "log(PD) emmeans",
       color = "Visit") +
  scale_color_visit_outer +
  facet_grid(~country) +
    
    xlab('') +
    ylab('log-PD emmean') +
    theme_bw() +
    theme(plot.title = element_text(size=11)) +
    theme(legend.position = 'none'
                 , legend.title.position = 'top'
                 , panel.grid = element_blank()) +
    theme(panel.spacing = unit(0, 'lines')) +
    theme(axis.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5)
    , axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

## Change facet label colors
pdf(file = NULL)
g = ggplot_gtable(ggplot_build(PD_emm_plot))
dev.off()
stripr <- which(grepl('strip-t', g$layout$name))
fills <- col_pal_country_light

k = 1
for (i in stripr) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k = k+1
}

PD_emm_plot2 = g

plot(PD_emm_plot2)

```

### EMR
```{r, fig.width = 3, fig.height = 2.5}

pab4 = ggplot(alphadivnona) +
    geom_violin(aes(y = log(EMR)
                   , x = relabely[visit]
                   , fill = visit
                   , color = visit)
                 , outlier.shape = NA) +
    
    geom_beeswarm(aes(y = log(EMR)
                  , x = relabely[visit]
                  , color = factor(visit)
                  )
                , size = 0.75
                , dodge.width = 0.5
                ) +
    scale_color_visit_outer +
    scale_fill_visit_pale +
    facet_grid(~country) +

    xlab('') +
    ylab('Eff. Micr. Rich.') +
    theme_bw() +
    theme(plot.title = element_text(size=11)) +
    theme(legend.position = 'none'
                     , legend.title.position = 'top'
                     , panel.grid = element_blank()) +
    theme(panel.spacing = unit(0, 'lines')) +
    theme(axis.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5)
        , axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

## Change facet label colors
pdf(file = NULL)
g = ggplot_gtable(ggplot_build(pab4))
dev.off()
stripr <- which(grepl('strip-t', g$layout$name))
fills <- col_pal_country_light

k = 1
for (i in stripr) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k = k+1
}

alpha_violins_country4 = g

plot(alpha_violins_country4)
```

### Emmeans for EMR
```{r, fig.width = 3, fig.height = 2.5}
fixedeff = c('formula_bm', 'visit', 'abx_status', 'country', 'sex'
             , 'inf_birth_weight', 'gest_age', 'delivery'
             , 'Seroconverted', 'Secretor')

EMR_modl = lmm_modl_fit(alphadivnona, 'log(EMR)', fixedeff, randeff)

EMR_emms = emmeans(EMR_modl, ~ visit + country)
EMR_emm_df = as.data.frame(EMR_emms)

# Plot EMMs with confidence intervals using ggplot2
EMR_emm_plot = ggplot(EMR_emm_df, aes(x = relabely[visit], y = emmean, group = relabely[visit])) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
            width = 0.15,
            position = position_dodge(width = 1)) +
  geom_line(aes(group = country), position = position_dodge(width = 1)) +
  geom_point(aes(color = visit), position = position_dodge(width = 1), size = 2) +
  labs(x = "Country", y = "log(EMR) emmeans",
       color = "Visit") +
  scale_color_visit_outer +
  facet_grid(~country) +
    
    xlab('') +
    ylab('log-EMR emmean') +
    theme_bw() +
    theme(plot.title = element_text(size=11)) +
    theme(legend.position = 'none'
                 , legend.title.position = 'top'
                 , panel.grid = element_blank()) +
    theme(panel.spacing = unit(0, 'lines')) +
    theme(axis.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5)
    , axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

## Change facet label colors
pdf(file = NULL)
g = ggplot_gtable(ggplot_build(EMR_emm_plot))
dev.off()
stripr <- which(grepl('strip-t', g$layout$name))
fills <- col_pal_country_light

k = 1
for (i in stripr) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k = k+1
}

EMR_emm_plot2 = g

plot(EMR_emm_plot2)

```

### Shannon
```{r, fig.width = 3, fig.height = 3.5}

pab5 = ggplot(alphadivnona) +
    geom_violin(aes(y = shannon
                   , x = relabely[visit]
                   , fill = visit
                   , color = visit)
                 , outlier.shape = NA) +
    
    geom_beeswarm(aes(y = shannon
                  , x = relabely[visit]
                  , color = factor(visit)
                  )
                , size = 0.75
                , dodge.width = 0.5
                ) +
    scale_color_visit_outer +
    scale_fill_visit_pale +
    facet_grid(~country) +

    xlab('') +
    ylab('Shannon') +
    theme_bw() +
    theme(plot.title = element_text(size=11)) +
    theme(legend.title.position = 'top'
                     , legend.position = 'bottom'
                     , panel.grid = element_blank()) +
    theme(panel.spacing = unit(0, 'lines')) +
    theme(axis.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5)
        , axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

## Change facet label colors
pdf(file = NULL)
g = ggplot_gtable(ggplot_build(pab5))
dev.off()
stripr <- which(grepl('strip-t', g$layout$name))
fills <- col_pal_country_light

k = 1
for (i in stripr) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k = k+1
}

alpha_violins_country5 = g

plot(alpha_violins_country5)
```

### Emmeans for Shannon
```{r, fig.width = 3, fig.height = 2.5}
fixedeff = c('formula_bm', 'visit', 'abx_status', 'country', 'sex'
             , 'inf_birth_weight', 'gest_age', 'delivery'
             , 'Seroconverted', 'Secretor')

Shannon_modl = lmm_modl_fit(alphadivnona, 'shannon', fixedeff, randeff)

Shannon_emms = emmeans(Shannon_modl, ~ visit + country)
Shannon_emm_df = as.data.frame(Shannon_emms)

# Plot EMMs with confidence intervals using ggplot2
Shannon_emm_plot = ggplot(Shannon_emm_df, aes(x = relabely[visit], y = emmean, group = relabely[visit])) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
            width = 0.15,
            position = position_dodge(width = 1)) +
  geom_line(aes(group = country), position = position_dodge(width = 1)) +
  geom_point(aes(color = visit), position = position_dodge(width = 1), size = 2) +
  labs(x = "Country", y = "Shannon emmeans",
       color = "Visit") +
  scale_color_visit_outer +
  facet_grid(~country) +
    
    xlab('') +
    ylab('Shannon emmean') +
    theme_bw() +
    theme(plot.title = element_text(size=11)) +
    theme(legend.position = 'none'
                 , legend.title.position = 'top'
                 , panel.grid = element_blank()) +
    theme(panel.spacing = unit(0, 'lines')) +
    theme(axis.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5)
    , axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

## Change facet label colors
pdf(file = NULL)
g = ggplot_gtable(ggplot_build(Shannon_emm_plot))
dev.off()
stripr <- which(grepl('strip-t', g$layout$name))
fills <- col_pal_country_light

k = 1
for (i in stripr) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k = k+1
}

Shannon_emm_plot2 = g

plot(Shannon_emm_plot2)

```

## Alpha divs supp fig
```{r, fig.height = 10, fig.width = 6}
alpha_divs_supp_fig = plot_grid(
          plot_grid(alpha_violins_country1, SR_emm_plot2)
        , plot_grid(alpha_violins_country2, simpson_emm_plot2)
        , plot_grid(alpha_violins_country3, PD_emm_plot2)
        , plot_grid(alpha_violins_country4, EMR_emm_plot2)
        , plot_grid(alpha_violins_country5, plot_grid(Shannon_emm_plot2,'', rel_heights = c(6, 2), ncol = 1))
        , labels = c('A', 'B', 'C', 'D', 'E')
        , nrow = 5, ncol = 1, rel_heights = c(2, 2,
                                              2, 2, 3)
        )


svglite('figures/figure-supp-alpha.svg', width = 6.5, height = 10, fix_text_size = FALSE)
    plot(alpha_divs_supp_fig)
dev.off()
```

## Beta diversity ordinations, longitudinal
```{r, fig.width = 5, fig.height = 3}
library('ape')
library('lme4')
library('emmeans')
library('vegan')

ords = ape::pcoa(vegdist(spp_clr[rownames(rdacounts),], method = 'euclidean'))
setkey(rdamd, sample_name)
ord_data = cbind(ords$vectors, rdamd[rownames(ords$vectors),])

ord_data[,visit.centroid.Axis.1 := median(Axis.1), by = .(visit)]
ord_data[,visit.centroid.Axis.2 := median(Axis.2), by = .(visit)]
ord_data[,country.centroid.Axis.1 := median(Axis.1), by = .(country)]
ord_data[,country.centroid.Axis.2 := median(Axis.2), by = .(country)]

## Background ellipses
p1 = ggplot(data = ord_data) +
    stat_ellipse(mapping = aes(fill = visit, x = Axis.1, y = Axis.2)
                 , geom = "polygon", alpha = 0.15, show.legend = FALSE)
    
## Add spiders
#for(vi in ord_data[,visit] %>% unique()) p1 = p1 +
#    geom_segment(data = ord_data[visit == vi,]
#               , mapping = aes(
#                   x = Axis.1, xend = visit.centroid.Axis.1
#                 , y = Axis.2, yend = visit.centroid.Axis.2
#                 , color = visit)
#               , alpha = 0.25
#               , size = 0.65)

## Add in points
p1 = p1 + geom_point(mapping = aes(
               x = Axis.1
             , y = Axis.2
             , color = visit)
             , size = 0.8)

## Centroids
for(vi in ord_data[,visit] %>% unique()) p1 = p1 +
    geom_point(data = ord_data[visit == vi,][!duplicated(visit),]
             , mapping = aes(
                 x = visit.centroid.Axis.1
               , y = visit.centroid.Axis.2
               , fill = visit)
             , color = neutral_gray
             , size = 2.5
             , shape = 21)

## Pvalues for this var
p1 = p1 + annotate(geom = 'text'
                 , x = -55, y = 55
                 , label = 'P <0.01\nF = 2.15'
                 , hjust = 0, vjust = 1
              , lineheight = 0.95
              , color = neutral_gray)

## Theme elements
p1 = p1 +
    scale_color_visit_outer +
    scale_fill_visit_outer +
    
    xlab('Axis1') +
    ylab('Axis2') +
    theme_bw() + 
    theme(plot.title = element_text(size=11), legend.title = element_text(size=11)) +
    ylim(c(-55, 55)) +
    xlim(c(-55, 55)) +
    coord_equal() +
    theme(legend.position = 'bottom'
        , legend.title.position = 'top'
        , panel.grid = element_blank()
        , legend.text = element_text(margin = margin(l=unit(0.05, 'cm')))
        , legend.spacing.y = unit(0,'cm')
        , legend.key.spacing.x = unit(0.1,'cm')
        )

## By country

## Background ellipses
p2 = ggplot(data = ord_data) +
    stat_ellipse(mapping = aes(fill = country, x = Axis.1, y = Axis.2)
              , geom = "polygon", alpha = 0.15, show.legend = FALSE)
    
## Add spiders
#for(co in ord_data[,country] %>% unique()) p2 = p2 +
#    geom_segment(data = ord_data[country == co,]
#               , mapping = aes(
#                   x = Axis.1, xend = country.centroid.Axis.1
#                 , y = Axis.2, yend = country.centroid.Axis.2
#                 , color = country)
#               , alpha = 0.25
#               , size = 0.65)

## Add in points
p2 = p2 + geom_point(mapping = aes(
               x = Axis.1
             , y = Axis.2
             , color = country)
             , size = 0.8)

## Centroids
for(co in ord_data[,country] %>% unique()) p2 = p2 +
    geom_point(data = ord_data[country == co,][!duplicated(country),]
             , mapping = aes(
                 x = country.centroid.Axis.1
               , y = country.centroid.Axis.2
               , fill = country)
             , color = neutral_gray
             , size = 2.5
             , shape = 21)

p2 = p2 + annotate(geom = 'text'
                 , x = -55, y = 55
                 , label = 'P <0.01\n F = 3.36'
                 , hjust = 0, vjust = 1
                 , lineheight = 0.95
                 , color = neutral_gray)

## Theme elements
p2 = p2 +
    scale_color_country_outer +
    scale_fill_country_outer +
    
    xlab('Axis1') +
    ylab('') +
    ylim(c(-55, 55)) +
    xlim(c(-55, 55)) +
    theme_bw() +
    coord_equal() +
    theme(plot.title = element_text(size=11), legend.title = element_text(size=11)) +
    theme(legend.position = 'bottom'
        , legend.title.position = 'top'
        , panel.grid = element_blank()
        , legend.text = element_text(margin = margin(l=unit(0.05, 'cm')))
        , legend.spacing.y = unit(0,'cm')
        , legend.key.spacing.x = unit(0.1,'cm')
        )

allv_ords_visit = p1 + ggtitle('Aitchison PCoA, Month')
allv_ords_country = p2 + ggtitle('Aitchison PCoA, Country')
plot_allv_ords = plot_grid(p1 + ggtitle('Aitchison PCoA, Month')
                         , p2 + ggtitle('Aitchison PCoA, Country'))
## All aitchison ordination plots
plot(plot_allv_ords)
```


## Emmeans for ogu richness + sero
```{r, fig.width = 3, fig.height = 2.5}
vs = alphadivdata[,Seroconverted] %>% unique() %>% na.omit()
vl = length(vs)

axl  = rep(abbreviate.response(vs) %>% unique(), 2) %>% sort()

## identify breaks by finding duped labels, from both directions (removing ends)
idx = axl %>% rev() %>% duplicated() %>% not() %>% rev() %>% which()
brk = axl %>% duplicated() %>% not() %>% which()

sero_modl = lmm_modl_fit(alphadivnona, 'log(SR)', fixedeff, randeff, label = 'OGU Richness')

sero_emms = emmeans(sero_modl, ~ visit + country + Seroconverted)
sero_emm_df = as.data.frame(sero_emms)

# Plot EMMs with confidence intervals using ggplot2
sero_emm_plot = ggplot(sero_emm_df, aes(x = interaction(visit, abbreviate.response(Seroconverted))# %>%
                                         #  factor() %>%
                                          # as.numeric()
                                    , y = emmean
                                    , group = Seroconverted)) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
            width = 0.15,
            position = position_dodge(width = 1)) +
  geom_line(aes(group = interaction(country, Seroconverted)), position = position_dodge(width = 1)) +
  geom_point(aes(color = visit), position = position_dodge(width = 1), size = 2) +
  labs(color = "Visit") +
  scale_color_visit_outer +
  facet_grid(~country) +
    
  xlab('') +
  #scale_x_continuous(labels = abbreviate.response(vs), breaks = (idx + brk)/2) +
  ylab('Emmean') +
  theme_bw() +
  ggtitle('log-OGU Rich. emmeans') +
  theme(plot.title = element_text(size=11)) +
  theme(legend.title = element_text(size=11)) +

  theme(legend.position = 'bottom'
      , legend.title.position = 'top'
      , panel.grid = element_blank()) +
  theme(panel.spacing = unit(0, 'lines')) +
  theme(axis.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5)
      , axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

## Change facet label colors
pdf(file = NULL)
g = ggplot_gtable(ggplot_build(sero_emm_plot))
dev.off()
stripr <- which(grepl('strip-t', g$layout$name))
fills <- col_pal_country_light

k = 1
for (i in stripr) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k = k+1
}

sero_emm_plot2 = g

plot(sero_emm_plot2)



```

## Beta diversity bubbleplots
```{r, fig.width = 4, fig.height = 10.5}
library('cowplot')
library('vegan')
library('ggpubr')
library('khroma')

bubbleplot_betadivs_v = function(adolist)
{
    relabelx = setNames(
    c('Age', 'Delivery','FORM/MOM', 'Secretor', 'Solids', 'Inf. BW', 'Inf. GA', 'Country','Inf. Sex','Seroconv') %>% factor(levels = c('Seroconv', 'Country', 'Age', 'Inf. Sex', 'Delivery','Inf. BW','Inf. GA', 'FORM/MOM','Solids', 'Secretor') %>% rev())
    ,
    c('age', 'delivery', 'formula_bm','Secretor','solid_food_introduced','inf_birth_weight','gest_age','country','sex','Seroconverted')
)
    boxwh = 0.85
    timepoints = md[,visit] %>% unique()
    timepointdt = timepoints %>% lapply(
    function(x) 
    {
        z = adolist[[x]][[1]] %>% as.data.table(keep.rownames = TRUE)
        z[, timepoint := x]
    }
    ) %>% do.call('rbind', .)

    
    timepointdt = timepointdt[! (rn %in% c('Residual', 'Total')), ]
    colnames(timepointdt) = make.names(colnames(timepointdt))
    
    timepointdt[,timepoint := relabely[as.character(timepoint)]]
    timepointdt[,rn := relabelx[as.character(rn)]]
    
    ## Convert infinite F-values to zero
    timepointdt[is.na(Pr..F.), F := 0]
    ## Get negative log pval.
    timepointdt[,neglogpval := -log10(Pr..F.)]
    ## Significant entries are those with pvalue < 0.05
    timepointdt[, significant := Pr..F. < 0.05]

    ggplot(timepointdt, aes(x = timepoint, y = rn)) +
        
        
        geom_point(data = timepointdt[!is.na(significant),]
                 , aes(size = R2, color = R2)) +
        scale_colour_gradientn(colours = color("YlOrBr")(9)
                              , limits = c(0, 0.04)
                              , breaks = 1:4*(1/100)
                              #, (labels = 1:4*(1/100)) %>% as.character()
                              , guide = 'legend') +
        
        scale_size_continuous(range = c(0, 8)
                          , limits = c(0, 0.04)
                          , breaks = 1:4*(1/100)
                          #, labels = (1:4*(1/100)) %>% as.character()
                          , guide = 'legend') +
        ## background boxes
        geom_tile(data = timepointdt
                , color = col_pal_signif['n.s.']
                , size = 1.25, width=1, height=1
                , fill = NA) +
        
        ## NAs; invariant boxes
        geom_tile(data = timepointdt[is.na(significant),]
                , color = col_pal_signif['n.s.']
                , size = 1.25, width=boxwh, height=boxwh
                , fill = col_pal_signif_pale['n.s.']) +
        
        ## Nonsignificant, non invariant
        geom_tile(data = timepointdt[Pr..F. >= 0.1 ,]
                , color = col_pal_signif['n.s.']
                , size = 1.25, width=boxwh, height=boxwh
                , fill = NA) +
        
        ## Significant at 0.1
        geom_tile(data = timepointdt[Pr..F. < 0.1,]
                , color = col_pal_signif['.']
                , size = 1.25, width=boxwh, height=boxwh
                , fill = NA) +
        
        ## Significant at 0.05
        geom_tile(data = timepointdt[Pr..F. < 0.05,]
                , color = col_pal_signif['*']
                , size = 1.25, width=boxwh, height=boxwh
                , fill = NA) +
        
        scale_x_discrete(expand = c(0,0)) +
        scale_y_discrete(expand = c(0,0)) +
        coord_equal(ratio = 1) +
        theme_bw() +
        theme(plot.title = element_text(size=11)) +
        theme(
            axis.title.x = element_blank()
          , axis.title.y = element_blank()
          , panel.grid = element_blank()
          , panel.border = element_blank() #element_rect(linewidth = 1.5)
          , legend.position = 'none'
          , axis.ticks = element_blank()
          , axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)
        )
}


bubbleplot_betadivs_h = function(adolist)
{
    relabelx = setNames(
    c('Age', 'Delivery','FORM/MOM', 'Secretor', 'Solids', 'Inf. BW', 'Inf. GA', 'Country','Inf. Sex','Seroconv') %>% factor(levels = c('Seroconv', 'Country', 'Age', 'Inf. Sex', 'Delivery','Inf. BW','Inf. GA', 'FORM/MOM','Solids', 'Secretor') %>% rev())
    ,
    c('age', 'delivery', 'formula_bm','Secretor','solid_food_introduced','inf_birth_weight','gest_age','country','sex','Seroconverted')
)
    boxwh = 0.85
    timepoints = md[,visit] %>% unique()
    timepointdt = timepoints %>% lapply(
    function(x) 
    {
        z = adolist[[x]][[1]] %>% as.data.table(keep.rownames = TRUE)
        z[, timepoint := x]
    }
    ) %>% do.call('rbind', .)

    
    timepointdt = timepointdt[! (rn %in% c('Residual', 'Total')), ]
    colnames(timepointdt) = make.names(colnames(timepointdt))
    
    timepointdt[,timepoint := relabely[as.character(timepoint)]]
    timepointdt[,rn := relabelx[as.character(rn)]]
    
    ## Convert infinite F-values to zero
    timepointdt[is.na(Pr..F.), F := 0]
    ## Get negative log pval.
    timepointdt[,neglogpval := -log10(Pr..F.)]
    ## Significant entries are those with pvalue < 0.05
    timepointdt[, significant := Pr..F. < 0.05]

    ggplot(timepointdt, aes(y = timepoint, x = rn)) +
        
        
        geom_point(data = timepointdt[!is.na(significant),]
                 , aes(size = R2, color = R2)) +
        scale_colour_gradientn(colours = color("YlOrBr")(9)
                              , limits = c(0, 0.04)
                              , breaks = 1:4*(1/100)
                              #, (labels = 1:4*(1/100)) %>% as.character()
                              , guide = 'legend') +
        
        scale_size_continuous(range = c(0, 8)
                          , limits = c(0, 0.04)
                          , breaks = 1:4*(1/100)
                          #, labels = (1:4*(1/100)) %>% as.character()
                          , guide = 'legend') +
        ## background boxes
        geom_tile(data = timepointdt
                , color = col_pal_signif['n.s.']
                , size = 1.25, width=1, height=1
                , fill = NA) +
        
        ## NAs; invariant boxes
        geom_tile(data = timepointdt[is.na(significant),]
                , color = col_pal_signif['n.s.']
                , size = 1.25, width=boxwh, height=boxwh
                , fill = col_pal_signif_pale['n.s.']) +
        
        ## Nonsignificant, non invariant
        geom_tile(data = timepointdt[Pr..F. >= 0.1 ,]
                , color = col_pal_signif['n.s.']
                , size = 1.25, width=boxwh, height=boxwh
                , fill = NA) +
        
        ## Significant at 0.1
        geom_tile(data = timepointdt[Pr..F. < 0.1,]
                , color = col_pal_signif['.']
                , size = 1.25, width=boxwh, height=boxwh
                , fill = NA) +
        
        ## Significant at 0.05
        geom_tile(data = timepointdt[Pr..F. < 0.05,]
                , color = col_pal_signif['*']
                , size = 1.25, width=boxwh, height=boxwh
                , fill = NA) +
        
        scale_x_discrete(expand = c(0,0)) +
        scale_y_discrete(expand = c(0,0)) +
        coord_equal(ratio = 1) +
        theme_bw() +
        theme(plot.title = element_text(size=11)) +
        theme(
            axis.title.x = element_blank()
          , axis.title.y = element_blank()
          , panel.grid = element_blank()
          , panel.border = element_blank() #element_rect(linewidth = 1.5)
          , legend.position = 'none'
          , axis.ticks = element_blank()
          , axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)
        )
}

## Define the range of F-stats that we're going to see
range_R2s = lapply(c(br_adonis_list, eu_adonis_list, wu_adonis_list, uwu_adonis_list), function(x) x[[1]] %>% as.data.table()) %>% do.call('rbind',.) %>% na.omit() %>% dplyr::select(R2)

#x_min_range = c(0, max(range_Fs))
F_breaks = c(1.5, 4.5, 7.5)

## Bray
bubbleplot_bray = bubbleplot_betadivs_h(br_adonis_list) + ggtitle('Bray-Curtis') +
    theme(axis.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5)
        , axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

## Aitchison
bubbleplot_aitchison = bubbleplot_betadivs_h(eu_adonis_list) + ggtitle('Aitchison') +
    theme(axis.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5)
        , axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

## Weighted unifrac
bubbleplot_uwu = bubbleplot_betadivs_h(wu_adonis_list) + ggtitle('W-UniFrac') +
    theme(axis.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5)
        , axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

## Unweighted unifrac
bubbleplot_wu = bubbleplot_betadivs_h(uwu_adonis_list) + ggtitle('UW-UniFrac') +
    theme(axis.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5)
        , axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

## Legend
pdf(file = NULL)
    pl = ggpubr::get_legend(bubbleplot_bray + theme(legend.position = 'bottom'))
dev.off()

beta_bubbleplots = plot_grid(bubbleplot_bray
        , bubbleplot_aitchison
        , bubbleplot_uwu
        , bubbleplot_wu
        , pl
        , labels = c('A','B','C','D'), ncol = 1)

plot(beta_bubbleplots)

svglite('figures/extended-data-betadiv-bubbleplots.svg', width = 6, height = 10, fix_text_size = FALSE)
    plot(beta_bubbleplots)
dev.off()
```

## Beta diversity ordinations, cross-sectional
```{r, fig.width = 5, fig.height = 7.5}

## Beta diversity ordinations
plot_country_seroconversion_ords = function(x) 
{
    ord_data =  x[[2]]$data
    permanova = x[[1]] %>% as.data.table(keep.rownames = TRUE)
    
    ord_data[,sero.centroid.Axis.1 := median(Axis.1), by = .(Seroconverted)]
    ord_data[,sero.centroid.Axis.2 := median(Axis.2), by = .(Seroconverted)]
    ord_data[,country.centroid.Axis.1 := median(Axis.1), by = .(country)]
    ord_data[,country.centroid.Axis.2 := median(Axis.2), by = .(country)]
    
    ## Background ellipses
    p1 = ggplot(data = ord_data) +
        stat_ellipse(mapping = aes(fill = country, x = Axis.1, y = Axis.2), geom = "polygon", alpha = 0.15)
        
    ## Add spiders
    #for(co in ord_data[,country] %>% unique()) p1 = p1 +
    #    geom_segment(data = ord_data[country == co,]
    #               , mapping = aes(
    #                   x = Axis.1, xend = country.centroid.Axis.1
    #                 , y = Axis.2, yend = country.centroid.Axis.2
    #                 , color = country)
    #               , alpha = 0.25
    #               , size = 0.65)
    
    ## Add in points
    p1 = p1 + geom_point(mapping = aes(
                   x = Axis.1
                 , y = Axis.2
                 , color = country)
                 , size = 0.8)
    
    ## Centroids
    for(co in ord_data[,country] %>% unique()) p1 = p1 +
        geom_point(data = ord_data[country == co,][!duplicated(country),]
                 , mapping = aes(
                     x = country.centroid.Axis.1
                   , y = country.centroid.Axis.2
                   , fill = country)
                 , color = neutral_gray
                 , size = 2.5
                 , shape = 21)
    
    ## Build pvals.
    p.value = permanova[rn == 'country', `Pr(>F)`]
    f.stat = permanova[rn == 'country', `F`] %>% round(3)
    pval = ifelse(p.value < 0.01, '<0.01', paste0('= ', as.character(signif(p.value, digits=3))))
    plabs =  paste0('P ', pval, '\nF = ', f.stat)
    
    ## annotate
    p1 = p1 + annotate(geom = 'text'
                 , x = ord_data[,(min(Axis.1) %>% abs())] * ord_data[,min(Axis.1) %>% sign()]*1.2
                 , y = ord_data[,(max(Axis.2) %>% abs())] * ord_data[,max(Axis.2) %>% sign()]*1.1
                 , label = plabs
                 , hjust = 0, vjust = 1
                 , lineheight = 0.95
                 , color = neutral_gray)

    ## Theme elements
    p1 = p1 +
        scale_color_country_outer +
        scale_fill_country_outer +
        facet_grid(~relabely[visit]) +
        
        xlab('Axis1') +
        ylab('Axis2') +
        theme_bw() +
        theme(plot.title = element_text(size=11)) +
        theme(legend.position = 'none', panel.grid = element_blank())
    
    ## Change facet label colors
    pdf(file = NULL)
    g = ggplot_gtable(ggplot_build(p1))
    dev.off()
    stripr <- which(grepl('strip-t', g$layout$name))
    fills <- col_pal_visit_light[intersect(names(col_pal_visit_light), ord_data[,visit])]
    
    k = 1
    for (i in stripr) {
      j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
      g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
      k = k+1
    }
    
    pl1 = ggplotify::as.ggplot(g)

    ## By seroconversion
    
    ## Background ellipses
    p2 = ggplot(data = ord_data) +
        stat_ellipse(mapping = aes(fill = Seroconverted, x = Axis.1, y = Axis.2), geom = "polygon", alpha = 0.15)
        
    ## Add spiders
    #for(sc in ord_data[,Seroconverted] %>% unique()) p2 = p2 +
    #    geom_segment(data = ord_data[Seroconverted == sc,]
    #               , mapping = aes(
    #                   x = Axis.1, xend = sero.centroid.Axis.1
    #                 , y = Axis.2, yend = sero.centroid.Axis.2
    #                 , color = Seroconverted)
    #               , alpha = 0.25
    #               , size = 0.65)
    ## Add in points
    p2 = p2 + geom_point(mapping = aes(
                   x = Axis.1
                 , y = Axis.2
                 , color = Seroconverted)
                 , size = 0.8)
    
    ## Centroids
    for(sc in ord_data[,Seroconverted] %>% unique()) p2 = p2 +
                
        geom_point(data = ord_data[Seroconverted == sc,][!duplicated(Seroconverted),]
                 , mapping = aes(
                     x = sero.centroid.Axis.1
                   , y = sero.centroid.Axis.2
                   , fill = Seroconverted)
                 , color = neutral_gray
                 , size = 2.5
                 , shape = 21)
    
    ## Build pvals.
    p.value = permanova[rn == 'Seroconverted', `Pr(>F)`]
    f.stat = permanova[rn == 'Seroconverted', `F`] %>% round(3)
    pval = ifelse(p.value < 0.01, '<0.01', paste0('= ', as.character(signif(p.value, digits=3))))
    plabs =  paste0('P ', pval, '\nF = ', f.stat)
    
    p2 = p2 + annotate(geom = 'text'
                     , x = ord_data[,(min(Axis.1) %>% abs())] * ord_data[,min(Axis.1) %>% sign()]*1.2
                     , y = ord_data[,(max(Axis.2) %>% abs())] * ord_data[,max(Axis.2) %>% sign()]*1.1
                     , label = plabs
                     , hjust = 0, vjust = 1
                     , lineheight = 0.95
                     , color = neutral_gray)

    p2 = p2 +
        scale_color_sero_outer +
        scale_fill_sero_outer +
        facet_grid(~relabely[visit]) +
        
        xlab('Axis1') +
        ylab('Axis2') +
        theme_bw() +
        theme(plot.title = element_text(size=11)) +
        theme(legend.position = 'none', panel.grid = element_blank())
    
    ## Change facet label colors
    pdf(file = NULL)
    g = ggplot_gtable(ggplot_build(p2))
    dev.off()
    stripr <- which(grepl('strip-t', g$layout$name))
    fills <- col_pal_visit_light[intersect(names(col_pal_visit_light), ord_data[,visit])]
    
    k = 1
    for (i in stripr) {
      j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
      g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
      k = k+1
    }
    
    pl2 = ggplotify::as.ggplot(g)
    
    list(pl1, pl2)
    
}

## get data from ordination
aitchison_ordplots = lapply(eu_adonis_list, plot_country_seroconversion_ords) %>% do.call('c',.)
bray_ordplots = lapply(br_adonis_list, plot_country_seroconversion_ords) %>% do.call('c',.)
wu_ordplots = lapply(uwu_adonis_list, plot_country_seroconversion_ords) %>% do.call('c',.)
uwu_ordplots = lapply(wu_adonis_list, plot_country_seroconversion_ords) %>% do.call('c',.)

## All ordination plots
aitchison_plotgrid = plot_grid(plotlist = aitchison_ordplots, ncol = 2, nrow = 3)
bray_plotgrid = plot_grid(plotlist = bray_ordplots, ncol = 2, nrow = 3)
wu_plotgrid = plot_grid(plotlist = wu_ordplots, ncol = 2, nrow = 3)
uwu_plotgrid = plot_grid(plotlist = uwu_ordplots, ncol = 2, nrow = 3)

```

## Beta diversity ordinations, cross-sectional acPCoA (for aitchison only)
```{r, fig.width = 2.5, fig.height = 2.5}
## late revision: we don't want to use regular PCoA because our conclusion is
## about the data being highly confounded. Instead, we'll use acPCoA to get the
## point across to the reader.
library('caret')
silence(
{
    list.files('include/acPCoA/R/', pattern = '.R') %>%
        paste0('include/acPCoA/R/',.) %>%
        sapply(source)
})

set.seed(123) ## re-set the seed; I'm generating figures from blocks.
aitchison_acpcoa = lapply(1:length(eu_adonis_list),
function(i)
{
    dats = eu_adonis_list[[i]][[2]]$data
    ## get dm
    dm = eu_adonis_list[[i]][[3]]
    
    ## Dummy vars.
    confounders = dats[,.(country, delivery, formula_bm)] %>%
        dummyVars(data = ., formula = "~.") %>%
        predict(newdata=dats[,.(country, delivery, formula_bm)]) %>%
    cbind(dats[,.(age)])
    
    ## Fix the encoding of formula
    confounders[formula_bmBOTH == 1,formula_bmMOM := 1]
    confounders[formula_bmBOTH == 1,formula_bmFORMULA := 1]
    
    ## Remove redundant cols
    confounders[,formula_bmBOTH := NULL]
    confounders[,countryPanama := NULL]
    confounders[,deliveryvaginal := NULL]
    
    ## Run acPCoA.
    result_acPCoA = acPCoA(DistanceMatrix=dm
                         , ConfounderMatrix=confounders %>% as.matrix() %>% scale()
                         , nPC=2
                         , lambda=seq(0, 20, 0.05)
                         , kernel="linear"
                         , anov = F)
    
    ord_data = result_acPCoA$Xv %>% cbind(dats)
    
    ## vs. acpcoa
    message(i)
    message('vs acpcoa')
    glm(data = ord_data, factor(Seroconverted) ~ V1 + V2, family = 'binomial') %>% summary() %>% print()
    glm(data = ord_data, log(pre_rota_sero_iga) ~ V1 + V2, family = 'gaussian') %>% summary() %>% print()
    
    glm(data = ord_data, log(post_rota_sero_iga) ~ log(pre_rota_sero_iga) + V1 + V2, family = 'gaussian') %>% summary() %>% print()
        
    ## vs. pcoa
    message('vs pcoa')
    glm(data = ord_data, factor(Seroconverted) ~ Axis.1 + Axis.2, family = 'binomial') %>% summary() %>% print()
    
    glm(data = ord_data, log(pre_rota_sero_iga) ~ Axis.1 + Axis.2, family = 'gaussian') %>% summary() %>% print()
    
    glm(data = ord_data, log(post_rota_sero_iga) ~ log(pre_rota_sero_iga) + Axis.1 + Axis.2, family = 'gaussian') %>% summary() %>% print()
  
    ord_data[,sero.centroid.Axis.1 := median(V1), by = .(Seroconverted)]
    ord_data[,sero.centroid.Axis.2 := median(V2), by = .(Seroconverted)]
     
    permanova = eu_adonis_list[[i]][[1]] %>% as.data.table(keep.rownames = TRUE)
    ## Background ellipses
    p2 = ggplot(data = ord_data) +
        stat_ellipse(mapping = aes(fill = Seroconverted, x = V1, y = V2), geom = "polygon", alpha = 0.15)
        
    ## Add in points
    p2 = p2 + geom_point(mapping = aes(
                   x = V1
                 , y = V2
                 , color = Seroconverted)
                 , size = 0.8)
    
    ## Centroids
    for(sc in ord_data[,Seroconverted] %>% unique()) p2 = p2 +
                
        geom_point(data = ord_data[Seroconverted == sc,][!duplicated(Seroconverted),]
                 , mapping = aes(
                     x = sero.centroid.Axis.1
                   , y = sero.centroid.Axis.2
                   , fill = Seroconverted)
                 , color = neutral_gray
                 , size = 2.5
                 , shape = 21)
    
    ## Build pvals.
    p.value = permanova[rn == 'Seroconverted', `Pr(>F)`]
    f.stat = permanova[rn == 'Seroconverted', `F`] %>% round(3)
    pval = ifelse(p.value < 0.01, '<0.01', paste0('= ', as.character(signif(p.value, digits=3))))
    plabs =  paste0('P ', pval, '\nF = ', f.stat)
    
    p2 = p2 + annotate(geom = 'text'
                     , x = ord_data[,(min(V1) %>% abs())] * ord_data[,min(V1) %>% sign()]*1.2
                     , y = ord_data[,(max(V2) %>% abs())] * ord_data[,max(V2) %>% sign()]*1.1
                     , label = plabs
                     , hjust = 0, vjust = 1
                     , lineheight = 0.95
                     , color = neutral_gray)

    p2 = p2 +
        scale_color_sero_outer +
        scale_fill_sero_outer +
        facet_grid(~relabely[visit]) +
        
        xlab('Axis1') +
        ylab('Axis2') +
        theme_bw() +
        theme(plot.title = element_text(size=11)) +
        theme(legend.position = 'none', panel.grid = element_blank())
    
    ## Change facet label colors
    pdf(file = NULL)
    g = ggplot_gtable(ggplot_build(p2))
    dev.off()
    stripr <- which(grepl('strip-t', g$layout$name))
    fills <- col_pal_visit_light[intersect(names(col_pal_visit_light), ord_data[,visit])]
    
    k = 1
    for (i in stripr) {
      j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
      g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
      k = k+1
    }
    
    pl2 = ggplotify::as.ggplot(g)
    
    return(list(pl2, ord_data))
})
```

## Alpha diversity boxplots
```{r, fig.width = 3, fig.height = 3}
## Custom axis defs
vs = alphadivdata[,visit] %>% unique()
vl = length(vs)

axl  = rep(relabely, 2) %>% sort()

## identify breaks by finding duped labels, from both directions (removing ends)
idx = axl %>% rev() %>% duplicated() %>% not() %>% rev() %>% which()
brk = axl %>% duplicated() %>% not() %>% which()

## grab intersect and subtract half
sep = data.frame(sep = intersect(idx + 1, brk) - 0.5)

pab = ggplot(alphadivnona[!is.na(Seroconverted),]) +
    geom_boxplot(aes(y = log10(SR)
                   , x = paste0(visit, abbreviate.response(Seroconverted)) %>%
                               factor() %>%
                               as.numeric()
                   , group = interaction(visit, abbreviate.response(Seroconverted))
                   , fill = Seroconverted
                   , color = Seroconverted)
                 , outlier.shape = NA) +
        

    
    geom_beeswarm(aes(y = log10(SR)
                  , x = paste0(visit, abbreviate.response(Seroconverted)) %>%
                               factor() %>%
                               as.numeric()
                  , color = factor(Seroconverted)
                  )
                , size = 0.75
                , dodge.width = 0.5
                ) +
    
    scale_color_sero_outer +
    scale_fill_sero_pale +
    
        #geom_hline(yintercept = log2(20), linetype = 'dashed', color = neutral_gray) +
    geom_segment(data = sep
                 , aes(x = sep, xend = sep), y = 1.9, yend = 2.5, color = neutral_gray) +
    geom_segment(data = sep
                 , aes(x = sep, xend = sep), y = 1.8, yend = 2.6, color = neutral_gray
                 , linetype = 'dashed') +
    xlab('') +
    scale_x_continuous(labels = relabely[vs], breaks = (idx + brk)/2) +
    
    #facet_grid(relabely[visit]~country) +
    facet_grid(~country) +

    xlab('') +
    ylab('log10(OGU Richness)') +
    theme_bw() +
    theme(plot.title = element_text(size=11)) +
    theme(legend.position = 'bottom', legend.title.position = 'top'
                 , panel.grid = element_blank()) +
    theme(axis.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5)
        , axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

## Change facet label colors
pdf(file = NULL)
g = ggplot_gtable(ggplot_build(pab))
dev.off()
stripr <- which(grepl('strip-t', g$layout$name))
fills <- col_pal_country_light

k = 1
for (i in stripr) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k = k+1
}

stripr <- which(grepl('strip-r', g$layout$name))
fills <- col_pal_visit_light

k = 1
for (i in stripr) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k = k+1
}

alpha_boxplots_country_all = g

## just NB now
pab = ggplot(alphadivnona[visit == 'NB' & !is.na(Seroconverted),]) +
    geom_boxplot(aes(y = log10(SR)
                   , x = interaction(visit, abbreviate.response(Seroconverted))
                   , fill = Seroconverted
                   , color = Seroconverted)
                 , outlier.shape = NA) +
        

    
    geom_beeswarm(aes(y = log10(SR)
                  , x = interaction(visit, abbreviate.response(Seroconverted))
                  , color = factor(Seroconverted)
                  )
                , size = 0.75
                , dodge.width = 0.5
                ) +
    
    scale_color_sero_outer +
    scale_fill_sero_pale +
    
    #facet_grid(relabely[visit]~country) +
    facet_grid(~country) +

    xlab('') +
    ylab('log10(OGU Richness)') +
    theme_bw() +
    theme(plot.title = element_text(size=11)) +
    theme(legend.position = 'bottom', legend.title.position = 'top'
                     , panel.grid = element_blank()) +
    theme(axis.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5)
        , axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

## Change facet label colors
pdf(file = NULL)
g = ggplot_gtable(ggplot_build(pab))
dev.off()
stripr <- which(grepl('strip-t', g$layout$name))
fills <- col_pal_country_light

k = 1
for (i in stripr) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k = k+1
}

#stripr <- which(grepl('strip-r', g$layout$name))
#fills <- col_pal_visit_light['NB']

#k = 1
#for (i in stripr) {
#  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
#  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
#  k = k+1
#}

alpha_boxplots_country_NB = g

plot(alpha_boxplots_country_NB)
plot(alpha_boxplots_country_all)
```

## Alpha diversity bubbleplots
```{r, fig.width = 4, fig.height = 4.5}
## Alpha divs
boxwh = 0.85

relabelx = setNames(
    c('Visit', 'Delivery','FORM/MOM', 'Secretor', 'Solids', 'Inf. BW', 'Abx', 'Inf. GA', 'Country','Inf. Sex','Seroconv') %>% factor(levels = c('Seroconv', 'Country', 'Visit', 'Inf. Sex', 'Delivery','Inf. BW', 'Abx', 'Inf. GA', 'FORM/MOM','Solids', 'Secretor') %>% rev())
        ,
    c('visit', 'delivery', 'formula_bm','Secretor','solid_food_introduced','inf_birth_weight','abx_status', 'gest_age','country','sex','Seroconverted')
)
    
alpha_signifdt = list(
      SR_signifdt[,.(variable = rn, Fval = `F`, pval = `Pr(>F)`, metric = 'Richness')]
    , EMR_signifdt[,.(variable = rn, Fval = `F`, pval = `Pr(>F)`, metric = 'EMR')]
    , PD_signifdt[,.(variable = rn, Fval = `F`, pval = `Pr(>F)`, metric = 'PD')]
    , shannon_signifdt[,.(variable = rn, Fval = `F`, pval = `Pr(>F)`, metric = 'shannon')]
    , simpson_signifdt[,.(variable = rn, Fval = `F`, pval = `Pr(>F)`, metric = 'simpson')]
) %>% do.call('rbind', .)

alpha_signifdt = alpha_signifdt[variable != '(Intercept)',]
alpha_signifdt[, significant := pval < 0.05]

alpha_signifdt[,variable := relabelx[as.character(variable)]]
    
alpha_bubbleplot = ggplot(alpha_signifdt, aes(y = metric, x = variable)) +
        
    geom_point(data = alpha_signifdt[!is.na(significant),]
                 , aes(size = Fval, color = Fval)) +
    scale_colour_gradientn(colours = color("YlOrBr")(9)
                              , limits = c(0, 10)
                              , breaks = 1:4 * (2.5)
                              #, (labels = 1:4*(1/100)) %>% as.character()
                              , guide = 'legend') +
        
    scale_size_continuous(range = c(0, 8)
                          , limits = c(0, 10)
                          , breaks = 1:4 * (2.5)
                          #, labels = (1:4*(1/100)) %>% as.character()
                          , guide = 'legend') +
    ## background boxes
    geom_tile(data = alpha_signifdt
            , color = col_pal_signif_pale['n.s.']
            , size = 1.25, width=1, height=1
            , fill = NA) +
    
    ## NAs; invariant boxes
    geom_tile(data = alpha_signifdt[is.na(significant),]
            , color = col_pal_signif['n.s.']
            , size = 1.25, width=boxwh, height=boxwh
            , fill = col_pal_signif_pale['n.s.']) +
    
    ## Nonsignificant, non invariant
    geom_tile(data = alpha_signifdt[pval >= 0.1 ,]
            , color = col_pal_signif['n.s.']
            , size = 1.25, width=boxwh, height=boxwh
            , fill = NA) +
    
    ## Significant at 0.1
    geom_tile(data = alpha_signifdt[pval < 0.1,]
            , color = col_pal_signif['.']
            , size = 1.25, width=boxwh, height=boxwh
            , fill = NA) +
    
    ## Significant at 0.05
    geom_tile(data = alpha_signifdt[pval < 0.05,]
            , color = col_pal_signif['*']
            , size = 1.25, width=boxwh, height=boxwh
            , fill = NA) +
    
    scale_x_discrete(expand = c(0,0)) +
    scale_y_discrete(expand = c(0,0)) +
    coord_equal(ratio = 1) +
    theme_bw() +
    theme(plot.title = element_text(size=11)) +
    theme(
        axis.title.x = element_blank()
      , axis.title.y = element_blank()
      , panel.grid = element_blank()
      , panel.border = element_blank() #element_rect(linewidth = 1.5)
      , legend.position = 'right'
      , axis.ticks = element_blank()
      , axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)
    )

plot(alpha_bubbleplot)

```

## Render figure 2
```{r, fig.width = 8.5, fig.height = 8.5, eval = FALSE}

r1c1 = plot_grid(sero_emm_plot2, '', nrow = 2, rel_heights =c(1, 0.3))

r1 = plot_grid(r1c1, alpha_bubbleplot + ggtitle('Alpha diversity LMEM')
               , ncol = 2, labels = c('A','B')
               , rel_widths = c(0.5, 1))

r1 = plot_grid(r1, '', rel_widths = c(6.5, 2), ncol = 2)

#r1c2 = plot_grid(plotlist = aitchison_ordplots, labels = c('C','','','','',''), ncol = 2)
pp = plot_grid(plotlist=lapply(aitchison_acpcoa, function(x) x[[1]]), ncol = 3) 
r2c1 = plot_grid(allv_ords_visit + theme(legend.position ='none'), labels = c('C'))

r2c2 = plot_grid(pp
               , bubbleplot_aitchison +
                   theme(legend.title =element_text(size=11), legend.position = 'left') +
                   ggtitle('Aitchison PERMANOVA')
               , labels = c('D','E'), nrow = 2)

r2 = plot_grid(r2c1, r2c2, ncol = 2, rel_widths = c(1, 2.5))

fig2 = plot_grid(r1, r2, nrow = 2, rel_heights = c(2.5, 2.5))

plot(fig2)

svglite('figures/figure2-bad.svg', width = 8.5, height = 8.5, fix_text_size = FALSE)
    plot(fig2)
dev.off()

## Write source data to files.

## Alpha divs
fwrite(alpha_signifdt, 'source-data/fig2/alpha-diversity-terms.txt.xls', sep='\t') ## bubbleplots

fwrite(alphadivnona[,.(Seroconverted, country, visit, enrollment_id, SR, EMR, PD, shannon, simpson)], 'source-data/fig2/alpha-diversity.txt.xls', sep='\t') ## raw alpha divs (boxplots)

fwrite(sero_emm_df, 'source-data/fig2/alpha-div-emmeans.txt.xls',sep='\t') ## alpha emmeans

## aitchison
fwrite(ord_data[,.(Seroconverted, country, visit, enrollment_id, Axis.1, Axis.2)], 'source-data/fig2/aitchison-pcoa.txt.xls', sep='\t')

## cross sectional aitchison
rbind(aitchison_acpcoa[[1]][[2]][,.(Seroconverted, country, visit, enrollment_id, V1, V2)],
      aitchison_acpcoa[[2]][[2]][,.(Seroconverted, country, visit, enrollment_id, V1, V2)],
      aitchison_acpcoa[[3]][[2]][,.(Seroconverted, country, visit, enrollment_id, V1, V2)]) %>%

fwrite(., 'source-data/fig2/cross-sectional-aitchison-pcoa.txt.xls', sep='\t')

## Aitchison adonis2 bubbledata.
visits %>% lapply(
    function(x) 
    {
        z = eu_adonis_list[[x]][[1]] %>% as.data.table(keep.rownames = TRUE)
        z[, timepoint := x]
    }
    ) %>% do.call('rbind', .) %>%

fwrite(., 'source-data/fig2/aitchison-adonis2.txt.xls', sep='\t')
```


# Figure 3: Differential abundance
```{r}
library('cowplot')
library('vegan')
library('ggpubr')
library('khroma')

load('diffabund.RData')
load('beta-div.RData')
```

## dbRDA: boxplot & feature scores
```{r, fig.width = 6, fig.height = 2.5}
## Grab ordination data
ord_coords = dbrda_alr$CCA$wa
ord_data =  ord_coords %>% cbind(filtered_metadata[rownames(ord_coords),])
ord_data[,sero.centroid.dbRDA1 := median(dbRDA1), by = .(Seroconverted)]
ord_data[,sero.centroid.pre_iga := median(log2(pre_rota_sero_iga)), by = .(Seroconverted)]
ord_data[,sero.centroid.post_iga := median(log2(post_rota_sero_iga)), by = .(Seroconverted)]

ord_bp_data = ord_data[,.(dbRDA1, visit = 'Month 2', Seroconverted, country, val = log2(pre_rota_sero_iga))] %>%
        rbind(ord_data[,.(dbRDA1, visit = 'Month 6', Seroconverted, country, val = log2(post_rota_sero_iga))])# %>%
        #rbind(ord_data[,.(dbRDA1, visit = 'Foldchange', country, val = log2(rota_sero_iga_fc))])

g = ggplot(ord_data) +

    geom_boxplot(aes(y = dbRDA1
                   , x = abbreviate.response(Seroconverted)
                   , fill = Seroconverted
                   , color = Seroconverted
                 )
                 , outlier.shape = NA) +
    
    scale_color_sero_outer +
    scale_fill_sero_pale +
    
    geom_beeswarm(aes(y = dbRDA1
                  , x = abbreviate.response(Seroconverted)
                  , color = factor(Seroconverted)
                  )
                , size = 0.75
                , dodge.width = 0.5
                ) +
    

    facet_grid(relabely[visit]~country) +
    xlab('') +
    xlim(abbreviate.response(rev(names(col_pal_sero_dark)))) + 
    ylab('dbRDA1') +
    theme_bw() + 
    theme(plot.title = element_text(size=11)) + 
    theme(legend.position = 'bottom', legend.title.position = 'top', panel.grid = element_blank())


## Change facet label colors
pdf(file = NULL)
g = ggplot_gtable(ggplot_build(g))
dev.off()

stripr <- which(grepl('strip-t', g$layout$name))
fills <- col_pal_country_light

k = 1
for (i in stripr) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k = k+1
}


stripr <- which(grepl('strip-r', g$layout$name))
fills <- col_pal_visit_light[intersect(names(col_pal_visit_light), ord_data[,visit])]

k = 1
for (i in stripr) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k = k+1
}

dbrda_sero_jitterbox = g

## species contributions
dbrda_spp_data = sppscores[order(percent) %>% rev(),][1:10,]
dbrda_spp_data = dbrda_spp_data[order(percent),]
dbrda_spp_data[,numerator := spp %>% gsub('\\/.*','',.)]
dbrda_spp_data[,numerator := factor(numerator, levels = numerator)]

dbrda_sero_bars = ggplot(data = dbrda_spp_data) +
    geom_vline(xintercept = 0, linetype = 'dashed') +
    geom_bar(aes(x = score, y = numerator)
             , stat = 'identity'
             , color = col_pal_sero_dark['Non-responder']
             , fill = col_pal_sero_pale['Non-responder']) +
    geom_vline(xintercept = 0, linetype = 'dashed', color = neutral_gray) +
    geom_text(aes(x = score, y = numerator, label = (round(percent, 3)*100) %>% paste0(.,'%'))
            , position = position_nudge(x = -0.0025), hjust = 1, vjust = 0.5
            , color = neutral_gray, size = 3.2) +
    ylab('') +
    xlab('Axis Score') +
    theme_bw() + 
    theme(plot.title = element_text(size=11)) +
    theme(panel.grid = element_blank(), legend.position = 'none') +
    ggtitle('Contributing genera (dbRDA)')
    #facet_grid(~'dbRDA1 contributing genera')

plot_grid(dbrda_sero_jitterbox, dbrda_sero_bars, labels = c('A','B'))
```

## Differential abundance analysis plot at NB
### Taxonomic
```{r, fig.width = 4, fig.height = 5}
load('diffabund.RData')
library('ggrepel')
library('ggplotify')

## Diffabund volcano
plottingdata = maaslin2_out_NB$results %>%  dplyr::filter(metadata == 'Seroconverted') %>% as.data.table()
plottingdata[,signif := assign_stars(qval)]

plottingdata[,spp := gsub('[a-z]+\\.', '. ', feature)]
plottingdata[,feature2 := gsub('\\.', '\n', feature)]

p1 = ggplot(plottingdata) +
    geom_point(aes(x = coef, y = -log10(pval), color = signif)) +
    geom_vline(xintercept = 0, linetype = 'dashed', color = neutral_gray) +
    geom_text_repel(data = plottingdata[signif != 'n.s.' | spp == 'K. pneumoniae',]
                  , aes(x = coef, y = -log10(pval), label = spp)
                  , min.segment.length = 0
                  , size = 3) +
    #scale_y_log10() +
    ylab('-log10(pvalue)') +
    xlab('Seroconverison coef') +
    scale_color_signif + 
    theme_bw() + 
    theme(plot.title = element_text(size=11)) +
    theme(legend.title = element_text(size=11)) +
    theme(legend.position = 'bottom', legend.title.position = 'top') +
    theme(panel.grid = element_blank()) +
    ggtitle('M1 Maaslin2') #+
    #facet_grid('Month 1'~'Maaslin2, seroconversion')


## Change facet label colors
pdf(file = NULL)
g = ggplot_gtable(ggplot_build(p1))
dev.off()

stripr <- which(grepl('strip-r', g$layout$name))
fills <- col_pal_visit_light['NB']

k = 1
for (i in stripr) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k = k+1
}

diffabund_volcano = as.ggplot(g)

## Diffabund coefs with se
p2 = ggplot(data = plottingdata[signif != 'n.s.',]) +
    geom_vline(xintercept = 0, linetype = 'dashed', color = neutral_gray) +
    geom_point(aes(x = coef, y = feature2, color = signif), size = 2) +
    geom_errorbar(aes(y = feature2, xmax = coef + 2*stderr, xmin = coef - 2*stderr), width = 0.15) +
    scale_color_signif + 
    ylab('') +
    xlab('Beta coefficient') +
    theme_bw() + 
    theme(plot.title = element_text(size=11)) +
    theme(panel.grid = element_blank(), legend.position = 'none')# +
   # facet_grid('Month 1'~'Enrichment among seroconverters')

## Change facet label colors
pdf(file = NULL)
g = ggplot_gtable(ggplot_build(p2))
dev.off()

stripr <- which(grepl('strip-r', g$layout$name))
fills <- col_pal_visit_light['NB']

k = 1
for (i in stripr) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k = k+1
}

diffabund_bars = as.ggplot(g)

#diffabund_plots = g1 %>% insert_bottom(g2)
diffabund_plots = plot_grid(plotlist = list(diffabund_bars, diffabund_volcano), labels = c('C', ''), nrow = 2, rel_heights = c(1, 1.5))
plot(diffabund_plots)
```

```{r, fig.width = 3, fig.height = 4}
metacyc_nb_lps_all = fread('source-data/fig3/metacyc_lps_nb_all_results.tsv', sep='\t', header=TRUE)

## Diffabund volcano
plottingdata = metacyc_nb_lps_all %>%  dplyr::filter(metadata == 'Seroconverted') %>% as.data.table()
plottingdata[,signif := assign_stars(qval)]

plottingdata[,spp := feature]

metacyc_nb_lps_all = ggplot(plottingdata) +
    geom_point(aes(x = coef, y = -log10(pval), color = signif)) +
    geom_vline(xintercept = 0, linetype = 'dashed', color = neutral_gray) +
    geom_text_repel(data = plottingdata[signif == '*',]
                  , aes(x = coef, y = -log10(pval), label = spp)
                  , min.segment.length = 0
                  , size = 3) +
    #scale_y_log10() +
    ylab('-log10(pvalue)') +
    xlab('Seroconverison coef') +
    scale_color_signif + 
    theme_bw() + 
    theme(plot.title = element_text(size=11)) +
    theme(legend.title = element_text(size=11)) +
    theme(legend.position = 'bottom', legend.title.position = 'top') +
    theme(panel.grid = element_blank()) +
    ggtitle('MetaCyc Maaslin2 (M1)') #+
    #facet_grid('Month 1'~'Maaslin2, seroconversion')

plot(metacyc_nb_lps_all)

```

### KO-lps
```{r, fig.width = 3, fig.height = 4}
KO_lps_nb = fread('source-data/fig3/KO_lps_nb_all_results.tsv', sep='\t', header=TRUE)

## Diffabund volcano
plottingdata = KO_lps_nb %>%  dplyr::filter(metadata == 'Seroconverted') %>% as.data.table()
plottingdata[,signif := assign_stars(qval)]

plottingdata[,spp := feature]
plottingdata[,spp := gsub('Klebsiella.michiganensis.', 'kom:', spp)]
plottingdata[,spp := gsub('Klebsiella.grimontii.', 'kgr:', spp)]

ko_lps_nb_volcano = ggplot(plottingdata) +
    geom_point(aes(x = coef*-1, y = -log10(pval), color = signif)) +
    geom_vline(xintercept = 0, linetype = 'dashed', color = neutral_gray) +
    geom_text_repel(data = plottingdata[signif != 'n.s.',]
                  , aes(x = coef*-1, y = -log10(pval), label = spp)
                  , min.segment.length = 0
                  , size = 3) +
    #scale_y_log10() +
    ylab('-log10(pvalue)') +
    xlab('Maaslin2 coef') +
    scale_color_signif + 
    theme_bw() + 
    theme(plot.title = element_text(size=11)) +
    theme(legend.title = element_text(size=11)) +
    theme(legend.position = 'bottom', legend.title.position = 'top') +
    theme(panel.grid = element_blank()) +
    ggtitle('Species:KEGG Maaslin2 (M1)') #+
    #facet_grid('Month 1'~'Maaslin2, seroconversion')

plot(ko_lps_nb_volcano)

svglite('figures/figure-3D.svg', width = 3, height = 4, fix_text_size = FALSE)
    plot(ko_lps_nb_volcano)
dev.off()

#fwrite(plottingdata, 'source-data/ed-fig3/ko-lps-results.txt.xls', sep = '\t')
```

### MetaCyc-lps
```{r, fig.width = 3, fig.height = 2.5}
metacyc_nb_lps_all = fread('source-data/fig3/metacyc_lps_nb_all_results.tsv', sep='\t', header=TRUE)

## Diffabund volcano
plottingdata = metacyc_nb_lps_all %>%  dplyr::filter(metadata == 'Seroconverted') %>% as.data.table()
plottingdata[,signif := assign_stars(qval)]

plottingdata[order(coef), spp := factor(feature, levels = feature)]

metacyc_nb_lps_plot = ggplot(plottingdata) +
    
    geom_vline(xintercept = 0, linetype = 'dashed', color = neutral_gray) +
    geom_bar(aes(x = coef*-1, y = spp, fill = signif), stat = 'identity') +
    geom_errorbar(aes(y = spp, xmax = (coef + 2*stderr)*-1, xmin = (coef - 2*stderr)*-1), width = 0.15) +
    #scale_color_signif + 
    scale_fill_signif + 
    ylab('') +
    xlab('Maaslin2 coef') +
    theme_bw() + 
    ggtitle('MetaCyc LPS\nDiff. Abund.') +
    theme(plot.title = element_text(size=11)) +
    theme(panel.grid = element_blank()
        , legend.position = 'bottom'
        , legend.title.position = 'top'
        , legend.key.size = unit(0.25, 'cm')
        , legend.key.height = unit(0.25, 'cm')
        , legend.key.width = unit(0.25, 'cm'))

plot(metacyc_nb_lps_plot)

svglite('figures/figure-3C.svg', width = 3, height = 2.5, fix_text_size = FALSE)
    plot(metacyc_nb_lps_plot)
dev.off()

#fwrite(plottingdata, 'source-data/ed-fig3/metacyc-lps-results.txt.xls', sep = '\t')
```

### MetaCyc-all
```{r, fig.width = 3.5, fig.height = 7.5}
metacyc_nb_all = fread('source-data/fig3/metacyc_nb_all_results.tsv', sep='\t', header=TRUE)

## Diffabund volcano
plottingdata = metacyc_nb_all %>%  dplyr::filter(metadata == 'Seroconverted') %>% as.data.table()
plottingdata[,signif := assign_stars(qval)]

plottingdata[order(coef), spp := factor(feature, levels = feature)]

metacyc_nb_plot = ggplot(plottingdata[signif != 'n.s.',]) +
    
    geom_vline(xintercept = 0, linetype = 'dashed', color = neutral_gray) +
    geom_bar(aes(x = coef*-1, y = spp, fill = signif), stat = 'identity') +
    geom_errorbar(aes(y = spp, xmax = (coef + 2*stderr)*-1, xmin = (coef - 2*stderr)*-1), width = 0.15) +
    #scale_color_signif + 
    scale_fill_signif + 
    ylab('') +
    xlab('Maaslin2 coef') +
    theme_bw() + 
    ggtitle('MetaCyc\nDiff. Abund. (M1)') +
    theme(plot.title = element_text(size=11)) +
    theme(panel.grid = element_blank()
        , legend.position = 'bottom'
        , legend.title.position = 'top'
        , legend.key.size = unit(0.25, 'cm')
        , legend.key.height = unit(0.25, 'cm')
        , legend.key.width = unit(0.25, 'cm'))

plot(metacyc_nb_plot)

svglite('figures/extended-data-figure-3B.svg', width = 3.5, height = 7.5, fix_text_size = FALSE)
    plot(metacyc_nb_plot)
dev.off()

fwrite(plottingdata, 'source-data/ed-fig3/metacyc-results.txt.xls', sep = '\t')

```

### Diffabund-Full
```{r, fig.width = 3, fig.height = 7.5}
plottingdata = results[['global']]

## Render volcano-ish plot.
plottingdata[,signif := assign_stars(padj)]

plottingdata[,errmin := effectSize - 2*SE]
plottingdata[,errmax := effectSize + 2*SE]
signif_spp = plottingdata[padj < qval_threshold, Species] %>% unique()

errdata = plottingdata[Species %in% signif_spp,][,.(errcords = c(errmin,errmax)), by = .(Species, method)]

p1 =  ggplot(data = plottingdata) +
    geom_point(aes(x = effectSize, y = -log10(pval), color = signif)) +
    facet_wrap(method~., scales = 'free_x', ncol = 1) +
    scale_y_log10() +
    ylab('-log10(pvalue)') +
    xlab('Maaslin2 coef') +
    scale_color_signif + 
    theme_bw() + 
    theme(plot.title = element_text(size=11)) +
    theme(legend.title = element_text(size=11)) +
    theme(legend.position = 'bottom', legend.title.position = 'top') +
    theme(panel.grid = element_blank()) +
    
    ggtitle('Taxonomic Differential Abundance') #+

p2 = ggplot(data = plottingdata[Species %in% signif_spp,]) +
    geom_bar(aes(x = effectSize, y = Species, fill = signif), stat = 'identity') +
    facet_wrap(method~., scales = 'free_x', ncol = 1) +
    geom_errorbar(aes(y = Species, xmax = effectSize + 2*SE, xmin = effectSize - 2*SE), width = 0.15) +
    #scale_color_signif + 
    scale_fill_signif + 
    ylab('') +
    xlab('Maaslin2 coef') +
    theme_bw() + 
    ggtitle('Species\nDiff. Abund.') +
    theme(plot.title = element_text(size=11)) +
    theme(panel.grid = element_blank()
        , legend.position = 'bottom'
        , legend.title.position = 'top'
        , legend.key.size = unit(0.25, 'cm')
        , legend.key.height = unit(0.25, 'cm')
        , legend.key.width = unit(0.25, 'cm'))


svglite('figures/extended-data-figure-3A.svg', width = 3.5, height = 7.5, fix_text_size = FALSE)
    plot(p2)
dev.off()

fwrite(plottingdata, 'source-data/ed-fig3/maaslin-results.txt.xls', sep = '\t')
```

## Render Figure 3
```{r, fig.width = 6.5, fig.height = 3}
#s1 = plot_grid(dbrda_sero_bars, '', rel_heights = c(1, 0.25), ncol = 1)
#s2 = plot_grid(dbrda_sero_jitterbox, s1, labels = c('A','B'), nrow = 1)
#r1 = plot_grid(s2, '', nrow = 1, rel_widths = c(6, 1.5))
#r2c1 = diffabund_plots

c1 = plot_grid(dbrda_sero_bars,'', rel_heights = c(1, 0.35), nrow = 2)
c2 = diffabund_volcano

r1 = plot_grid(c1, c2, labels = c('A', 'B'), ncol = 2)

plot(r1)

plot_grid(r1, r2, nrow = 2, rel_heights = c(3, 2))

svglite('figures/figure-3AB.svg', width = 6.5, height = 2.8, fix_text_size = FALSE)
    plot(r1)
dev.off()


svglite('figures/figure-3CD.svg', width = 7, height = 5.5, fix_text_size = FALSE)
    plot(r2)
dev.off()

## dbRDA species scores
fwrite(sppscores[order(percent) %>% rev(),]
     , 'source-data/fig3/dbrda-contributing-genera.txt.xls',sep='\t')

## Differential abundance
fwrite(results[['global']], 'source-data/fig3/combined-diffabund.txt.xls', sep='\t')
fwrite(maaslin2_out_NB$results, 'source-data/fig3/m1-maaslin2.txt.xls', sep = '\t')
fwrite(maaslin2_out_2mo$results, 'source-data/fig3/m2-maaslin2.txt.xls', sep = '\t')
fwrite(maaslin2_out_4mo$results, 'source-data/fig3/m4-maaslin2.txt.xls', sep = '\t')

## Functional diffabunds are already written to the file.
```


# Figure 4: DMMs
```{r}
load(file = 'dmms.RData')
```

## Multinomial plots
```{r MultinomialPlots, fig.width = 2.5, fig.height = 5}
library('ggplot2')
library('data.table')
library('magrittr')
## Vs cs-GCT3
# > print(seroconverted_pdmm3_NB)
#                 pval        OR     coef        se signif
# cs.DMM1 0.0090270784  4.994811 1.608400 0.6160027     **
# cs.DMM2 0.0502142029  3.316546 1.198924 0.6122786      .
# cs.DMM4 0.0992615591  2.902451 1.065555 0.6464007      .
# cs.DMM5 0.0275939527  5.226373 1.653718 0.7506623      *
# cs.DMM6 0.0003025822 21.039282 3.046391 0.8431565    ***
dats = as.data.table(seroconverted_pdmm3_NB, keep.rownames=TRUE)
dats[,Significance := gsub('\\*\\*+','\\*', signif)] ## single-star significance level

sero_odds_vdmm3 = ggplot(data = dats) +
        
    ## add in error bars
    geom_errorbar(aes(x = gsub('cs.DMM','cs-GCT', rn), ymax = coef + 2*se, ymin = coef - 2*se), width = 0.15) +
    
    ## render points
    geom_point(aes(x = gsub('cs.DMM','cs-GCT', rn), y = coef, color = Significance), size = 2) +
    
    ## color scale for significance
    scale_color_signif + 
    ylab('log(OR)') +
    xlab('') +
    theme_bw() + 
    theme(plot.title = element_text(size=11)) +
    ggtitle('Seroconversion odds') +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    theme(panel.grid = element_blank()
        , legend.position = 'none'
        , axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
    facet_grid(~'M1 type vs. cs-GCT3') +
    ylim(c(-0.5,6))

## Shedding
dats = as.data.table(stool_pcr_pdmm3_NB, keep.rownames=TRUE)
dats[,Significance := gsub('\\*\\*+','\\*', signif)]

shedding_odds_vdmm3 = ggplot(data = dats) +
    geom_errorbar(aes(x = gsub('cs.DMM','cs-GCT', rn), ymax = coef + 2*se, ymin = coef - 2*se), width = 0.15) +
    geom_point(aes(x = gsub('cs.DMM','cs-GCT', rn), y = coef, color = Significance), size = 2) +
    scale_color_signif + 
    ylab('log(OR)') +
    xlab('') +
    theme_bw() + 
    theme(plot.title = element_text(size=11)) +
    ggtitle('Shedding odds') +
    geom_hline(yintercept = 0, linetype = 'dashed', color = neutral_gray) +
    theme(panel.grid = element_blank()
        , legend.position = 'none'
        , axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
    facet_grid(~'M1 type vs. cs-GCT3') +
    ylim(c(-0.5,6))

dats = as.data.table(seroconverted_pdmm6_NB, keep.rownames=TRUE)
dats[,Significance := gsub('\\*\\*+','\\*', signif)]

## Vs cs-GCT6
## Sero
sero_odds_vdmm6 = ggplot(data = dats) +
    geom_errorbar(aes(x = gsub('cs.DMM','cs-GCT', rn), ymax = coef + 2*se, ymin = coef - 2*se), width = 0.15) +
    geom_point(aes(x = gsub('cs.DMM','cs-GCT', rn), y = coef, color = Significance), size = 2) +
    scale_color_signif + 
    ylab('log(Odds Ratio) difference') +
    xlab('') +
    theme_bw() + 
    theme(plot.title = element_text(size=11)) +
    ggtitle('Seroconversion') +
    geom_hline(yintercept = 0, linetype = 'dashed', color = neutral_gray) +
    theme(panel.grid = element_blank()
        , legend.position = 'none'
        , axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
    facet_grid(~'M1 type vs. cs-GCT6') +
    ylim(c(-6,0.5))

pdf(file = NULL)
    pl = ggpubr::get_legend(sero_odds_vdmm6 + theme(legend.position = 'bottom', legend.title.position = 'top'))
dev.off()

odds_multinom = plot_grid(sero_odds_vdmm3, shedding_odds_vdmm3, nrow = 2, rel_widths = c(1, 1), align = "v", axis = "l") %>% plot_grid(., pl, rel_heights = c(1, 0.2), nrow = 2)

plot(odds_multinom)

svglite('figures/odds-multinom.svg', width = 2.5, height = 5, fix_text_size = FALSE)
    plot(odds_multinom)
dev.off()

fwrite(seroconverted_pdmm3_NB %>% as.data.table(keep.rownames = TRUE), 'source-data/fig4/multinomial-or-sero-dmm3.txt.xls', sep = '\t')
fwrite(stool_pcr_pdmm3_NB %>% as.data.table(keep.rownames = TRUE), 'source-data/fig4/multinomial-or-pcr-dmm3.txt.xls', sep = '\t')

```

## Taxon barplots
```{r}
library('ggplot2')
library('rlang')
library('ggnewscale')
```

```{r NB_taxon_bars, fig.width = 5, fig.height = 3.5}
dmm_NB_spp_melt = data.table::melt(dmm_NB_spp) %>% as.data.table()
ncoll_spp = dmm_NB_spp_melt[value > collapse_thresh, Var2] %>%
    unique() %>%
    as.character()

breaks = c(ncoll_spp, collapse_str)
n = length(ncoll_spp)
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)) %>%
    unlist()

## Seed set to 113.
set.seed(113)
colpal = sample(col_vector, n) %>%
    c(. , '#AAAAAA') %>%
    setNames(., breaks)

## set order based on relabund
breakorder = dmm_NB_spp_plot[order(-value), Var2] %>% unique() %>% as.character() %>% setdiff(collapse_str) %>% c(collapse_str)

## Phylum associations
taxon_phyl_assoc = taxonomy[g %in% breakorder,.(p, g)] %>% unique() %>% split(by = 'p')

dmm_NB_spp_plot[,csGCT := gsub('cs.DMM','M1 cs-GCT', NB)]
taxon_bars_nb_all = ggplot() + 
    geom_bar(aes(fill=factor(Var2, levels = breakorder), y=value, x=Var1)
             , color = NA, position='stack', stat='identity'
             , data = dmm_NB_spp_plot) +
    facet_wrap(~csGCT, scales = 'free_x', ncol = 2) + 
    theme_bw() + 
    theme(plot.title = element_text(size=11)) +
    theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
    theme(legend.key.size = unit(0.25, 'cm')
        , legend.key.height = unit(0.25, 'cm')
        , legend.key.width = unit(0.25, 'cm')
        , axis.ticks = element_blank()
        #, legend.title = element_text(size=14)
        #, legend.text = element_text(size=10)
        ) +
    scale_y_continuous(labels = scales::percent, expand = c(0,0)) +
    scale_x_discrete(expand = c(0,0)) +
    xlab('Individual') +
    ylab('Genus-level composition (%)') +
    scale_fill_manual(values = colpal, breaks = breakorder, name = 'Genus', guide = guide_legend(ncol = 1))


## Change facet label colors
pdf(file = NULL)
g = ggplot_gtable(ggplot_build(taxon_bars_nb_all))
dev.off()

stripr <- which(grepl('strip-t', g$layout$name))
fills <- col_pal_visit_light['NB']

k = 1
for (i in stripr) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  #k = k+1
}

taxon_bars_nb_all = as.ggplot(g)

plot(taxon_bars_nb_all)
```

```{r 2mo_taxon_bars}
dmm_2mo_spp_melt = melt(dmm_2mo_spp) %>% as.data.table()
ncoll_spp = dmm_2mo_spp_melt[value > collapse_thresh, Var2] %>%
    unique() %>%
    as.character()

breaks = c(ncoll_spp, collapse_str)
n = length(ncoll_spp)
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)) %>%
    unlist()

## use the previous palette for  consistency between plots.
diffs = setdiff(ncoll_spp, names(colpal))

if(length(diffs) != 0)
{
    col_vector = setdiff(col_vector, colpal)
    n2 = length(diffs)
    colpal2 = sample(col_vector, n2) %>%
        setNames(., names(diffs))
    
    colpal = c(colpal2, colpal)
}

## set order based on relabund
breakorder = dmm_2mo_spp_plot[order(-value), Var2] %>% unique() %>% as.character() %>% setdiff(collapse_str) %>% c(collapse_str)

dmm_2mo_spp_plot[,csGCT := gsub('cs.DMM','M2 cs-GCT', X2mo)]

taxon_bars_2mo_all = ggplot() + 
    geom_bar(aes(fill=factor(Var2, levels = breakorder), y=value, x=Var1)
             , color = NA, position='stack', stat='identity'
             , data = dmm_2mo_spp_plot) +
    facet_wrap(~csGCT, scales = 'free_x') + 
    theme_bw() + 
    theme(plot.title = element_text(size=11)) +
    theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
    theme(legend.key.size = unit(0.25, 'cm')
        , legend.key.height = unit(0.25, 'cm')
        , legend.key.width = unit(0.25, 'cm')
        , axis.ticks = element_blank()
        #, legend.title = element_text(size=14)
        #, legend.text = element_text(size=10)
        ) +
    scale_y_continuous(labels = scales::percent, expand = c(0,0)) +
    scale_x_discrete(expand = c(0,0)) +
    xlab('Individual') +
    ylab('Genus-level composition (%)') +
    scale_fill_manual(values = colpal, breaks = breakorder, name = 'Genus', guide = guide_legend(ncol = 1))


## Change facet label colors
pdf(file = NULL)
g = ggplot_gtable(ggplot_build(taxon_bars_2mo_all))
dev.off()

stripr <- which(grepl('strip-t', g$layout$name))
fills <- col_pal_visit_light['2mo']

k = 1
for (i in stripr) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  #k = k+1
}

taxon_bars_2mo_all = as.ggplot(g)

plot(taxon_bars_2mo_all)

filtered_eids = cs_dmm_data[!is.na(X2mo), enrollment_id]

dmm_2mo_sel_coll = dmm_2mo_sel_coll = dmm_2mo_spp_melt[!(Var2 %in% ncoll_spp)
                                  & Var1 %in% filtered_eids
                                  , .(Var2 = collapse_str
                                  , value =  sum(value)), by = Var1][,.(Var1, Var2, value)] %>%
    rbind(dmm_2mo_spp_melt[(Var2 %in% ncoll_spp) & Var1 %in% filtered_eids,],.)

fwrite(dmm_2mo_sel_coll, 'source-data/ed-fig4/m2-genus-composition.txt.xls', sep = '\t')
```

```{r 4mo_taxon_bars}
dmm_4mo_spp_melt = melt(dmm_4mo_spp) %>% as.data.table()
ncoll_spp = dmm_4mo_spp_melt[value > collapse_thresh, Var2] %>%
    unique() %>%
    as.character()

breaks = c(ncoll_spp, collapse_str)
n = length(ncoll_spp)
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)) %>%
    unlist()

## use the previous palette for  consistency between plots.
diffs = setdiff(ncoll_spp, names(colpal))

if(length(diffs) != 0)
{
    col_vector = setdiff(col_vector, colpal)
    n2 = length(diffs)
    colpal2 = sample(col_vector, n2) %>%
        setNames(., names(diffs))
    
    colpal = c(colpal2, colpal)
}

dmm_4mo_spp_plot[,csGCT := gsub('cs.DMM','M4 cs-GCT', X4mo)]
taxon_bars_4mo_all = ggplot() + 
    geom_bar(aes(fill=factor(Var2, levels = breakorder), y=value, x=Var1)
             , color = NA, position='stack', stat='identity'
             , data = dmm_4mo_spp_plot) +
    facet_wrap(~csGCT, scales = 'free_x') + 
    theme_bw() + 
    theme(plot.title = element_text(size=11)) +
    theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
    theme(legend.key.size = unit(0.25, 'cm')
        , legend.key.height = unit(0.25, 'cm')
        , legend.key.width = unit(0.25, 'cm')
        , axis.ticks = element_blank()
        #, legend.title = element_text(size=14)
        #, legend.text = element_text(size=10)
        ) +
    scale_y_continuous(labels = scales::percent, expand = c(0,0)) +
    scale_x_discrete(expand = c(0,0)) +
    xlab('Individual') +
    ylab('Genus-level composition (%)') +
    scale_fill_manual(values = colpal, breaks = breakorder, name = 'Genus', guide = guide_legend(ncol = 1))

## Change facet label colors
pdf(file = NULL)
g = ggplot_gtable(ggplot_build(taxon_bars_4mo_all))
dev.off()

stripr <- which(grepl('strip-t', g$layout$name))
fills <- col_pal_visit_light['4mo']

k = 1
for (i in stripr) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  #k = k+1
}

filtered_eids = cs_dmm_data[!is.na(X4mo), enrollment_id]

dmm_4mo_sel_coll = dmm_4mo_sel_coll = dmm_4mo_spp_melt[
    !(Var2 %in% ncoll_spp) & Var1 %in% filtered_eids
, .(Var2 = collapse_str, value =  sum(value)), by = Var1][,.(Var1, Var2, value)] %>%
    rbind(dmm_4mo_spp_melt[(Var2 %in% ncoll_spp) & Var1 %in% filtered_eids,],.)

fwrite(dmm_4mo_sel_coll, 'source-data/ed-fig4/m4-genus-composition.txt.xls', sep = '\t')

taxon_bars_4mo_all = as.ggplot(g)

plot(taxon_bars_4mo_all)
```

```{r, fig.width = 10, fig.height = 2.5}
library('ggh4x')
## barplots with just M1 csGCT3 & csGCT6
filtered_eids = cs_dmm_data[!is.na(NB), enrollment_id]

ncoll_spp = dmm_NB_spp_melt[value > collapse_thresh & Var1 %in% filtered_eids, Var2] %>%
    unique() %>%
    as.character()

breaks = c(ncoll_spp, collapse_str)
n = length(ncoll_spp)
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)) %>%
    unlist()

## use the previous palette for  consistency between plots.
diffs = setdiff(ncoll_spp, names(colpal))

if(length(diffs) != 0)
{
    col_vector = setdiff(col_vector, colpal)
    n2 = length(diffs)
    colpal2 = sample(col_vector, n2) %>%
        setNames(., names(diffs))
    
    colpal = c(colpal2, colpal)
}

dmm_NB_sel_coll = dmm_NB_spp_melt[!(Var2 %in% ncoll_spp)
                                  & Var1 %in% filtered_eids
                                  , .(Var2 = collapse_str
                                  , value =  sum(value)), by = Var1][,.(Var1, Var2, value)] %>%
    rbind(dmm_NB_spp_melt[(Var2 %in% ncoll_spp) & Var1 %in% filtered_eids,],.)

dmm_NB_sel_plot = setkey(copy(cs_dmm_data), enrollment_id)[dmm_NB_sel_coll[,Var1] %>% as.character(),] %>%
    cbind(dmm_NB_sel_coll,.)

## set order based on relabund
breakorder = dmm_NB_sel_plot[order(-value), Var2] %>% unique() %>% as.character() %>% setdiff(collapse_str) %>% c(collapse_str)

## Phylum associations
taxon_phyl_assoc = taxonomy[g %in% breakorder,.(p, g)] %>% unique() %>% split(by = 'p')

dmm_NB_sel_plot[!is.na(NB), csGCT := gsub('cs.DMM','M1 cs-GCT', NB)]
taxon_bars_nb_sel = ggplot() + 
    geom_bar(aes(fill=factor(Var2, levels = breakorder), y=value, x=Var1)
             , color = NA, position='stack', stat='identity'
             , data = dmm_NB_sel_plot) +
    facet_nested(.~csGCT + country, scales = 'free', space = 'free', nest_line = element_line(linetype = 1)) + 
    theme_bw() + 
    theme(panel.spacing = unit(0, "cm")) +
    theme(plot.title = element_text(size=11)) +
    theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
    theme(legend.key.size = unit(0.25, 'cm')
        , legend.key.height = unit(0.25, 'cm')
        , legend.key.width = unit(0.25, 'cm')
        , axis.ticks = element_blank()
        #, legend.title = element_text(size=14)
        #, legend.text = element_text(size=10)
        ) +
    scale_y_continuous(labels = scales::percent, expand = c(0,0)) +
    scale_x_discrete(expand = c(0,0)) +
    xlab('Individual') +
    ylab('Genus-level composition (%)') +
    scale_fill_manual(values = colpal, breaks = breakorder, name = 'Genus', guide = guide_legend(ncol = 2))

## Change facet label colors
pdf(file = NULL)
g = ggplot_gtable(ggplot_build(taxon_bars_nb_sel))
dev.off()

stripr <- which(grepl('strip-t', g$layout$name))
fills <- c(rep(col_pal_visit_light['NB'],6), rep(col_pal_country_light, 6))
colors <- c(rep(col_pal_visit_dark['NB'],6), rep(col_pal_country_dark, 6))

k = 1
for (i in stripr) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$col <- colors[k]
  k = k+1
}

taxon_bars_nb_sel = as.ggplot(g)

plot(taxon_bars_nb_sel)

svglite('figures/genus-taxon-bars.svg', width = 11, height = 2.5, fix_text_size = FALSE)
    plot(taxon_bars_nb_sel)
dev.off()

fwrite(dmm_NB_sel_coll, 'source-data/fig4/genus-level-composition.txt.xls', sep = '\t')

```

```{r, fig.width = 6.5, fig.height = 2.8}
library('ggh4x')
## barplots with just M1 csGCT3 & csGCT6
filtered_eids = cs_dmm_data[NB %in% c('cs.DMM3', 'cs.DMM6'), enrollment_id]

ncoll_spp = dmm_NB_spp_melt[value > collapse_thresh & Var1 %in% filtered_eids, Var2] %>%
    unique() %>%
    as.character()

breaks = c(ncoll_spp, collapse_str)
n = length(ncoll_spp)
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)) %>%
    unlist()

## use the previous palette for  consistency between plots.
diffs = setdiff(ncoll_spp, names(colpal))

if(length(diffs) != 0)
{
    col_vector = setdiff(col_vector, colpal)
    n2 = length(diffs)
    colpal2 = sample(col_vector, n2) %>%
        setNames(., names(diffs))
    
    colpal = c(colpal2, colpal)
}

dmm_NB_sel_coll = dmm_NB_spp_melt[!(Var2 %in% ncoll_spp)
                                  & Var1 %in% filtered_eids
                                  , .(Var2 = collapse_str
                                  , value =  sum(value)), by = Var1][,.(Var1, Var2, value)] %>%
    rbind(dmm_NB_spp_melt[(Var2 %in% ncoll_spp) & Var1 %in% filtered_eids,],.)

dmm_NB_sel_plot = setkey(copy(cs_dmm_data), enrollment_id)[dmm_NB_sel_coll[,Var1] %>% as.character(),] %>%
    cbind(dmm_NB_sel_coll,.)

## set order based on relabund
breakorder = dmm_NB_sel_plot[order(-value), Var2] %>% unique() %>% as.character() %>% setdiff(collapse_str) %>% c(collapse_str)

## Phylum associations
taxon_phyl_assoc = taxonomy[g %in% breakorder,.(p, g)] %>% unique() %>% split(by = 'p')

dmm_NB_sel_plot[NB %in% c('cs.DMM3', 'cs.DMM6'), csGCT := gsub('cs.DMM','M1 cs-GCT', NB)]
taxon_bars_nb_sel = ggplot() + 
    geom_bar(aes(fill=factor(Var2, levels = breakorder), y=value, x=Var1)
             , color = NA, position='stack', stat='identity'
             , data = dmm_NB_sel_plot) +
    facet_nested(.~csGCT + country, scales = 'free', space = 'free', nest_line = element_line(linetype = 1)) + 
    theme_bw() + 
    theme(plot.title = element_text(size=11)) +
    theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
    theme(legend.key.size = unit(0.25, 'cm')
        , legend.key.height = unit(0.25, 'cm')
        , legend.key.width = unit(0.25, 'cm')
        , axis.ticks = element_blank()
        #, legend.title = element_text(size=14)
        #, legend.text = element_text(size=10)
        ) +
    scale_y_continuous(labels = scales::percent, expand = c(0,0)) +
    scale_x_discrete(expand = c(0,0)) +
    xlab('Individual') +
    ylab('Genus-level composition (%)') +
    scale_fill_manual(values = colpal, breaks = breakorder, name = 'Genus', guide = guide_legend(ncol = 1))

## Change facet label colors
pdf(file = NULL)
g = ggplot_gtable(ggplot_build(taxon_bars_nb_sel))
dev.off()

stripr <- which(grepl('strip-t', g$layout$name))
fills <- c(rep(col_pal_visit_light['NB'],2), rep(col_pal_country_light, 2))

k = 1
for (i in stripr) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k = k+1
}

taxon_bars_nb_sel = as.ggplot(g)

plot(taxon_bars_nb_sel)
```


## DMM Differential abundances
```{r vDMM6, fig.width = 6.5, fig.height = 2.5}

dmms = cs_dmm_data[,NB] %>% unique() %>% na.omit() %>% setdiff('cs.DMM6')

dmmdiffs = lapply(dmms, function(dmm)
{
    res = ancombc2_dmm6$res %>%
        dplyr::select(taxon, c(paste0('lfc_NB', dmm)
                           , paste0('se_NB', dmm)
                           , paste0('p_NB', dmm)
                           , paste0('q_NB', dmm))) %>%
        as.data.table()
    colnames(res) = c('taxon', 'l2fc', 'se', 'p', 'q')
    res$group = dmm
    res$signif =  ifelse(res$q < 0.1, '.', '') %>% ifelse(res$q < 0.05, '*', .) 
    return(res)
}) %>% do.call('rbind',.)
dmmdiffs$ref = 'cs.DMM6'

anydifftax = dmmdiffs[signif != '', taxon] %>% unique()
dmmdiffs[,csGCT := gsub('cs.DMM','M1 cs-GCT', group)]
dmmdiffs[signif == '', signif := 'n.s.']
vs_dmm6_coefplot = ggplot(data = dmmdiffs[taxon %in% anydifftax,]) +

    geom_vline(xintercept = 0, linetype = 'dashed', color = neutral_gray) +
    geom_bar(aes(x = l2fc, y = taxon, fill = signif), stat = 'identity') +
    geom_errorbar(aes(y = taxon, xmax = l2fc + 2*se, xmin = l2fc - 2*se), width = 0.15) +
    #scale_color_signif + 
    scale_fill_signif + 
    ylab('') +
    xlab('ANCOM-BC2 coefficient') +
    theme_bw() + 
    ggtitle('Differential abundance\nvs. cs-GCT6') +
    theme(plot.title = element_text(size=11)) +
    theme(panel.grid = element_blank()
        , legend.position = 'bottom'
        , legend.title.position = 'top'
        , legend.key.size = unit(0.25, 'cm')
        , legend.key.height = unit(0.25, 'cm')
        , legend.key.width = unit(0.25, 'cm')) +
    facet_grid(~ csGCT)

svglite('figures/dmm6-diffabunds.svg', width = 6.5, height = 2.5, fix_text_size = FALSE)
    plot(vs_dmm6_coefplot)
dev.off()

fwrite(dmmdiffs, 'source-data/fig4/diffabund-vs-dmm6.txt.xls', sep='\t')

```


```{r, fig.width=4, fig.height = 2.5}
dmm6_vs_dmms_coefplot = ggplot(data = dmmdiffs[group %in% c('cs.DMM3', 'cs.DMM2') & taxon %in% anydifftax,]) +
    geom_vline(xintercept = 0, linetype = 'dashed', color = neutral_gray) +
    #geom_point(aes(x = l2fc, y = taxon, color = signif), size = 2) +
    geom_bar(aes(x = l2fc, y = taxon, fill = signif), stat = 'identity') +
    geom_errorbar(aes(y = taxon, xmax = l2fc + 2*se, xmin = l2fc - 2*se), width = 0.15) +
    #scale_color_signif + 
    scale_fill_signif + 
    ylab('') +
    xlab('ANCOM-BC2 l2fc') +
    theme_bw() + 
    ggtitle('ANCOM-BC2 vs. M1 cs-GCT6') +
    theme(plot.title = element_text(size=11)) +
    theme(panel.grid = element_blank()
        , legend.position = 'bottom'
        , legend.title.position = 'top'
        ,legend.key.size = unit(0.25, 'cm')
        , legend.key.height = unit(0.25, 'cm')
        , legend.key.width = unit(0.25, 'cm')) +
    facet_grid(~csGCT)
plot(dmm6_vs_dmms_coefplot)

svglite('figures/dmm6-diffabunds-limited.svg', width = 4, height = 2.5, fix_text_size = FALSE)
    plot(dmm6_vs_dmms_coefplot)
dev.off()

```


```{r vDMM3, fig.width = 6.5, fig.height = 3.5}
dmms = cs_dmm_data[,NB] %>% unique() %>% na.omit() %>% setdiff('cs.DMM3')

dmmdiffs = lapply(dmms, function(dmm)
{
    res = ancombc2_dmm3$res %>%
        dplyr::select(taxon, c(paste0('lfc_NB', dmm)
                           , paste0('se_NB', dmm)
                           , paste0('p_NB', dmm)
                           , paste0('q_NB', dmm))) %>%
        as.data.table()
    colnames(res) = c('taxon', 'l2fc', 'se', 'p', 'q')
    res$group = dmm
    res$signif =  ifelse(res$q < 0.1, '.', '') %>% ifelse(res$q < 0.05, '*', .) 
    return(res)
}) %>% do.call('rbind',.)
dmmdiffs$ref = 'cs.DMM3'

anydifftax = dmmdiffs[signif != '', taxon] %>% unique()
dmmdiffs[,csGCT := gsub('cs.DMM','M1 cs-GCT', group)]
dmmdiffs[signif == '', signif := 'n.s.']

vs_dmm3_coefplot = ggplot(data = dmmdiffs[taxon %in% anydifftax,]) +

    geom_vline(xintercept = 0, linetype = 'dashed', color = neutral_gray) +
    geom_bar(aes(x = l2fc, y = taxon, fill = signif), stat = 'identity') +
    geom_errorbar(aes(y = taxon, xmax = l2fc + 2*se, xmin = l2fc - 2*se), width = 0.15) +
    #scale_color_signif + 
    scale_fill_signif + 
    ylab('') +
    xlab('ANCOM-BC2 l2fc') +
    theme_bw() + 
    ggtitle('Differential abundance\nvs. cs-GCT3') +
    theme(plot.title = element_text(size=11)) +
    theme(panel.grid = element_blank()
        , legend.position = 'bottom'
        , legend.title.position = 'top'
        , legend.key.size = unit(0.25, 'cm')
        , legend.key.height = unit(0.25, 'cm')
        , legend.key.width = unit(0.25, 'cm')) +
    facet_grid(~ csGCT)

plot(vs_dmm3_coefplot)

svglite('figures/dmm3-diffabunds.svg', width = 6.5, height = 3.5, fix_text_size = FALSE)
    plot(vs_dmm3_coefplot)
dev.off()

fwrite(dmmdiffs, 'source-data/ed-fig5/diffabunds.txt.xls', sep = '\t')

```

### Extended data multinomial plots.
```{r, fig.width = 2.5, fig.height = 2}
## Pre-vaccine IgA vs DMM1 at 2mo
dats = as.data.table(pre_pdmm1_2mo, keep.rownames=TRUE)
dats[,Significance := gsub('\\*\\*+','\\*', signif)] ## single-star significance level

pre_odds_vdmm1 = ggplot(data = dats) +
    ## render points
    geom_point(aes(x = gsub('cs.DMM','cs-GCT', rn), y = coef, color = Significance), size = 2) +
    
    ## add in error bars
    geom_errorbar(aes(x = gsub('cs.DMM','cs-GCT', rn), ymax = coef + 2*se, ymin = coef - 2*se), width = 0.15) +
    
    ## color scale for significance
    scale_color_signif + 
    ylab('log(OR)') +
    xlab('') +
    theme_bw() +
    ggtitle('Multinomial pre-RV1 Iga (M2)') +
    theme(plot.title = element_text(size=11)) +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    theme(panel.grid = element_blank()
        , legend.position = 'none'
        , axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
    facet_grid(~'vs M2 cs-GCT1')


plot(pre_odds_vdmm1)

svglite('figures/odds-multinom-pre-iga.svg', width = 2.5, height = 2, fix_text_size = FALSE)
    plot(pre_odds_vdmm1)
dev.off()

fwrite(dats, 'source-data/ed-fig4/pre-vaccine-iga-multinom.txt.xls', sep='\t')
```

```{r, fig.width = 2.5, fig.height = 5}
## Pre-vaccine IgA vs DMM6 at NB
dats = as.data.table(post_pdmm6_NB, keep.rownames=TRUE)
dats[,Significance := gsub('\\*\\*+','\\*', signif)] ## single-star significance level

post_odds_vdmm6 = ggplot(data = dats) +
    ## render points
    geom_point(aes(x = gsub('cs.DMM','cs-GCT', rn), y = coef, color = Significance), size = 2) +
    
    ## add in error bars
    geom_errorbar(aes(x = gsub('cs.DMM','cs-GCT', rn), ymax = coef + 2*se, ymin = coef - 2*se), width = 0.15) +
    
    ## color scale for significance
    scale_color_signif + 
    ylab('log(OR)') +
    xlab('') +
    ggtitle('Multinomial post-RV1 IgA') +
    theme_bw() + 
    theme(plot.title = element_text(size=11)) +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    theme(panel.grid = element_blank()
        , legend.position = 'none'
        #, legend.title.position = 'top'
        , axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
    facet_grid(~'vs M1 cs-GCT6')

plot(post_odds_vdmm6)

## Pre-vaccine IgA vs DMM6 at NB
dats = as.data.table(pre_pdmm6_NB, keep.rownames=TRUE)
dats[,Significance := gsub('\\*\\*+','\\*', signif)] ## single-star significance level

pre_odds_vdmm6 = ggplot(data = dats) +
    ## render points
    geom_point(aes(x = gsub('cs.DMM','cs-GCT', rn), y = coef, color = Significance), size = 2) +
    
    ## add in error bars
    geom_errorbar(aes(x = gsub('cs.DMM','cs-GCT', rn), ymax = coef + 2*se, ymin = coef - 2*se), width = 0.15) +
    
    ## color scale for significance
    scale_color_signif + 
    ylab('log(OR)') +
    xlab('') +
    ggtitle('Multinomial pre-RV1 IgA') +
    theme_bw() + 
    theme(plot.title = element_text(size=11)) +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    theme(panel.grid = element_blank()
        , legend.position = 'none'
        #, legend.title.position = 'top'
        , axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
    facet_grid(~'vs. M1 cs-GCT6')

pdf(file = NULL)
    pl = ggpubr::get_legend(pre_odds_vdmm6 + theme(legend.position = 'bottom', legend.title.position = 'top'))
dev.off()

odds_multinom_iga = plot_grid(pre_odds_vdmm6, post_odds_vdmm6, nrow = 2, rel_widths = c(1, 1), align = "v", axis = "l") %>% plot_grid(., pl, rel_heights = c(1, 0.2), nrow = 2)

plot(odds_multinom_iga)

svglite('figures/odds-multinom-iga.svg', width = 2.5, height = 5, fix_text_size = FALSE)
    plot(odds_multinom_iga)
dev.off()
```

## DMM-Adonis
```{r}
library('khroma')
## Define palette for each timepoint

nb_dmms = cs_dmm_data[,unique(NB)] %>% na.omit() %>% sort()
nb_dmm_pal = (color("bright")(7))[1:length(nb_dmms)] %>% setNames(nb_dmms)
nb_csgt_pal = nb_dmm_pal %>% setNames(gsub('cs.DMM','M1 cs-GCT', nb_dmms))

## 50% transparent mix for light version
nb_csgt_pal_light = nb_csgt_pal %>%
    lapply(function(co) hex(mixcolor(alpha = 0.5, color1 = hex2RGB(co), color2 = hex2RGB('#FFFFFF')))) %>%
    do.call('c',.)

## 25% transparent mix for light version
nb_csgt_pal_pale = nb_csgt_pal_light %>%
    lapply(function(co) hex(mixcolor(alpha = 0.5, color1 = hex2RGB(co), color2 = hex2RGB('#FFFFFF')))) %>%
    do.call('c',.)

## Plot ordinations vs NB dmms function
plot_nb_ordination_dmm = function(ord_data, permanova)
{
    ord_data[,csGCT := gsub('cs.DMM','M1 cs-GCT', NB)]

    permanova = permanova %>% as.data.table(keep.rownames = TRUE)
    
    ord_data[,nbdmm.centroid.Axis.1 := median(Axis.1), by = .(csGCT)]
    ord_data[,nbdmm.centroid.Axis.2 := median(Axis.2), by = .(csGCT)]
    
    ## Background ellipses
    p1 = ggplot(data = ord_data) +
        stat_ellipse(mapping = aes(fill = csGCT, x = Axis.1, y = Axis.2), geom = "polygon", alpha = 0.15)
        
    ## Add spiders
    #for(dm in ord_data[,csGCT] %>% unique()) p1 = p1 +
    #    geom_segment(data = ord_data[csGCT == dm,]
    #               , mapping = aes(
    #                   x = Axis.1, xend = nbdmm.centroid.Axis.1
    #                 , y = Axis.2, yend = nbdmm.centroid.Axis.2
    #                 , color = csGCT)
    #               , alpha = 0.25
    #               , size = 0.65)
    
    ## Add in points
    p1 = p1 + geom_point(mapping = aes(
                   x = Axis.1
                 , y = Axis.2
                 , color = csGCT)
                 , size = 0.8)
    
    ## Centroids
    for(dm in ord_data[,csGCT] %>% unique()) p1 = p1 +
        geom_point(data = ord_data[csGCT == dm,][!duplicated(csGCT),]
                 , mapping = aes(
                      x = nbdmm.centroid.Axis.1
                    , y = nbdmm.centroid.Axis.2
                    , fill = csGCT)
                 , color = neutral_gray
                 , size = 2.5
                 , shape = 21)
    
    ## Build pvals.
    p.value = permanova[rn == 'NB', `Pr(>F)`]
    f.stat = permanova[rn == 'NB', `F`] %>% round(3)
    pval = ifelse(p.value < 0.01, '<0.01', paste0('= ', as.character(signif(p.value, digits=3))))
    plabs =  paste0('P ', pval, '\nF = ', f.stat)
    
    ## annotate
    p1 = p1 + annotate(geom = 'text'
                 , x = ord_data[,(min(Axis.1) %>% abs())] * ord_data[,min(Axis.1) %>% sign()]*1.2
                 , y = ord_data[,(max(Axis.2) %>% abs())] * ord_data[,max(Axis.2) %>% sign()]*1.1
                 , label = plabs
                 , hjust = 0, vjust = 1
                 , lineheight = 0.95
                 , color = neutral_gray)

    ## Theme elements
    p1 = p1 +
        scale_color_manual(values = nb_csgt_pal, breaks = names(nb_csgt_pal)) +
        scale_fill_manual(values = nb_csgt_pal, breaks = names(nb_csgt_pal)) +
        facet_grid(~relabely[visit]) +
        
        xlab('Axis1') +
        ylab('Axis2') +
        theme_bw() + 
    theme(plot.title = element_text(size=11)) +
        theme(legend.position = 'none', panel.grid = element_blank())
    
    ## Change facet label colors
    pdf(file = NULL)
    g = ggplot_gtable(ggplot_build(p1))
    dev.off()
    stripr <- which(grepl('strip-t', g$layout$name))
    fills <- col_pal_visit_light[intersect(names(col_pal_visit_light), ord_data[,visit])]
    
    k = 1
    for (i in stripr) {
      j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
      g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
      k = k+1
    }
    
    pl1 = ggplotify::as.ggplot(g)

    return(pl1)
}

## Plot ordinations vs NB dmms legend
plot_nb_ordination_dmm_leg = function(ord_data)
{
    ord_data[,csGCT := gsub('cs.DMM','M1 cs-GCT', NB)]
    ord_data[,nbdmm.centroid.Axis.1 := median(Axis.1), by = .(csGCT)]
    ord_data[,nbdmm.centroid.Axis.2 := median(Axis.2), by = .(csGCT)]
    
    ## Centroids
    for(dm in ord_data[,csGCT] %>% unique()) p1 = p1 +
        geom_point(data = ord_data[csGCT == dm,][!duplicated(csGCT),]
                 , mapping = aes(
                      x = nbdmm.centroid.Axis.1
                    , y = nbdmm.centroid.Axis.2
                    , fill = csGCT)
                 , color = neutral_gray
                 , size = 2.5
                 , shape = 21)
    ## Theme elements
    p1 = p1 +
        scale_fill_manual(values = nb_csgt_pal, breaks = names(nb_csgt_pal)) +
        xlab('Axis1') +
        ylab('Axis2') +
        theme_bw() + 
    theme(plot.title = element_text(size=11)) +
        theme(legend.position = 'bottom', legend.title.position = 'top', panel.grid = element_blank())
    return(p1)
}

```

```{r, fig.width = 5, fig.height = 3.5}
## Iterate adonis over visit number
aitch_dmm_data_NB = visits %>%
    lapply(
function(v)
{
    thisvisitsams = metadata[visit == v 
                             & enrollment_id %in% alphadivdmm[!is.na(NB), enrollment_id]
                            , sample_name] %>%
        intersect(colnames(rarecounts))
    
    thisvisitcounts = spp_clr[thisvisitsams,]
    
    thisvisitmeta = setkey(alphadivdmm, enrollment_id)[metadata[thisvisitsams, enrollment_id],] %>%
        setkey(sample_name)
    
    thisvisitmeta$visit = v

    v_euclidean = vegdist(thisvisitcounts, method = 'euclidean')
    
    ## Draw PCoA
    v_euclidean_pcoa_ord = pcoa(v_euclidean)
    v_euclidean_pcoa_coords = v_euclidean_pcoa_ord$vectors
    v_euclidean_pcoa_data = thisvisitmeta %>% cbind(v_euclidean_pcoa_coords)
    return(v_euclidean_pcoa_data)
}) %>% setNames(visits)

## NB
permanova = aitch_adonis_list_NB[[1]][[1]]
p_nb_nb = plot_nb_ordination_dmm(aitch_dmm_data_NB[[1]], permanova)

## 2mo
permanova = aitch_adonis_list_NB[[2]][[1]]
p_nb_2mo = plot_nb_ordination_dmm(aitch_dmm_data_NB[[2]], permanova)

## 4mo
permanova = aitch_adonis_list_NB[[3]][[1]]
p_nb_4mo = plot_nb_ordination_dmm(aitch_dmm_data_NB[[3]], permanova)

pdf(file = NULL)
    legend = ggpubr::get_legend(plot_nb_ordination_dmm_leg(aitch_dmm_data_NB[[1]]))
dev.off()

fig4e = plot_grid(p_nb_2mo, p_nb_4mo, labels = c('A',''), ncol = 2) %>% plot_grid(., legend, rel_heights = c(1, 0.4), nrow = 2)

plot(fig4e)
```


## Render figure 4
```{r, fig.width = 5, fig.height = 8}

## Sero/shedding odds from multinom plot as figure 4 A and B
fig4ab = odds_multinom
svglite('figures/figure4-AB.svg', width = 5, height = 2.5, fix_text_size = FALSE)
    plot(fig4ab)
dev.off()

## DMM differences as panels C and D
fig4cd = plot_grid(taxon_bars_nb_sel, '', rel_heights = c(2.8, 0.7), ncol = 1) %>%
plot_grid(., dmm6_vs_dmms_coefplot, rel_widths = c(1,0.45), ncol = 2)

svglite('figures/figure4-CD.svg', width = 9.8, height = 3.5, fix_text_size = FALSE)
    plot(fig4cd)
dev.off()

## Panel E (print this to a file, but the final version doesn't use it.)
svglite('figures/figure4-E.svg', width = 5, height = 3.5, fix_text_size = FALSE)
    plot(fig4e)
dev.off()

plot_grid(fig4cd, fig4ab, fig4e, nrow = 3)
```

```{r, fig.width = 6.5, fig.height = 5}
pg = plot_grid(taxon_bars_2mo_all,
taxon_bars_4mo_all, ncol = 1, labels = c('A','B'))

pg2 = plot_grid(taxon_bars_nb_all, vs_dmm3_coefplot, vs_dmm6_coefplot, labels = c('A','B','C'))

svglite('figures/extended-data-taxon-barplots-other-timepoints.svg', width = 6.5, height = 5, fix_text_size = FALSE)
    plot(pg)
dev.off()
```

# Figure 5: Calprotectin
```{r}
library('ggplotify')
library('ggsignif')
load('calprotectin.Rdata')
```

## PCR M1
```{r, fig.width = 2, fig.height = 2.5}
## Calprotectin boxplots vs shedding
p1 = ggplot(data = cal_md[!is.na(rota_stool_pcr),]) +
    scale_color_pcr_outer +
    scale_fill_pcr_pale +
    geom_boxplot(outlier.shape = NA
         , aes(x = rota_stool_pcr
         , y = cal_2wk
         , color = rota_stool_pcr
         , fill = rota_stool_pcr)) +
    geom_beeswarm(
          aes(x = rota_stool_pcr
         , y = cal_2wk
         , color = rota_stool_pcr
         , fill = rota_stool_pcr), size = 0.75) +
    scale_x_discrete(breaks = cal_md[,rota_stool_pcr] %>% factor() %>% levels()
                   , labels = c('-','+')) + 
    
    ggpubr::geom_signif(
        data = data.frame(xmx = 'Negative', xmn = 'Positive', label = 'P = 0.043 *', yp = 12),
        mapping = aes(xmin = xmn, xmax = xmx, annotations = label, y_position = yp),
        textsize = 3, vjust = -0.2,
        manual = TRUE,
        inherit.aes = FALSE
    ) +
    
    ylim(c(NA,12.5)) +
    
    theme_bw() + 
    theme(plot.title = element_text(size=11)) +
    xlab('RV1 shedding, M4') +
    ylab('log2-Calprotectin, M1') +
    #theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
    #theme(legend.position = 'bottom', legend.title.position = 'top') +
    theme(legend.position = 'none') +
    theme(panel.spacing = unit(0, "lines")) +
    theme(panel.grid = element_blank())

calprotectin_pcr_m1_boxplots = p1
plot(calprotectin_pcr_m1_boxplots)
```

## PCR M2
```{r, fig.width = 2, fig.height = 2.5}
## Calprotectin boxplots vs shedding
p1 = ggplot(data = cal_md[!is.na(rota_stool_pcr),]) +
    scale_color_pcr_outer +
    scale_fill_pcr_pale +
    geom_boxplot(outlier.shape = NA
         , aes(x = rota_stool_pcr
         , y = cal_2mo
         , color = rota_stool_pcr
         , fill = rota_stool_pcr)) +
    geom_beeswarm(
          aes(x = rota_stool_pcr
         , y = cal_2mo
         , color = rota_stool_pcr
         , fill = rota_stool_pcr), size = 0.75) +
    scale_x_discrete(breaks = cal_md[,rota_stool_pcr] %>% factor() %>% levels()
                   , labels = c('-','+')) + 
    
    ggpubr::geom_signif(
        data = data.frame(xmx = 'Negative', xmn = 'Positive', label = 'P = 0.50', yp = 12),
        mapping = aes(xmin = xmn, xmax = xmx, annotations = label, y_position = yp),
        textsize = 3, vjust = -0.2,
        manual = TRUE,
        inherit.aes = FALSE
    ) +
    
    ylim(c(NA,12.5)) +
    
    theme_bw() + 
    theme(plot.title = element_text(size=11)) +
    xlab('RV1 shedding, M4') +
    ylab('log2-Calprotectin, M2') +
    #theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
    #theme(legend.position = 'bottom', legend.title.position = 'top') +
    theme(legend.position = 'none') +
    theme(panel.spacing = unit(0, "lines")) +
    theme(panel.grid = element_blank())

calprotectin_pcr_m2_boxplots = p1
plot(calprotectin_pcr_m2_boxplots)
```

## Sero M1
```{r, fig.width = 2, fig.height = 2.5}
## Calprotectin boxplots vs sero, m1
p1 = ggplot(data = cal_md[!is.na(Seroconverted),]) +
    scale_color_sero_outer +
    scale_fill_sero_pale +
    geom_boxplot(outlier.shape = NA
         , aes(x = Seroconverted
         , y = cal_2wk
         , color = Seroconverted
         , fill = Seroconverted)) +
    geom_beeswarm(
          aes(x = Seroconverted
         , y = cal_2wk
         , color = Seroconverted
         , fill = Seroconverted), size = 0.75) +
    scale_x_discrete(breaks = cal_md[,Seroconverted] %>% factor() %>% levels()
                   , labels = c('NR','R')) + 
    
    ggpubr::geom_signif(
        data = data.frame(xmx = 'Non-responder', xmn = 'Responder', label = 'P = 0.86', yp = 12),
        mapping = aes(xmin = xmn, xmax = xmx, annotations = label, y_position = yp),
        textsize = 3, vjust = -0.2,
        manual = TRUE,
        inherit.aes = FALSE
    ) +
    
    ylim(c(NA,12.5)) +
    
    theme_bw() + 
    theme(plot.title = element_text(size=11)) +
    xlab('RV1 Seroconversion') +
    ylab('log2-Calprotectin, M1') +
    #theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
    theme(axis.text.x = element_text()) +
    #theme(legend.position = 'bottom', legend.title.position = 'top') +
    theme(legend.position = 'none') +
    theme(panel.spacing = unit(0, "lines")) +
    theme(panel.grid = element_blank())

calprotectin_sero_m1_boxplots = p1
plot(calprotectin_sero_m1_boxplots)
```

## Sero M2
```{r, fig.width = 2, fig.height = 2.5}
## Calprotectin boxplots vs shedding
p1 = ggplot(data = cal_md[!is.na(Seroconverted),]) +
    scale_color_sero_outer +
    scale_fill_sero_pale +
    geom_boxplot(outlier.shape = NA
         , aes(x = Seroconverted
         , y = cal_2mo
         , color = Seroconverted
         , fill = Seroconverted)) +
    geom_beeswarm(
          aes(x = Seroconverted
         , y = cal_2mo
         , color = Seroconverted
         , fill = Seroconverted), size = 0.75) +
    scale_x_discrete(breaks = cal_md[,Seroconverted] %>% factor() %>% levels()
                   , labels = c('NR','R')) + 
    
    ggpubr::geom_signif(
        data = data.frame(xmx = 'Non-responder', xmn = 'Responder', label = 'P = 0.014 *', yp = 12),
        mapping = aes(xmin = xmn, xmax = xmx, annotations = label, y_position = yp),
        textsize = 3, vjust = -0.2,
        manual = TRUE,
        inherit.aes = FALSE
    ) +
    
    ylim(c(NA,12.5)) +
    
    theme_bw() + 
    theme(plot.title = element_text(size=11)) +
    xlab('RV1 Seroconversion') +
    ylab('log2-Calprotectin, M2') +
    #theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
    theme(axis.text.x = element_text()) +
    #theme(legend.position = 'bottom', legend.title.position = 'top') +
    theme(legend.position = 'none') +
    theme(panel.spacing = unit(0, "lines")) +
    theme(panel.grid = element_blank())

calprotectin_sero_m2_boxplots = p1
plot(calprotectin_sero_m2_boxplots)
```

```{r, fig.width = 4, fig.height = 5}
calprotectin_pg = plot_grid(calprotectin_pcr_m1_boxplots
                          , calprotectin_pcr_m2_boxplots
                          , calprotectin_sero_m1_boxplots
                          , calprotectin_sero_m2_boxplots
                          , labels = c('A','B','C','D'))

plot(calprotectin_pg)

svglite('figures/calprotectin-extended-data.svg', width = 4, height = 5)
    plot(calprotectin_pg)
dev.off()

cal_md[,.(Seroconverted, rota_stool_pcr, cal_m1 = cal_2wk, cal_m2 = cal_2mo, csGCT_m1 = NB)] %>% 
    fwrite('source-data/fig5/calprotectin-data.txt.xls', sep = '\t')
```

```{r, fig.width = 2.5, fig.height = 2.5}
feats = fread('.tmp.Maaslin2_cal2wk_NB/features/filtered_data_norm.tsv', sep = '\t', header=TRUE)
bifidos_nb = feats[,Bifidobacterium.longum] %>% setNames(feats[,feature])
names(bifidos_nb) = filtered_metadata[names(bifidos_nb), enrollment_id]

cal_bif_nb = cal_md[names(bifidos_nb),] %>% cbind(bifidos_nb)

bifidos_nb_plot = ggplot(data = cal_bif_nb[!is.na(country),], aes(x = cal_2wk, y = asin(sqrt(bifidos_nb)))) + geom_point(color = '#4477AA', size = 0.75) + geom_smooth(method = 'lm') +
    theme_bw() + 
    theme(plot.title = element_text(size=11)) +
    ylab('asin(sqrt(B. longum)), M1') +
    xlab('log2-Calprotectin, M1') +
    theme(panel.grid = element_blank())

plot(bifidos_nb_plot)

cal_bif_nb[!is.na(country), .(m1_cal = cal_2wk, ast_bifidos  = asin(sqrt(bifidos_nb)))] %>% 
    fwrite('source-data/fig5/bifidos-data.txt.xls', sep = '\t')

```

```{r, fig.width = 4, fig.height = 2.5}
library('DescTools')
cal_md[,csGCT := gsub('cs.DMM','cs-GCT', NB)]

z = DunnTest(cal_md[!is.na(NB),cal_2wk], cal_md[!is.na(NB),NB], method = 'BH')

ps = z$pmat %>% reshape2::melt() %>% as.data.table()

sigs = ps[value < 0.1,]

sigs[,pv:= round(value, 3)]
sigs[,label := paste0('P = ', pv, '*')]

p3 = ggplot(data = cal_md[!is.na(csGCT),]) +
        #scale_color_manual(values = nb_csgt_pal, breaks = names(nb_csgt_pal)) +
        #scale_fill_manual(values = nb_csgt_pal_pale, breaks = names(nb_csgt_pal)) +
        #facet_grid(~relabely['NB']) +
   
    geom_boxplot(aes(x = csGCT, y = cal_2wk)
               , outlier.shape = NA, color = neutral_gray
               , fill = neutral_gray_pale) +
    geom_beeswarm(aes(x = csGCT, y = cal_2wk), size = 0.75, color = neutral_gray) +
    
    ggpubr::geom_signif(
        data = data.frame(xmx = 'cs-GCT1', xmn = 'cs-GCT2', label = 'P = 0.038', yp = 12),
        mapping = aes(xmin = xmn, xmax = xmx, annotations = label, y_position = yp),
        textsize = 3, vjust = -0.2,
        manual = TRUE,
        inherit.aes = FALSE
    ) +
    
    theme_bw() + 
    
    ylim(c(NA, 12.5)) +
    
    theme(plot.title = element_text(size=11)) +
    xlab('M1 cross-sectional Gut Community Type (cs-GCT)') +
    ylab('log2-Calprotectin, M1') +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
    theme(legend.position = 'none') +
    theme(panel.grid = element_blank())

calprotectin_country_dmm_plots = p3

## print
plot(calprotectin_country_dmm_plots)
svglite('figures/calrprotectin-dmms.svg', width = 4, height = 2.5, fix_text_size = FALSE)
    plot(calprotectin_country_dmm_plots)
dev.off()


```

## Render figure 5
```{r}

## I ended up taking this plot apart to make figure 5.
gridplot = plot_grid(calprotectin_pcr_m1_boxplots, bifidos_nb_plot, ncol = 2, labels = c('A','B')) %>%
    list(.,calprotectin_country_dmm_plots) %>%
    plot_grid(plotlist = ., nrow = 2, rel_heights = c(0.8, 1), labels = c('','C'))

svglite('figures/figure-5.svg', width = 6, height = 6, fix_text_size = FALSE)
    plot(gridplot)
dev.off()


```

# ED-Figure 6:  Machine learning
```{r}
## Load in the ML dataset
load('rota-lgbm.Rdata')
```

```{r fig.width = 6, fig.height = 6.5}


get_xtabs = function(model, data, truth)
{
    truth = truth %>% unlist() %>% unname() %>% make.names() %>% factor()
    data = data[!is.na(truth),]
    predictions = predict(model, data, type = 'prob')
    colnames(predictions) = make.names(colnames(predictions))
    
    predictions$.pred_class = colnames(predictions)[predictions %>% apply(1, which.max)] %>%
        gsub('^\\.pred_','',.) %>%
        factor(levels = levels(truth))
    
    xtab = table(predictions$.pred_class, truth)
    return(xtab)
}

library('tidymodels')
dependentvar = 'Seroconverted'

## NB timepoint
test_curves = get_roc_pr_curves(model = nb_bfmodl
                  , data = nb_data_test
                  , truth = nb_data_test[,dependentvar]
                  , label = 'holdout')
nb_curves = test_curves

nb_roc_plot = ggplot() +
  geom_line(data = test_curves[[1]], aes(x = 1 - specificity, y = sensitivity), color = '#4477AA', size = 0.75) +
  geom_abline(
        lty = 2, alpha = 0.5,
        color = "gray50",
        size = 1
    ) +
    facet_grid(~'M1') + 
    theme_bw() + 
    theme(plot.title = element_text(size=11)) + 
    theme(panel.grid = element_blank()) +
    theme(strip.background = element_rect(fill = col_pal_visit_light['NB']))

noskill_pr_test = nb_data_test[,dependentvar] %>% unlist() %>% table() %>% (function(x) x[1] / (x[2] + x[1]))
nb_pr_plot = ggplot() +
  geom_line(data = test_curves[[2]], aes(x = recall, y = precision), color = '#44AA99', size = 1) +
  geom_hline(
        lty = 2, alpha = 0.5,
        color = "gray50",
        size = 1,
        yintercept = noskill_pr_test
    ) +
    facet_grid(~'M1') + 
    theme_bw() + 
    theme(plot.title = element_text(size=11)) + 
    theme(panel.grid = element_blank()) +
    theme(strip.background = element_rect(fill = col_pal_visit_light['NB']))

nb_test_xtabs = get_xtabs(model = nb_bfmodl
                  , data = nb_data_test
                  , truth = nb_data_test[,dependentvar])
rownames(nb_test_xtabs) = abbreviate.response(rownames(nb_test_xtabs))
colnames(nb_test_xtabs) = abbreviate.response(colnames(nb_test_xtabs))
nb_xtab_plot = ggtexttable(nb_test_xtabs)

## M2 timepoint
test_curves = get_roc_pr_curves(model = m2_bfmodl
                  , data = m2_data_test
                  , truth = m2_data_test[,dependentvar]
                  , label = 'holdout')
m2_curves = test_curves


m2_test_xtabs = get_xtabs(model = m2_bfmodl
                  , data = m2_data_test
                  , truth = m2_data_test[,dependentvar])
rownames(m2_test_xtabs) = abbreviate.response(rownames(m2_test_xtabs))
colnames(m2_test_xtabs) = abbreviate.response(colnames(m2_test_xtabs))
m2_xtab_plot = ggtexttable(m2_test_xtabs)

m2_roc_plot = ggplot() +
  geom_line(data = test_curves[[1]], aes(x = 1 - specificity, y = sensitivity), color = '#4477AA', size = 0.75) +
  geom_abline(
        lty = 2, alpha = 0.5,
        color = "gray50",
        size = 1
    ) +
    facet_grid(~'M2') + 
    theme_bw() + 
    theme(plot.title = element_text(size=11)) + 
    theme(panel.grid = element_blank()) +
    theme(strip.background = element_rect(fill = col_pal_visit_light['2mo']))

noskill_pr_test = m2_data_test[,dependentvar] %>% unlist() %>% table() %>% (function(x) x[1] / (x[2] + x[1]))
m2_pr_plot = ggplot() +
  geom_line(data = test_curves[[2]], aes(x = recall, y = precision), color = '#44AA99', size = 1) +
  geom_hline(
        lty = 2, alpha = 0.5,
        color = "gray50",
        size = 1,
        yintercept = noskill_pr_test
    ) +
    facet_grid(~'M2') + 
    theme_bw() + 
    theme(plot.title = element_text(size=11)) + 
    theme(panel.grid = element_blank()) +
    theme(strip.background = element_rect(fill = col_pal_visit_light['2mo']))


## M4 timepoint
test_curves = get_roc_pr_curves(model = m4_bfmodl
                  , data = m4_data_test
                  , truth = m4_data_test[,dependentvar]
                  , label = 'holdout')
m4_curves = test_curves


m4_test_xtabs = get_xtabs(model = m4_bfmodl
                  , data = m4_data_test
                  , truth = m4_data_test[,dependentvar])

rownames(m4_test_xtabs) = abbreviate.response(rownames(m4_test_xtabs))
colnames(m4_test_xtabs) = abbreviate.response(colnames(m4_test_xtabs))
m4_xtab_plot = ggtexttable(m4_test_xtabs)

m4_roc_plot = ggplot() +
  geom_line(data = test_curves[[1]], aes(x = 1 - specificity, y = sensitivity), color = '#4477AA', size = 0.75) +
  geom_abline(
        lty = 2, alpha = 0.5,
        color = "gray50",
        size = 1
    ) +
    facet_grid(~'M4') + 
    theme_bw() + 
    theme(plot.title = element_text(size=11)) + 
    theme(panel.grid = element_blank()) +
    theme(strip.background = element_rect(fill = col_pal_visit_light['4mo']))

noskill_pr_test = m4_data_test[,dependentvar] %>% unlist() %>% table() %>% (function(x) x[1] / (x[2] + x[1]))
m4_pr_plot = ggplot() +
  geom_line(data = test_curves[[2]], aes(x = recall, y = precision), color = '#44AA99', size = 1) +
  geom_hline(
        lty = 2, alpha = 0.5,
        color = "gray50",
        size = 1,
        yintercept = noskill_pr_test
    ) +
    facet_grid(~'M4') + 
    theme_bw() + 
    theme(plot.title = element_text(size=11)) + 
    theme(panel.grid = element_blank()) +
    theme(strip.background = element_rect(fill = col_pal_visit_light['4mo']))

nb_row = plot_grid(nb_roc_plot, nb_pr_plot, nb_xtab_plot, rel_widths = c(1, 1, 0.5), nrow = 1)
m2_row = plot_grid(m2_roc_plot, m2_pr_plot, m2_xtab_plot, rel_widths = c(1, 1, 0.5), nrow = 1)
m4_row = plot_grid(m4_roc_plot, m4_pr_plot, m4_xtab_plot, rel_widths = c(1, 1, 0.5), nrow = 1)

plot_grid(nb_row, m2_row, m4_row, labels = c('A','B','C'), ncol = 1)

saveRDS(nb_bfmodl, file = 'm1-lgbm.rds')
saveRDS(m2_bfmodl, file = 'm2-lgbm.rds')
saveRDS(m4_bfmodl, file = 'm4-lgbm.rds')

```

## Render figure 6
```{r}
p6 = plot_grid(nb_row, m2_row, m4_row, labels = c('A','B','C'), ncol = 1)
svglite('figures/figure-6.svg', width = 6, height = 6.5, fix_text_size = FALSE)
    plot(p6)
dev.off()

## Save curves to file

roc_curves = rbind(
    cbind(nb_curves[[1]], t = 'm1'),
    cbind(m2_curves[[1]], t = 'm2'),
    cbind(m4_curves[[1]], t = 'm4')
)

pr_curves = rbind(
    cbind(nb_curves[[2]], t = 'm1'),
    cbind(m2_curves[[2]], t = 'm2'),
    cbind(m4_curves[[2]], t = 'm4')
)

fwrite(roc_curves, "source-data/ed-fig6/roc-curves.tsv.xls", sep = '\t')
fwrite(pr_curves, "source-data/ed-fig6/pr-curves.tsv.xls", sep = '\t')

```

# Tables
```{r}
cal_md %>% setkey(enrollment_id)

## OGU matrix and taxonomy
cbind(taxonomy, countmatrix) %>% fwrite('tables/ogu-matrix.txt.xls', sep = '\t')

metadata[,include := FALSE]
metadata[!is.na(Seroconverted) & sample_name %in% colnames(goodcounts),include := TRUE]

## Sample metadata
metadata %>%
    cbind(.,cal_md[metadata[,enrollment_id],.(cal_2wk, cal_2mo)]) %>%
    fwrite('tables/sample_metadata.txt.xls', sep = '\t')

```

# Suppfigs
```{r, fig.width = 6, fig.height = 6}

svglite('figures/suppfig-alphadiv-regressions.svg', width = 6, height = 6, fix_text_size = FALSE)
    norm_plot(alphadivnona[,SR], transform = 'log', label = 'OGU Richness')
dev.off()

norm_plot(alphadivnona[,EMR], transform = 'log', label = 'EM Richness')  

norm_plot(alphadivnona[,PD], transform = 'log', label = "Faith's PD")  

norm_plot(alphadivnona[,simpson], transform = 'logit', label = "Simpson")  

norm_plot(alphadivnona[,shannon], transform = 'log', label = "Shannon")  

```

```{r}
sessionInfo()

```
