---
title: "ROTA Metagenomics: Differential abundance"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: true
    theme: lumen
---

# Summary
This document contains the scripts used to identify differentially abundant taxa on the Rota metagenome dataset.

# Setup
In this section are the chunks that set up the script for differential abundance analysis. Load in libraries, set parameters, and reformat the TSE object.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.width=10, fig.height=8)
```

```{r libraries}
## override library function to silence all package messages.
silence = function(x) suppressPackageStartupMessages(suppressWarnings({invisible({capture.output(x)})}))
library = function(...) silence({base::library(...)})

library('data.table')
library('tidyverse')
library('ggplot2')
library('ANCOMBC')
library('mia')
library('MicrobiomeStat')
library('Maaslin2') ## bioconductor
library('dplyr')
library('UpSetR')
library('cowplot')
library('phyloseq')
        
load(file="initialized-dataset.RData")
set.seed(42)

qval_threshold = 0.1 ## alter to make stricter/more lenient.

```

```{r DataManip}

## Function to assign significance at 0.1/0.05 level.
assign_stars = function(pvalues) ifelse(pvalues < 0.1, '.','n.s.') %>%
                 ifelse(pvalues < 0.05, '*',.)

## Taxa-are-rows
subsetByAbundantFeatures = function(x, abundance)
{
    rowProps = x %>% assay() %>% apply(2, function(y) y/sum(y))
    x[(rowSums(rowProps >= abundance) > 0) %>% which(),]
}

physeq_decon = merge_phyloseq(otudecon, taxdecon, sample_data(samdf))
tse_decon = makeTreeSEFromPhyloseq(physeq_decon)
tse_spp = agglomerateByRank(tse_decon, rank = 'Species')

## Library size
colData(tse_spp)$library_size = colSums(assay(tse_spp, "counts"))
colData(tse_spp)$log_pre_rota_sero_iga = log(tse_spp$pre_rota_sero_iga)
colData(tse_spp)$log_post_rota_sero_iga = log(tse_spp$post_rota_sero_iga)

countries = metadata[,country] %>% unique() %>% na.omit()
visits = metadata[,visit] %>% unique() %>% na.omit()
# slice_grid = expand.grid(countries, visits) %>% na.omit()

tse_sero = tse_spp[, !is.na(tse_decon$Seroconverted)] ## we have pre and post both

tse_sero_subset = tse_sero %>%
    subsetByAbundantFeatures(abundance = 0.01) %>%
    subsetByPrevalentFeatures(prevalence = 0.05)

```


# Longitudinal Differential abundance
Execute diffabund using MAASLin2
```{r Maaslin2}

## Define ref levels
reference = c('country,Panama', 'formula_bm,BOTH', 'visit,NB')
fixed_effects = c('visit', 'country', 'formula_bm', 'delivery', 'gest_age', 'library_size', 'Seroconverted')
random_effects = 'enrollment_id'

maaslin2_out = Maaslin2(input_data = tse_sero_subset %>% assay() %>% t() %>% as.data.frame(),
             input_metadata = tse_sero_subset %>% colData() %>% as.data.frame(),
             output = "results/Maaslin2",
             transform = "AST",
             fixed_effects = fixed_effects,
             random_effects = random_effects,
             reference = reference,
             normalization = "TSS",
             plot_heatmap = FALSE,
             plot_scatter = FALSE,
             save_scatter = FALSE,
             save_models = FALSE,
             standardize = FALSE)

maaslin2_res = maaslin2_out$results %>%
    filter(qval < qval_threshold) %>% print()

```

Execute diffabund using LinDA
```{r LinDA}

assign('linda_out',
    linda(as.data.frame(assay(tse_sero_subset, "counts")),
        as.data.frame(colData(tse_sero_subset)),
        formula = "~ visit + country + formula_bm + delivery + gest_age + Seroconverted + library_size + (1 | enrollment_id)",
        alpha = qval_threshold
        )
) %>% invisible() %>% capture.output()

linda_res = linda_out$output$SeroconvertedResponder

linda_res %>%
  filter(reject) %>%
  dplyr::select(log2FoldChange, stat, padj) %>%
  rownames_to_column(var = "feature")  %>% print()

```

Execute diffabund using ANCOM-BC2
```{r ANCOMBC2}
## Function to execute ancombc2 repeatedly.
execute_ancombc2 = function(x, tax_levels, fix_formula, rand_formula, group)
{
    tt = 'Species'
    neg_lb = FALSE
    ancombc2(data = x
        , assay.type = "counts"
        , fix_formula = fix_formula
        , rand_formula = rand_formula
        , p_adj_method = "fdr"
        , group = group
        , struc_zero = TRUE
        , neg_lb = TRUE
        , alpha = qval_threshold
        , global = TRUE
    )
}

## Exec ancombc2
ancombc2_out = execute_ancombc2(
      x = tse_sero_subset
    , tax_levels = c('Species')
    , fix_formula = "visit + country + formula_bm + delivery + gest_age + Seroconverted + library_size"
    , rand_formula = "(1 | enrollment_id)"
    , group = 'Seroconverted'
)

## Print results
ancombc2_out$res %>%
  dplyr::select(taxon, lfc_SeroconvertedResponder, q_SeroconvertedResponder) %>%
  filter(q_SeroconvertedResponder < qval_threshold) %>%
  arrange(q_SeroconvertedResponder) %>% print()

```

```{r Coallate}

results = list()

results[['global']] = list(
    as.data.table(ancombc2_out$res)[,
       .(Species = taxon %>% make.names()
        , effectSize = lfc_SeroconvertedResponder / log(2)
        , SE = se_SeroconvertedResponder / log(2)
        , padj = q_SeroconvertedResponder
        , pval = p_SeroconvertedResponder
        , method = 'ANCOM-BC2'
        , timepoint = 'global'
        )]
    ,
    
    as.data.table(linda_out$output$SeroconvertedResponder, keep.rownames = TRUE)[,
        .(Species = rn %>% make.names()
        , effectSize = log2FoldChange
        , SE = lfcSE
        , padj
        , pval = pvalue
        , method = 'LinDA'
        , timepoint = 'global'
        )]
    ,
    as.data.table(maaslin2_out$results)[metadata == 'Seroconverted',
        .(Species = feature %>% make.names()
        , effectSize = coef
        , SE = stderr
        , padj = qval
        , pval
        , method = 'Maaslin2'
        , timepoint = 'global'
        )]
) %>% do.call('rbind', .)
```

```{r PlotDiffabunds, fig.height = 6, fig.width = 8}

plottingdata = results[['global']]

## Render volcano-ish plot.
plottingdata[,signif := assign_stars(padj)]

plottingdata[,errmin := effectSize - 2*SE]
plottingdata[,errmax := effectSize + 2*SE]
signif_spp = plottingdata[padj < qval_threshold, Species] %>% unique()

errdata = plottingdata[Species %in% signif_spp,][,.(errcords = c(errmin,errmax)), by = .(Species, method)]

p1 =  ggplot(data = plottingdata) +
    geom_point(aes(x = effectSize, y = -log10(pval), color = signif)) +
    facet_grid(~method, scales = 'free') +
    scale_y_log10() +
    theme_bw()

p2 = ggplot(data = plottingdata[Species %in% signif_spp,]) +
    geom_bar(aes(x = effectSize, y = Species, fill = signif), stat = 'identity') +
    geom_line(data = errdata, aes(x=errcords, y = Species, group = Species)) +
    facet_grid(~method, scales = 'free') +
    theme_bw()


plottingdt = plottingdata[padj < qval_threshold,] %>% dcast(Species ~ method, fun.aggregate = length)
colnames(plottingdt) = colnames(plottingdt) %>% make.names()
plottingmat = plottingdt[,-1] %>% as.matrix()


intermediatelabels = c(2:(ncol(plottingmat)-1)) %>% lapply(function(m) combn(colnames(plottingmat), m, FUN = function(pp) paste0(pp, collapse='&'))) %>% unlist()
intermediates = c(2:(ncol(plottingmat)-1)) %>% lapply(function(m) combn(1:ncol(plottingmat), m) %>% apply(2, function(ii) rowSums(plottingmat[,c(ii)]) - rowSums(matrix(plottingmat[,-ii], ncol = ncol(plottingmat) - m)) == m) %>% colSums()) %>% unlist()

## Render upset plot
upset(
    fromExpression(
        c(
            plottingmat[rowSums(plottingmat) == 1,] %>% colSums(),
            setNames(intermediates, intermediatelabels),
            sum(rowSums(plottingmat) == ncol(plottingmat)) %>% setNames(paste0(colnames((plottingmat)), collapse='&'))
        )
    )
)

plot_grid(plotlist = list(p1, p2), nrow = 2)

```

# Cross-sectional differential abundance
Execute diffabund using MAASLin2
```{r Maaslin2_CS, fig.height = 4, fig.width = 3}

## Define ref levels
reference = c('country,Panama', 'formula_bm,BOTH')
fixed_effects = c('age', 'country', 'formula_bm', 'delivery', 'gest_age', 'library_size', 'Seroconverted')

for (visit in visits)
{
    tse_sero_subset_v = tse_sero[,colData(tse_sero)$visit == visit] %>%
        subsetByAbundantFeatures(abundance = 0.01) %>%
        subsetByPrevalentFeatures(prevalence = 0.05)
    
    assign(paste0('maaslin2_out_', visit)
           , Maaslin2(input_data = tse_sero_subset_v %>% assay() %>% t() %>% as.data.frame(),
                 input_metadata = tse_sero_subset_v %>% colData() %>% as.data.frame(),
                 output = "results/Maaslin2" %>% paste0(.,'_',visit),
                 transform = "AST",
                 fixed_effects = fixed_effects,
                 reference = reference,
                 normalization = "TSS",
                 standardize = FALSE,
                 min_prevalence = 0,
                 plot_heatmap = FALSE,
                 plot_scatter = FALSE,
                 save_scatter = FALSE,
                 save_models = FALSE)
    ) %>% invisible() %>% capture.output()
}

maaslin2_res_NB = maaslin2_out_NB$results %>%
    filter(qval < qval_threshold) %>%
    filter(metadata == 'Seroconverted') %>%
    print()

maaslin2_res_2mo = maaslin2_out_2mo$results %>%
    filter(qval < qval_threshold) %>%
    filter(metadata == 'Seroconverted') %>%
    print()

maaslin2_res_4mo = maaslin2_out_4mo$results %>%
    filter(qval < qval_threshold) %>%
    filter(metadata == 'Seroconverted') %>%
    print()

plottingdata = maaslin2_out_NB$results %>%  filter(metadata == 'Seroconverted') %>% as.data.table()
plottingdata[,signif := assign_stars(qval)]

p1 = ggplot(plottingdata) +
    geom_point(aes(x = coef, y = -log10(pval), color = signif)) +
    scale_y_log10() +
    theme_bw()


p2 = ggplot(data = plottingdata[signif != 'n.s.',]) +
    geom_bar(aes(x = coef, y = feature, fill = signif), stat = 'identity') +
    geom_segment(aes(x = coef - 2*stderr, xend = coef + 2*stderr,  y = feature, yend = feature, group = feature)) +
    theme_bw()

plot_grid(plotlist = list(p1, p2), nrow = 2)

```

# End
```{r Save}
save(list = ls(all.names = TRUE), file = 'diffabund.RData')
sessionInfo()
```
