---
title: "ROTA Metagenomics: Machine learning"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: true
    theme: lumen
---
# Summary
This document contains the scripts used to run & tune lightGBM models.

# Setup
In this section are the chunks that call libraries into the environment and intitalize the dataset.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.width=6, fig.height=5)
```

```{r libraries}
## override library function to silence all package messages.
silence = function(x) suppressPackageStartupMessages(suppressWarnings({invisible({capture.output(x)})}))
library = function(...) silence({base::library(...)})

library('data.table')
library('rlang')
library('lightgbm')
library('bonsai')
library('themis')
library('yardstick')
library('tidymodels')
library('tidyverse')
library('ggplot2')
library('kernelshap')
library('shapviz')
library('caret')
library('DescTools')
library('pROC')
library('cowplot')
library('dials')

load(file="calprotectin.RData")

evaluate.model = function(truth, prediction)
{
    cm = table(truth, prediction)
    
    # Evaluation Metrics
    n = sum(cm) # number of instances
    nc = nrow(cm) # number of classes
    diag = diag(cm) # number of correctly classified instances per class 
    rowsums = apply(cm, 1, sum) # number of instances per class
    colsums = apply(cm, 2, sum) # number of predictions per class
    p = rowsums / n # distribution of instances over the actual classes
    q = colsums / n # distribution of instances over the predicted classes
    accuracy = sum(diag) / n 
    message(paste0("Overall accuracy of the model is: ", accuracy) %>% message())  
    precision = diag / colsums 
    recall = diag / rowsums 
    f1 = 2 * precision * recall / (precision + recall) 
    metrics = data.table(accuracy = c(accuracy[1], NA %>% rep(nc-1))
                        , precision
                        , recall
                        , f1
                        )
    
    list(confusion.matrix = cm, metrics)
}

get_roc_pr_curves = function(model, data, truth, strata = NULL, label = NULL)
{
    truth = truth %>% unlist() %>% unname() %>% make.names() %>% factor()
    data = data[!is.na(truth),]
    predictions = predict(model, data, type = 'prob')
    colnames(predictions) = make.names(colnames(predictions))
    
    predictions$.pred_class = colnames(predictions)[predictions %>% apply(1, which.max)] %>%
        gsub('^\\.pred_','',.) %>%
        factor(levels = levels(truth))
    
    if(ncol(predictions) == 3)
    {
        predictions$.truth = truth
        predictions$.prob = 1 - (predictions[,colnames(predictions)[2]] %>% unlist())
        
        if(is.null(strata))
        {
            ROC = roc_curve(data = predictions, '.prob', truth = .truth)
            PR = pr_curve(data = predictions, '.prob', truth = .truth)
            
            ROC$dataset = ''
            PR$dataset = ''
            print(evaluate.model(predictions$.pred_class, predictions$.truth))
        } else {
            ROC = lapply(data[,strata] %>% unlist() %>% unique(),
            function(st)
            {
                stdata = data[data[,strata] == st,]
                stpred = predictions[data[,strata] == st,]
                
                classobs = stpred[,'.truth'] %>% unlist() %>% unique() %>% length()
                if(classobs == 1)
                {
                    warning('Only one class observed in truth.')
                    return(NULL)
                }
                
                stROC = roc_curve(data = stpred, '.prob', truth = .truth)
                stROC$dataset = st
                
                return(stROC)
            }) %>% do.call('rbind', .)
            
            PR = lapply(data[,strata] %>% unlist() %>% unique(),
            function(st)
            {
                stdata = data[data[,strata] == st,]
                stpred = predictions[data[,strata] == st,]
                
                classobs = stpred[,'.truth'] %>% unlist() %>% unique() %>% length()
                if(classobs == 1)
                {
                    warning('Only one class observed in truth.')
                    return(NULL)
                }
                
                stPR = pr_curve(data = stpred, '.prob', truth = .truth)
                stPR$dataset = st
                
                return(stPR)
            }) %>% do.call('rbind', .)
        }
    } else {
        ROC = NULL
        PR = NULL
    }
    ROC$dataset = paste0(ROC$dataset %>% unlist(), '.', label)
    PR$dataset = paste0(PR$dataset %>% unlist(), '.', label)
    return(list(ROC, PR))
}

get_xtabs = function(model, data, truth)
{
    truth = truth %>% unlist() %>% unname() %>% make.names() %>% factor()
    data = data[!is.na(truth),]
    predictions = predict(model, data, type = 'prob')
    colnames(predictions) = make.names(colnames(predictions))
    
    predictions$.pred_class = colnames(predictions)[predictions %>% apply(1, which.max)] %>%
        gsub('^\\.pred_','',.) %>%
        factor(levels = levels(truth))
    
    xtab = table(predictions$.pred_class, truth)
    return(xtab)
}

agglomerate = function(counts, taxonomy, rank)
{
    mcounts = reshape2::melt(counts) %>% as.data.table() %>% setkey(Var1)
    mtax = taxonomy[mcounts,][,.(value = sum(value)), by = .(Var2, R = get(rank))]
    
    dtax = dcast(mtax, R ~ Var2, value.var = 'value')
    aggmatrix = dtax[,-1] %>% as.matrix()
    rownames(aggmatrix) = dtax[,R]
    
    return(aggmatrix)
}


min_prevalence = 0.05
min_abundance = 0.01
pseudocount = 1

agglomerate = function(counts, taxonomy, rank)
## Agglomerate function for large sample-sizes.
## 1. Convert from wide to long
## 2. Bind to taxonomy
## 3. Sum counts within selected rank
## 4. Convert from long back to wide, add zeroes back into the matrix & return.
## Params:
## counts = a count matrix, where columns are samples, to agglomerate.
## taxonomy = a matrix or data frame where rows are OTUs, representing taxonomy
## rank = One of the columns of the taxonomy matrix (a string) to agglomerate on
## Returns:
## A count matrix where columns are samples and rows are the rank of interest.
{
    mcounts = reshape2::melt(counts) %>% as.data.table() %>% setkey(Var1)
    mtax = taxonomy[mcounts,][,.(value = sum(value)), by = .(Var2, R = get(rank))]
    
    dtax = data.table::dcast(mtax, R ~ Var2, value.var = 'value')
    aggmatrix = dtax[,-1] %>% as.matrix()
    rownames(aggmatrix) = dtax[,R]
    
    return(aggmatrix)
}

## Prevalence filter.
## Params:
## counts = a count matrix where samples are columns.
## min_prevalence = minimum prevalence threshold (as a decimal).
## Returns:
# A filtered count matrix with rows in fewer samples than the threshold of
# min_prevalence removed.
prevalence = function(counts, min_prevalence)
    counts[rowSums(counts > 0) > (ncol(counts)*min_prevalence),];

## Abundance filter.
## Params:
## counts = a count matrix where samples are columns.
## min_abundance = minimum abundance threshold (as a decimal).
## Returns:
## A filtered count matrix with samples at a lower abundance in any sample than
## the threshold removed.
abundance = function(counts, min_abundance)
    counts[(rowSums(apply(counts, 2, function(x) x/sum(x)) >= min_abundance) > 0) %>% which(),];

set.seed(42)
```

```{r Params}
## lightGBM params
lgb_ntrees = 500
lgb_k = 5
lgb_repeats = 1
lgb_gridsize = 100

## Generate train/test outer splits.
train_proportion = 0.7
min_abundance = 0.01
min_prevalence = 0.05
```

```{r FilterData}
## Bind alpha div data to filtered metadata.
filtered_metadata[alphadivs, c(colnames(alphadivs)) := mget(colnames(alphadivs))]
```

# Predicting delivery mode at NB
We're predicting delivery mode at NB because we should be able to in this cohort.
Using LightGBM.
```{r dvSplitData}
## Variable names for configs
dependentvar = 'delivery'
indivar = 'enrollment_id'
sampvar = 'sample_name'
timevar = 'visit'
timepoint = 'NB'
covars = c(colnames(alphadivs)) %>% setdiff('rn')
strata = c('country') ## stratify to ensure even coverage of countries in the training/testing split

ttmd = filtered_metadata[get(timevar) == timepoint
                , mget(c(sampvar, indivar, covars, dependentvar, strata))] %>% na.omit()

train_indivs = ttmd[,sample(get(indivar), ceiling(.N*train_proportion)), by = c(strata)][,V1]
test_indivs = setdiff(ttmd[,get(indivar)], train_indivs)
train_samples = ttmd[get(indivar) %in% train_indivs, get(sampvar)]
test_samples = ttmd[get(indivar) %in% test_indivs, get(sampvar)]
```

```{r dvTrainingSplit}
## Training covariate data
train_samples = ttmd[get(sampvar) %in% train_samples, get(sampvar)] ## Reorder according to the table.
train_covars = ttmd[get(sampvar) %in% train_samples, mget(covars)]
train_ids = ttmd[get(sampvar) %in% train_samples, mget(indivar)]
train_counts = goodcounts[,train_samples] %>% apply(2, function(x) x / sum(x))

## Response var
train_response = ttmd[get(sampvar) %in% train_samples, mget(dependentvar)]

## Agglomerate.
ogu_train_counts = train_counts %>%
    abundance(min_abundance) %>%
    prevalence(min_prevalence) %>%
    t() %>% 
    as_tibble() %>%
    rename_all(~ make.names(.)) %>%
    as.matrix()

spp_train_counts = train_counts %>%
    agglomerate(taxonomy, 's') %>%
    abundance(min_abundance) %>%
    prevalence(min_prevalence) %>%
    t() %>% 
    as_tibble() %>%
    rename_all(~ make.names(.)) %>%
    rename_all(~ paste0('s__',.)) %>%
    as.matrix()

gen_train_counts = train_counts %>%
    agglomerate(taxonomy, 'g') %>%
    abundance(min_abundance) %>%
    prevalence(min_prevalence) %>%
    t() %>% 
    as_tibble() %>%
    rename_all(~ make.names(.)) %>%
    rename_all(~ paste0('g__',.)) %>%
    as.matrix()

## Taxon-specific comps.
ogu_train_comp = ogu_train_counts[,-union(ogu_train_counts %>% nearZeroVar()
                                    , which(colSums(ogu_train_counts) == 0)
                                    )]
spp_train_comp = spp_train_counts[,-union(spp_train_counts %>% nearZeroVar()
                                    , which(colSums(spp_train_counts) == 0)
                                    )]
gen_train_comp = gen_train_counts[,-union(gen_train_counts %>% nearZeroVar()
                                    , which(colSums(gen_train_counts) == 0)
                                    )]

## Build training dataframe.
dv_data_train = cbind(train_response, train_covars, gen_train_comp, spp_train_comp) %>% as_tibble()
```

```{r dvTestingSplit}
## Training covariate data
test_samples = ttmd[get(sampvar) %in% test_samples, get(sampvar)] ## Reorder according to the table.
test_covars = ttmd[get(sampvar) %in% test_samples, mget(covars)]
test_ids = ttmd[get(sampvar) %in% test_samples, mget(indivar)]
test_counts = goodcounts[,test_samples] %>% apply(2, function(x) x / sum(x))

## Response var
test_response = ttmd[get(sampvar) %in% test_samples, mget(dependentvar)]

## Agglomerate.
ogu_test_counts = test_counts %>%
    t() %>% 
    as_tibble() %>%
    rename_all(~ make.names(.)) %>%
    dplyr::select(!!colnames(ogu_train_counts)) %>%
    as.matrix()

spp_test_counts = test_counts %>% agglomerate(taxonomy, 's') %>%
    t() %>% 
    as_tibble() %>%
    rename_all(~ make.names(.)) %>%
    rename_all(~ paste0('s__',.)) %>%
    dplyr::select(!!colnames(spp_train_counts)) %>%
    as.matrix()
 
gen_test_counts = test_counts %>% agglomerate(taxonomy, 'g') %>%
    t() %>% 
    as_tibble() %>%
    rename_all(~ make.names(.)) %>%
    rename_all(~ paste0('g__',.)) %>%
    dplyr::select(!!colnames(gen_train_counts)) %>%
    as.matrix()

## Taxon-specific comps.
ogu_test_comp = ogu_test_counts[,colnames(ogu_train_comp)]
spp_test_comp = spp_test_counts[,colnames(spp_train_comp)]
gen_test_comp = gen_test_counts[,colnames(gen_train_comp)]

## Build training dataframe.
dv_data_test = cbind(test_response, test_covars, gen_test_comp, spp_test_comp) %>% as_tibble()
selvars = c(covars, colnames(spp_train_comp), colnames(gen_train_comp))
```

```{r dvTuneParams}
lgb_classifier = boost_tree(
   trees = lgb_ntrees
 , tree_depth = tune()
 , min_n = tune()
 , loss_reduction = tune()
 , sample_size = tune()
 , mtry = tune()
 , learn_rate = tune()
 ) %>%
set_engine("lightgbm",
           is_unbalance = TRUE,
           #boosting = "dart",
           print_every_n=10) %>%
set_mode("classification")

modl = lgb_classifier

modl_formula = paste0(dependentvar, ' ~ .') %>% as.formula()
modl_recipe = recipe(modl_formula, data = dv_data_train)

modl_workflow = workflow() %>% 
    add_model(modl) %>%
    add_recipe(modl_recipe)

print(modl_workflow)
    
modl_cvfolds = vfold_cv(dv_data_train, strata = dependentvar, v = lgb_k, repeats = lgb_repeats);

## Make tuning grid.
lgb_tune_grid = grid_latin_hypercube(
   tree_depth()
 , min_n()
 , loss_reduction()
 , sample_size = sample_prop()
 , finalize(mtry(), dv_data_train)
 , learn_rate()
 #, alpha()
 , size = lgb_gridsize
 )
    
## Exec cross-validation of the model
message('Running cross-validation of model...')
delivery_lgb_res = tune_grid(
  modl_workflow
  , resamples = modl_cvfolds
  , grid = lgb_tune_grid
  , metric_set(yardstick::accuracy
             , yardstick::brier_class
             , yardstick::f_meas
             , yardstick::pr_auc
             , yardstick::kap
             , yardstick::roc_auc
             )
  , control = control_grid(save_pred = TRUE, verbose = TRUE, save_workflow = TRUE)
)

```

```{r dvFit}
delivery_bfmodl = fit_best(delivery_lgb_res, metric = 'roc_auc')
```

```{r dvEvalModel}

train_curves = get_roc_pr_curves(model = delivery_bfmodl
                  , data = dv_data_train
                  , truth = dv_data_train[,dependentvar]
                  , label = 'training')

## Eval holdout data
test_curves = get_roc_pr_curves(model = delivery_bfmodl
                  , data = dv_data_test
                  , truth = dv_data_test[,dependentvar]
                  , label = 'holdout')

## Plor ROC-curves.
train_roc_plot = ggplot() +
  geom_line(data = train_curves[[1]], aes(x = 1 - specificity, y = sensitivity), color = '#34A6FF', size = 1) +
  geom_abline(
        lty = 2, alpha = 0.5,
        color = "gray50",
        size = 1
    ) + theme_bw() + ggtitle('Training ROC Delivery NB')

test_roc_plot = ggplot() +
  geom_line(data = test_curves[[1]], aes(x = 1 - specificity, y = sensitivity), color = '#DD0534', size = 1) +
  geom_abline(
        lty = 2, alpha = 0.5,
        color = "gray50",
        size = 1
    ) + theme_bw() + ggtitle('Holdout ROC Delivery NB')

## Plot PR curves.
noskill_pr_test = dv_data_test[,dependentvar] %>% unlist() %>% table() %>% (function(x) x[1] / (x[2] + x[1]))
noskill_pr_train = dv_data_train[,dependentvar] %>% unlist() %>% table() %>% (function(x) x[1] / (x[2] + x[1]))

train_pr_plot = ggplot() +
  geom_line(data = train_curves[[2]], aes(x = recall, y = precision), color = '#34A6FF', size = 1) +
  geom_hline(
        lty = 2, alpha = 0.5,
        color = "gray50",
        size = 1,
        yintercept = noskill_pr_train
    ) + theme_bw() + ggtitle('Training PR Delivery NB')


test_pr_plot = ggplot() +
  geom_line(data = test_curves[[2]], aes(x = recall, y = precision), color = '#DD0534', size = 1) +
  geom_hline(
        lty = 2, alpha = 0.5,
        color = "gray50",
        size = 1,
        yintercept = noskill_pr_test
    ) + theme_bw() + ggtitle('Holdout PR Delivery NB')

plot_grid(train_roc_plot, train_pr_plot, test_roc_plot, test_pr_plot)
```

```{r dvSHAPs}
## Bind the data back together and get SHAPs for the full dataset.
data_shap = dv_data_train %>% rbind(dv_data_test)

shap_values = kernelshap(delivery_bfmodl, data_shap, bg_X = data_shap, type = "prob")
sv = shapviz(shap_values, which_class = 1)

sv_importance(sv, kind = "bee")

sv_force(sv, row_id = 1)
sv_waterfall(sv, row_id = 1)

```

# Predicting seroconversion at NB
```{r nbSplitData}
## Variable names for configs
dependentvar = 'Seroconverted'
indivar = 'enrollment_id'
sampvar = 'sample_name'
timevar = 'visit'
timepoint = 'NB'
covars = c(colnames(alphadivs)) %>% setdiff('rn')
strata = c('country') ## stratify to ensure even coverage of countries in the training/testing split

ttmd = filtered_metadata[get(timevar) == timepoint, mget(c(sampvar, indivar, covars, dependentvar, strata))] %>% na.omit()

train_indivs = ttmd[,sample(get(indivar), ceiling(.N*train_proportion)), by = c(strata)][,V1]
test_indivs = setdiff(ttmd[,get(indivar)], train_indivs)
train_samples = ttmd[get(indivar) %in% train_indivs, get(sampvar)]
test_samples = ttmd[get(indivar) %in% test_indivs, get(sampvar)]
```

```{r nbTrainingSplit}
## Training covariate data
train_samples = ttmd[get(sampvar) %in% train_samples, get(sampvar)] ## Reorder according to the table.
train_covars = ttmd[get(sampvar) %in% train_samples, mget(covars)]
train_ids = ttmd[get(sampvar) %in% train_samples, mget(indivar)]
train_counts = goodcounts[,train_samples] %>% apply(2, function(x) x / sum(x))

## Response var
train_response = ttmd[get(sampvar) %in% train_samples, mget(dependentvar)]

## Agglomerate.
ogu_train_counts = train_counts %>%
    abundance(min_abundance) %>%
    prevalence(min_prevalence) %>%
    t() %>% 
    as_tibble() %>%
    rename_all(~ make.names(.)) %>%
    as.matrix()

spp_train_counts = train_counts %>%
    agglomerate(taxonomy, 's') %>%
    abundance(min_abundance) %>%
    prevalence(min_prevalence) %>%
    t() %>% 
    as_tibble() %>%
    rename_all(~ make.names(.)) %>%
    rename_all(~ paste0('s__',.)) %>%
    as.matrix()

gen_train_counts = train_counts %>%
    agglomerate(taxonomy, 'g') %>%
    abundance(min_abundance) %>%
    prevalence(min_prevalence) %>%
    t() %>% 
    as_tibble() %>%
    rename_all(~ make.names(.)) %>%
    rename_all(~ paste0('g__',.)) %>%
    as.matrix()

## Taxon-specific comps.
ogu_train_comp = ogu_train_counts[,-union(ogu_train_counts %>% nearZeroVar()
                                    , which(colSums(ogu_train_counts) == 0)
                                    )]
spp_train_comp = spp_train_counts[,-union(spp_train_counts %>% nearZeroVar()
                                    , which(colSums(spp_train_counts) == 0)
                                    )]
gen_train_comp = gen_train_counts[,-union(gen_train_counts %>% nearZeroVar()
                                    , which(colSums(gen_train_counts) == 0)
                                    )]

## Build training dataframe.
nb_data_train = cbind(train_response, train_covars, gen_train_comp, spp_train_comp) %>% as_tibble()
```

```{r nbTestingSplit}
## Training covariate data
test_samples = ttmd[get(sampvar) %in% test_samples, get(sampvar)] ## Reorder according to the table.
test_covars = ttmd[get(sampvar) %in% test_samples, mget(covars)]
test_ids = ttmd[get(sampvar) %in% test_samples, mget(indivar)]
test_counts = goodcounts[,test_samples] %>% apply(2, function(x) x / sum(x))

## Response var
test_response = ttmd[get(sampvar) %in% test_samples, mget(dependentvar)]

## Agglomerate.
ogu_test_counts = test_counts %>%
    t() %>% 
    as_tibble() %>%
    rename_all(~ make.names(.)) %>%
    dplyr::select(!!colnames(ogu_train_counts)) %>%
    as.matrix()

spp_test_counts = test_counts %>% agglomerate(taxonomy, 's') %>%
    t() %>% 
    as_tibble() %>%
    rename_all(~ make.names(.)) %>%
    rename_all(~ paste0('s__',.)) %>%
    dplyr::select(!!colnames(spp_train_counts)) %>%
    as.matrix()
 
gen_test_counts = test_counts %>% agglomerate(taxonomy, 'g') %>%
    t() %>% 
    as_tibble() %>%
    rename_all(~ make.names(.)) %>%
    rename_all(~ paste0('g__',.)) %>%
    dplyr::select(!!colnames(gen_train_counts)) %>%
    as.matrix()

## Taxon-specific comps.
ogu_test_comp = ogu_test_counts[,colnames(ogu_train_comp)]
spp_test_comp = spp_test_counts[,colnames(spp_train_comp)]
gen_test_comp = gen_test_counts[,colnames(gen_train_comp)]

## Build training dataframe.
nb_data_test = cbind(test_response, test_covars, gen_test_comp, spp_test_comp) %>% as_tibble()

## selection variables.
selvars = c(covars, colnames(spp_train_comp), colnames(gen_train_comp))
```

```{r nbTuneParams}
lgb_classifier = boost_tree(
   trees = lgb_ntrees
 , tree_depth = tune()
 , min_n = tune()
 , loss_reduction = tune()
 , sample_size = tune()
 , mtry = tune()
 , learn_rate = tune()
 ) %>%
set_engine("lightgbm",
           #is_unbalance = TRUE,
           #boosting = "dart",
           print_every_n=10) %>%
set_mode("classification")

modl = lgb_classifier

modl_formula = paste0(dependentvar, ' ~ .') %>% as.formula()
modl_recipe = recipe(modl_formula, data = nb_data_train)

modl_workflow = workflow() %>% 
    add_model(modl) %>%
    add_recipe(modl_recipe)

print(modl_workflow)
    
modl_cvfolds = vfold_cv(nb_data_train, strata = dependentvar, v = lgb_k, repeats = lgb_repeats);

## Make tuning grid.
lgb_tune_grid = grid_latin_hypercube(
   tree_depth()
 , min_n()
 , loss_reduction()
 , sample_size = sample_prop()
 , finalize(mtry(), nb_data_train)
 , learn_rate()
 #, alpha()
 , size = lgb_gridsize
 )
    
## Exec cross-validation of the model
message('Running cross-validation of model...')
nb_lgb_res = tune_grid(
  modl_workflow
  , resamples = modl_cvfolds
  , grid = lgb_tune_grid
  , metric_set(yardstick::accuracy
             , yardstick::brier_class
             , yardstick::f_meas
             , yardstick::pr_auc
             , yardstick::kap
             , yardstick::roc_auc
             )
  , control = control_grid(save_pred = TRUE, verbose = TRUE, save_workflow = TRUE)
)

```

```{r nbFit}
nb_bfmodl = fit_best(nb_lgb_res, metric = 'brier_class')
```

```{r nbEvalModel}


train_curves = get_roc_pr_curves(model = nb_bfmodl
                  , data = nb_data_train
                  , truth = nb_data_train[,dependentvar]
                  , label = 'training')

## Eval holdout data
test_curves = get_roc_pr_curves(model = nb_bfmodl
                  , data = nb_data_test
                  , truth = nb_data_test[,dependentvar]
                  , label = 'holdout')

## Plor ROC-curves.
train_roc_plot = ggplot() +
  geom_line(data = train_curves[[1]], aes(x = 1 - specificity, y = sensitivity), color = '#34A6FF', size = 1) +
  geom_abline(
        lty = 2, alpha = 0.5,
        color = "gray50",
        size = 1
    ) + theme_bw() + ggtitle('Training ROC Seroconversion NB')

test_roc_plot = ggplot() +
  geom_line(data = test_curves[[1]], aes(x = 1 - specificity, y = sensitivity), color = '#DD0534', size = 1) +
  geom_abline(
        lty = 2, alpha = 0.5,
        color = "gray50",
        size = 1
    ) + theme_bw() + ggtitle('Holdout ROC Seroconversion NB')

## Plot PR curves.
noskill_pr_test = nb_data_test[,dependentvar] %>% unlist() %>% table() %>% (function(x) x[1] / (x[2] + x[1]))
noskill_pr_train = nb_data_train[,dependentvar] %>% unlist() %>% table() %>% (function(x) x[1] / (x[2] + x[1]))

train_pr_plot = ggplot() +
  geom_line(data = train_curves[[2]], aes(x = recall, y = precision), color = '#34A6FF', size = 1) +
  geom_hline(
        lty = 2, alpha = 0.5,
        color = "gray50",
        size = 1,
        yintercept = noskill_pr_train
    ) + theme_bw() + ggtitle('Training PR Seroconversion NB')


test_pr_plot = ggplot() +
  geom_line(data = test_curves[[2]], aes(x = recall, y = precision), color = '#DD0534', size = 1) +
  geom_hline(
        lty = 2, alpha = 0.5,
        color = "gray50",
        size = 1,
        yintercept = noskill_pr_test
    ) + theme_bw() + ggtitle('Holdout PR Seroconversion NB')

plot_grid(train_roc_plot, train_pr_plot, test_roc_plot, test_pr_plot)


```

```{r nbSHAPs}
## Bind the data back together and get SHAPs for the full dataset.
data_shap = nb_data_train %>% rbind(nb_data_test)

shap_values = kernelshap(nb_bfmodl, data_shap, bg_X = data_shap, type = "prob")
sv = shapviz(shap_values, which_class = 1)

sv_importance(sv, kind = "bee")

sv_force(sv, row_id = 1)
sv_waterfall(sv, row_id = 1)

```

# Predicting seroconversion at 2mo
```{r m2SplitData}
## Variable names for configs
dependentvar = 'Seroconverted'
indivar = 'enrollment_id'
sampvar = 'sample_name'
timevar = 'visit'
timepoint = '2mo'
covars = c(colnames(alphadivs)) %>% setdiff('rn')
strata = c('country') ## stratify to ensure even coverage of countries in the training/testing split

ttmd = filtered_metadata[get(timevar) == timepoint, mget(c(sampvar, indivar, covars, dependentvar, strata))] %>% na.omit()

train_indivs = ttmd[,sample(get(indivar), ceiling(.N*train_proportion)), by = c(strata)][,V1]
test_indivs = setdiff(ttmd[,get(indivar)], train_indivs)
train_samples = ttmd[get(indivar) %in% train_indivs, get(sampvar)]
test_samples = ttmd[get(indivar) %in% test_indivs, get(sampvar)]
```

```{r m2TrainingSplit}
## Training covariate data
train_samples = ttmd[get(sampvar) %in% train_samples, get(sampvar)] ## Reorder according to the table.
train_covars = ttmd[get(sampvar) %in% train_samples, mget(covars)]
train_ids = ttmd[get(sampvar) %in% train_samples, mget(indivar)]
train_counts = goodcounts[,train_samples] %>% apply(2, function(x) x / sum(x))

## Response var
train_response = ttmd[get(sampvar) %in% train_samples, mget(dependentvar)]

## Agglomerate.
ogu_train_counts = train_counts %>%
    abundance(min_abundance) %>%
    prevalence(min_prevalence) %>%
    t() %>% 
    as_tibble() %>%
    rename_all(~ make.names(.)) %>%
    as.matrix()

spp_train_counts = train_counts %>%
    agglomerate(taxonomy, 's') %>%
    abundance(min_abundance) %>%
    prevalence(min_prevalence) %>%
    t() %>% 
    as_tibble() %>%
    rename_all(~ make.names(.)) %>%
    rename_all(~ paste0('s__',.)) %>%
    as.matrix()

gen_train_counts = train_counts %>%
    agglomerate(taxonomy, 'g') %>%
    abundance(min_abundance) %>%
    prevalence(min_prevalence) %>%
    t() %>% 
    as_tibble() %>%
    rename_all(~ make.names(.)) %>%
    rename_all(~ paste0('g__',.)) %>%
    as.matrix()

## Taxon-specific comps.
ogu_train_comp = ogu_train_counts[,-union(ogu_train_counts %>% nearZeroVar()
                                    , which(colSums(ogu_train_counts) == 0)
                                    )]
spp_train_comp = spp_train_counts[,-union(spp_train_counts %>% nearZeroVar()
                                    , which(colSums(spp_train_counts) == 0)
                                    )]
gen_train_comp = gen_train_counts[,-union(gen_train_counts %>% nearZeroVar()
                                    , which(colSums(gen_train_counts) == 0)
                                    )]

## Build training dataframe.
m2_data_train = cbind(train_response, train_covars, gen_train_comp, spp_train_comp) %>% as_tibble()
```

```{r m2TestingSplit}
## Training covariate data
test_samples = ttmd[get(sampvar) %in% test_samples, get(sampvar)] ## Reorder according to the table.
test_covars = ttmd[get(sampvar) %in% test_samples, mget(covars)]
test_ids = ttmd[get(sampvar) %in% test_samples, mget(indivar)]
test_counts = goodcounts[,test_samples] %>% apply(2, function(x) x / sum(x))

## Response var
test_response = ttmd[get(sampvar) %in% test_samples, mget(dependentvar)]

## Agglomerate.
ogu_test_counts = test_counts %>%
    t() %>% 
    as_tibble() %>%
    rename_all(~ make.names(.)) %>%
    dplyr::select(!!colnames(ogu_train_counts)) %>%
    as.matrix()

spp_test_counts = test_counts %>% agglomerate(taxonomy, 's') %>%
    t() %>% 
    as_tibble() %>%
    rename_all(~ make.names(.)) %>%
    rename_all(~ paste0('s__',.)) %>%
    dplyr::select(!!colnames(spp_train_counts)) %>%
    as.matrix()
 
gen_test_counts = test_counts %>% agglomerate(taxonomy, 'g') %>%
    t() %>% 
    as_tibble() %>%
    rename_all(~ make.names(.)) %>%
    rename_all(~ paste0('g__',.)) %>%
    dplyr::select(!!colnames(gen_train_counts)) %>%
    as.matrix()

## Taxon-specific comps.
ogu_test_comp = ogu_test_counts[,colnames(ogu_train_comp)]
spp_test_comp = spp_test_counts[,colnames(spp_train_comp)]
gen_test_comp = gen_test_counts[,colnames(gen_train_comp)]

## Build training dataframe.
m2_data_test = cbind(test_response, test_covars, gen_test_comp, spp_test_comp) %>% as_tibble()

```

```{r m2TuneParams}
lgb_classifier = boost_tree(
   trees = lgb_ntrees
 , tree_depth = tune()
 , min_n = tune()
 , loss_reduction = tune()
 , sample_size = tune()
 , mtry = tune()
 , learn_rate = tune()
 ) %>%
set_engine("lightgbm",
           #is_unbalance = TRUE,
           #boosting = "dart",
           print_every_n=10) %>%
set_mode("classification")

modl = lgb_classifier

modl_formula = paste0(dependentvar, ' ~ .') %>% as.formula()
modl_recipe = recipe(modl_formula, data = m2_data_train)

modl_workflow = workflow() %>% 
    add_model(modl) %>%
    add_recipe(modl_recipe)

print(modl_workflow)
    
modl_cvfolds = vfold_cv(m2_data_train, strata = dependentvar, v = lgb_k, repeats = lgb_repeats);

## Make tuning grid.
lgb_tune_grid = grid_latin_hypercube(
   tree_depth()
 , min_n()
 , loss_reduction()
 , sample_size = sample_prop()
 , finalize(mtry(), m2_data_train)
 , learn_rate()
 #, alpha()
 , size = lgb_gridsize
 )
    
## Exec cross-validation of the model
message('Running cross-validation of model...')
m2_lgb_res = tune_grid(
  modl_workflow
  , resamples = modl_cvfolds
  , grid = lgb_tune_grid
  , metric_set(yardstick::accuracy
             , yardstick::brier_class
             , yardstick::f_meas
             , yardstick::pr_auc
             , yardstick::kap
             , yardstick::roc_auc
             )
  , control = control_grid(save_pred = TRUE, verbose = TRUE, save_workflow = TRUE)
)

```

```{r m2Fit}
m2_bfmodl = fit_best(m2_lgb_res, metric = 'brier_class')
```

```{r m2EvalModel}
train_curves = get_roc_pr_curves(model = m2_bfmodl
                  , data = m2_data_train
                  , truth = m2_data_train[,dependentvar]
                  , label = 'training')

## Eval holdout data
test_curves = get_roc_pr_curves(model = m2_bfmodl
                  , data = m2_data_test
                  , truth = m2_data_test[,dependentvar]
                  , label = 'holdout')

## Plor ROC-curves.
train_roc_plot = ggplot() +
  geom_line(data = train_curves[[1]], aes(x = 1 - specificity, y = sensitivity), color = '#34A6FF', size = 1) +
  geom_abline(
        lty = 2, alpha = 0.5,
        color = "gray50",
        size = 1
    ) + theme_bw() + ggtitle('Training ROC Seroconversion 2mo')

test_roc_plot = ggplot() +
  geom_line(data = test_curves[[1]], aes(x = 1 - specificity, y = sensitivity), color = '#DD0534', size = 1) +
  geom_abline(
        lty = 2, alpha = 0.5,
        color = "gray50",
        size = 1
    ) + theme_bw() + ggtitle('Holdout ROC Seroconversion 2mo')

## Plot PR curves.
noskill_pr_test = m2_data_test[,dependentvar] %>% unlist() %>% table() %>% (function(x) x[1] / (x[2] + x[1]))
noskill_pr_train = m2_data_train[,dependentvar] %>% unlist() %>% table() %>% (function(x) x[1] / (x[2] + x[1]))

train_pr_plot = ggplot() +
  geom_line(data = train_curves[[2]], aes(x = recall, y = precision), color = '#34A6FF', size = 1) +
  geom_hline(
        lty = 2, alpha = 0.5,
        color = "gray50",
        size = 1,
        yintercept = noskill_pr_train
    ) + theme_bw() + ggtitle('Training PR Seroconversion 2mo')


test_pr_plot = ggplot() +
  geom_line(data = test_curves[[2]], aes(x = recall, y = precision), color = '#DD0534', size = 1) +
  geom_hline(
        lty = 2, alpha = 0.5,
        color = "gray50",
        size = 1,
        yintercept = noskill_pr_test
    ) + theme_bw() + ggtitle('Holdout PR Seroconversion 2mo')

plot_grid(train_roc_plot, train_pr_plot, test_roc_plot, test_pr_plot)


```

```{r m2SHAPs}
data_shap = m2_data_train %>% rbind(m2_data_test)

shap_values = kernelshap(m2_bfmodl, data_shap, bg_X = data_shap, type = "prob")
sv = shapviz(shap_values, which_class = 1)

sv_importance(sv, kind = "bee")

sv_force(sv, row_id = 1)
sv_waterfall(sv, row_id = 1)

```

# Predicting seroconversion at 4mo
```{r m4SplitData}
## Variable names for configs
dependentvar = 'Seroconverted'
indivar = 'enrollment_id'
sampvar = 'sample_name'
timevar = 'visit'
timepoint = '4mo'
covars = c(colnames(alphadivs)) %>% setdiff('rn')
strata = c('country') ## stratify to ensure even coverage of countries in the training/testing split

ttmd = filtered_metadata[get(timevar) == timepoint, mget(c(sampvar, indivar, covars, dependentvar, strata))] %>% na.omit()

train_indivs = ttmd[,sample(get(indivar), ceiling(.N*train_proportion)), by = c(strata)][,V1]
test_indivs = setdiff(ttmd[,get(indivar)], train_indivs)
train_samples = ttmd[get(indivar) %in% train_indivs, get(sampvar)]
test_samples = ttmd[get(indivar) %in% test_indivs, get(sampvar)]
```

```{r m4TrainingSplit}
## Training covariate data
train_samples = ttmd[get(sampvar) %in% train_samples, get(sampvar)] ## Reorder according to the table.
train_covars = ttmd[get(sampvar) %in% train_samples, mget(covars)]
train_ids = ttmd[get(sampvar) %in% train_samples, mget(indivar)]
train_counts = goodcounts[,train_samples] %>% apply(2, function(x) x / sum(x))

## Response var
train_response = ttmd[get(sampvar) %in% train_samples, mget(dependentvar)]

## Agglomerate.
ogu_train_counts = train_counts %>%
    abundance(min_abundance) %>%
    prevalence(min_prevalence) %>%
    t() %>% 
    as_tibble() %>%
    rename_all(~ make.names(.)) %>%
    as.matrix()

spp_train_counts = train_counts %>%
    agglomerate(taxonomy, 's') %>%
    abundance(min_abundance) %>%
    prevalence(min_prevalence) %>%
    t() %>% 
    as_tibble() %>%
    rename_all(~ make.names(.)) %>%
    rename_all(~ paste0('s__',.)) %>%
    as.matrix()

gen_train_counts = train_counts %>%
    agglomerate(taxonomy, 'g') %>%
    abundance(min_abundance) %>%
    prevalence(min_prevalence) %>%
    t() %>% 
    as_tibble() %>%
    rename_all(~ make.names(.)) %>%
    rename_all(~ paste0('g__',.)) %>%
    as.matrix()

## Taxon-specific comps.
ogu_train_comp = ogu_train_counts[,-union(ogu_train_counts %>% nearZeroVar()
                                    , which(colSums(ogu_train_counts) == 0)
                                    )]
spp_train_comp = spp_train_counts[,-union(spp_train_counts %>% nearZeroVar()
                                    , which(colSums(spp_train_counts) == 0)
                                    )]
gen_train_comp = gen_train_counts[,-union(gen_train_counts %>% nearZeroVar()
                                    , which(colSums(gen_train_counts) == 0)
                                    )]

## Build training dataframe.
m4_data_train = cbind(train_response, train_covars, gen_train_comp, spp_train_comp) %>% as_tibble()
```

```{r m4TestingSplit}
## Training covariate data
test_samples = ttmd[get(sampvar) %in% test_samples, get(sampvar)] ## Reorder according to the table.
test_covars = ttmd[get(sampvar) %in% test_samples, mget(covars)]
test_ids = ttmd[get(sampvar) %in% test_samples, mget(indivar)]
test_counts = goodcounts[,test_samples] %>% apply(2, function(x) x / sum(x))

## Response var
test_response = ttmd[get(sampvar) %in% test_samples, mget(dependentvar)]

## Agglomerate.
ogu_test_counts = test_counts %>%
    t() %>% 
    as_tibble() %>%
    rename_all(~ make.names(.)) %>%
    dplyr::select(!!colnames(ogu_train_counts)) %>%
    as.matrix()

spp_test_counts = test_counts %>% agglomerate(taxonomy, 's') %>%
    t() %>% 
    as_tibble() %>%
    rename_all(~ make.names(.)) %>%
    rename_all(~ paste0('s__',.)) %>%
    dplyr::select(!!colnames(spp_train_counts)) %>%
    as.matrix()
 
gen_test_counts = test_counts %>% agglomerate(taxonomy, 'g') %>%
    t() %>% 
    as_tibble() %>%
    rename_all(~ make.names(.)) %>%
    rename_all(~ paste0('g__',.)) %>%
    dplyr::select(!!colnames(gen_train_counts)) %>%
    as.matrix()

## Taxon-specific comps.
ogu_test_comp = ogu_test_counts[,colnames(ogu_train_comp)]
spp_test_comp = spp_test_counts[,colnames(spp_train_comp)]
gen_test_comp = gen_test_counts[,colnames(gen_train_comp)]

## Build training dataframe.
m4_data_test = cbind(test_response, test_covars, gen_test_comp, spp_test_comp) %>% as_tibble()

```

```{r m4TuneParams}
lgb_classifier = boost_tree(
   trees = lgb_ntrees
 , tree_depth = tune()
 , min_n = tune()
 , loss_reduction = tune()
 , sample_size = tune()
 , mtry = tune()
 , learn_rate = tune()
 ) %>%
set_engine("lightgbm",
           #is_unbalance = TRUE,
           #boosting = "dart",
           print_every_n=10) %>%
set_mode("classification")

modl = lgb_classifier

modl_formula = paste0(dependentvar, ' ~ .') %>% as.formula()
modl_recipe = recipe(modl_formula, data = m4_data_train)

modl_workflow = workflow() %>% 
    add_model(modl) %>%
    add_recipe(modl_recipe)

print(modl_workflow)
    
modl_cvfolds = vfold_cv(m4_data_train, strata = dependentvar, v = lgb_k, repeats = lgb_repeats);

## Make tuning grid.
lgb_tune_grid = grid_latin_hypercube(
   tree_depth()
 , min_n()
 , loss_reduction()
 , sample_size = sample_prop()
 , finalize(mtry(), m4_data_train)
 , learn_rate()
 #, alpha()
 , size = lgb_gridsize
 )
    
## Exec cross-validation of the model
message('Running cross-validation of model...')
m4_lgb_res = tune_grid(
  modl_workflow
  , resamples = modl_cvfolds
  , grid = lgb_tune_grid
  , metric_set(yardstick::accuracy
             , yardstick::brier_class
             , yardstick::f_meas
             , yardstick::pr_auc
             , yardstick::kap
             , yardstick::roc_auc
             )
  , control = control_grid(save_pred = TRUE, verbose = TRUE, save_workflow = TRUE)
)
```

```{r m4Fit}
m4_bfmodl = fit_best(m4_lgb_res, metric = 'brier_class')
```

```{r m4EvalModel}
train_curves = get_roc_pr_curves(model = m4_bfmodl
                  , data = m4_data_train
                  , truth = m4_data_train[,dependentvar]
                  , label = 'training')

## Eval holdout data
test_curves = get_roc_pr_curves(model = m4_bfmodl
                  , data = m4_data_test
                  , truth = m4_data_test[,dependentvar]
                  , label = 'holdout')

## Plor ROC-curves.
train_roc_plot = ggplot() +
  geom_line(data = train_curves[[1]], aes(x = 1 - specificity, y = sensitivity), color = '#34A6FF', size = 1) +
  geom_abline(
        lty = 2, alpha = 0.5,
        color = "gray50",
        size = 1
    ) + theme_bw() + ggtitle('Training ROC Seroconversion 4mo')

test_roc_plot = ggplot() +
  geom_line(data = test_curves[[1]], aes(x = 1 - specificity, y = sensitivity), color = '#DD0534', size = 1) +
  geom_abline(
        lty = 2, alpha = 0.5,
        color = "gray50",
        size = 1
    ) + theme_bw() + ggtitle('Holdout ROC Seroconversion 4mo')

## Plot PR curves.
noskill_pr_test = m4_data_test[,dependentvar] %>% unlist() %>% table() %>% (function(x) x[1] / (x[2] + x[1]))
noskill_pr_train = m4_data_train[,dependentvar] %>% unlist() %>% table() %>% (function(x) x[1] / (x[2] + x[1]))

train_pr_plot = ggplot() +
  geom_line(data = train_curves[[2]], aes(x = recall, y = precision), color = '#34A6FF', size = 1) +
  geom_hline(
        lty = 2, alpha = 0.5,
        color = "gray50",
        size = 1,
        yintercept = noskill_pr_train
    ) + theme_bw() + ggtitle('Training PR Seroconversion 4mo')


test_pr_plot = ggplot() +
  geom_line(data = test_curves[[2]], aes(x = recall, y = precision), color = '#DD0534', size = 1) +
  geom_hline(
        lty = 2, alpha = 0.5,
        color = "gray50",
        size = 1,
        yintercept = noskill_pr_test
    ) + theme_bw() + ggtitle('Holdout PR Seroconversion 4mo')

plot_grid(train_roc_plot, train_pr_plot, test_roc_plot, test_pr_plot)


```

```{r m4SHAPs}
data_shap = m4_data_train %>% rbind(m4_data_test)

shap_values = kernelshap(m4_bfmodl, data_shap, bg_X = data_shap, type = "prob")
sv = shapviz(shap_values, which_class = 1)

sv_importance(sv, kind = "bee")

sv_force(sv, row_id = 1)
sv_waterfall(sv, row_id = 1)

```

# Predicting shedding
```{r shSplitData}
## Variable names for configs
dependentvar = 'rota_stool_pcr'
indivar = 'enrollment_id'
sampvar = 'sample_name'
timevar = 'visit'
timepoint = 'NB'
covars = c(colnames(alphadivs)) %>% setdiff('rn')
strata = c('country') ## stratify to ensure even coverage of countries in the training/testing split

ttmd = filtered_metadata[get(timevar) == timepoint, mget(c(sampvar, indivar, covars, dependentvar, strata))] %>% na.omit()

train_indivs = ttmd[,sample(get(indivar), ceiling(.N*train_proportion)), by = c(strata)][,V1]
test_indivs = setdiff(ttmd[,get(indivar)], train_indivs)
train_samples = ttmd[get(indivar) %in% train_indivs, get(sampvar)]
test_samples = ttmd[get(indivar) %in% test_indivs, get(sampvar)]
```

```{r shTrainingSplit}
## Training covariate data
train_samples = ttmd[get(sampvar) %in% train_samples, get(sampvar)] ## Reorder according to the table.
train_covars = ttmd[get(sampvar) %in% train_samples, mget(covars)]
train_ids = ttmd[get(sampvar) %in% train_samples, mget(indivar)]
train_counts = goodcounts[,train_samples] %>% apply(2, function(x) x / sum(x))

## Response var
train_response = ttmd[get(sampvar) %in% train_samples, mget(dependentvar)]

## Agglomerate.
ogu_train_counts = train_counts %>%
    abundance(min_abundance) %>%
    prevalence(min_prevalence) %>%
    t() %>% 
    as_tibble() %>%
    rename_all(~ make.names(.)) %>%
    as.matrix()

spp_train_counts = train_counts %>%
    agglomerate(taxonomy, 's') %>%
    abundance(min_abundance) %>%
    prevalence(min_prevalence) %>%
    t() %>% 
    as_tibble() %>%
    rename_all(~ make.names(.)) %>%
    rename_all(~ paste0('s__',.)) %>%
    as.matrix()

gen_train_counts = train_counts %>%
    agglomerate(taxonomy, 'g') %>%
    abundance(min_abundance) %>%
    prevalence(min_prevalence) %>%
    t() %>% 
    as_tibble() %>%
    rename_all(~ make.names(.)) %>%
    rename_all(~ paste0('g__',.)) %>%
    as.matrix()

## Taxon-specific comps.
ogu_train_comp = ogu_train_counts[,-union(ogu_train_counts %>% nearZeroVar()
                                    , which(colSums(ogu_train_counts) == 0)
                                    )]
spp_train_comp = spp_train_counts[,-union(spp_train_counts %>% nearZeroVar()
                                    , which(colSums(spp_train_counts) == 0)
                                    )]
gen_train_comp = gen_train_counts[,-union(gen_train_counts %>% nearZeroVar()
                                    , which(colSums(gen_train_counts) == 0)
                                    )]

## Build training dataframe.
sh_data_train = cbind(train_response, train_covars, gen_train_comp, spp_train_comp) %>% as_tibble()
```

```{r shTestingSplit}
## Training covariate data
test_samples = ttmd[get(sampvar) %in% test_samples, get(sampvar)] ## Reorder according to the table.
test_covars = ttmd[get(sampvar) %in% test_samples, mget(covars)]
test_ids = ttmd[get(sampvar) %in% test_samples, mget(indivar)]
test_counts = goodcounts[,test_samples] %>% apply(2, function(x) x / sum(x))

## Response var
test_response = ttmd[get(sampvar) %in% test_samples, mget(dependentvar)]

## Agglomerate.
ogu_test_counts = test_counts %>%
    t() %>% 
    as_tibble() %>%
    rename_all(~ make.names(.)) %>%
    dplyr::select(!!colnames(ogu_train_counts)) %>%
    as.matrix()

spp_test_counts = test_counts %>% agglomerate(taxonomy, 's') %>%
    t() %>% 
    as_tibble() %>%
    rename_all(~ make.names(.)) %>%
    rename_all(~ paste0('s__',.)) %>%
    dplyr::select(!!colnames(spp_train_counts)) %>%
    as.matrix()
 
gen_test_counts = test_counts %>% agglomerate(taxonomy, 'g') %>%
    t() %>% 
    as_tibble() %>%
    rename_all(~ make.names(.)) %>%
    rename_all(~ paste0('g__',.)) %>%
    dplyr::select(!!colnames(gen_train_counts)) %>%
    as.matrix()

## Taxon-specific comps.
ogu_test_comp = ogu_test_counts[,colnames(ogu_train_comp)]
spp_test_comp = spp_test_counts[,colnames(spp_train_comp)]
gen_test_comp = gen_test_counts[,colnames(gen_train_comp)]

## Build training dataframe.
sh_data_test = cbind(test_response, test_covars, gen_test_comp, spp_test_comp) %>% as_tibble()

```

```{r shTuneParams}
lgb_classifier = boost_tree(
   trees = lgb_ntrees
 , tree_depth = tune()
 , min_n = tune()
 , loss_reduction = tune()
 , sample_size = tune()
 , mtry = tune()
 , learn_rate = tune()
 ) %>%
set_engine("lightgbm",
           is_unbalance = TRUE,
           print_every_n=10) %>%
set_mode("classification")

modl = lgb_classifier

modl_formula = paste0(dependentvar, ' ~ .') %>% as.formula()
modl_recipe = recipe(modl_formula, data = sh_data_train)

modl_workflow = workflow() %>% 
    add_model(modl) %>%
    add_recipe(modl_recipe)

print(modl_workflow)
    
modl_cvfolds = vfold_cv(sh_data_train, strata = dependentvar, v = lgb_k, repeats = lgb_repeats);

## Make tuning grid.
lgb_tune_grid = grid_latin_hypercube(
   tree_depth()
 , min_n()
 , loss_reduction()
 , sample_size = sample_prop()
 , finalize(mtry(), sh_data_train)
 , learn_rate()
 #, alpha()
 , size = lgb_gridsize
 )
    
## Exec cross-validation of the model
message('Running cross-validation of model...')
sh_lgb_res = tune_grid(
  modl_workflow
  , resamples = modl_cvfolds
  , grid = lgb_tune_grid
  , metric_set(yardstick::accuracy
             , yardstick::brier_class
             , yardstick::f_meas
             , yardstick::pr_auc
             , yardstick::kap
             , yardstick::roc_auc
             )
  , control = control_grid(save_pred = TRUE, verbose = TRUE, save_workflow = TRUE)
)
```

```{r shFit}
sh_bfmodl = fit_best(sh_lgb_res, metric = 'brier_class')
```

```{r shEvalModel}


train_curves = get_roc_pr_curves(model = sh_bfmodl
                  , data = sh_data_train
                  , truth = sh_data_train[,dependentvar]
                  , label = 'training')

## Eval holdout data
test_curves = get_roc_pr_curves(model = sh_bfmodl
                  , data = sh_data_test
                  , truth = sh_data_test[,dependentvar]
                  , label = 'holdout')

## Plor ROC-curves.
train_roc_plot = ggplot() +
  geom_line(data = train_curves[[1]], aes(x = 1 - specificity, y = sensitivity), color = '#34A6FF', size = 1) +
  geom_abline(
        lty = 2, alpha = 0.5,
        color = "gray50",
        size = 1
    ) + theme_bw() + ggtitle('Training ROC Take NB')

test_roc_plot = ggplot() +
  geom_line(data = test_curves[[1]], aes(x = 1 - specificity, y = sensitivity), color = '#DD0534', size = 1) +
  geom_abline(
        lty = 2, alpha = 0.5,
        color = "gray50",
        size = 1
    ) + theme_bw() + ggtitle('Holdout ROC Take NB')

## Plot PR curves.
noskill_pr_test = sh_data_test[,dependentvar] %>% unlist() %>% table() %>% (function(x) x[1] / (x[2] + x[1]))
noskill_pr_train = sh_data_train[,dependentvar] %>% unlist() %>% table() %>% (function(x) x[1] / (x[2] + x[1]))

train_pr_plot = ggplot() +
  geom_line(data = train_curves[[2]], aes(x = recall, y = precision), color = '#34A6FF', size = 1) +
  geom_hline(
        lty = 2, alpha = 0.5,
        color = "gray50",
        size = 1,
        yintercept = noskill_pr_train
    ) + theme_bw() + ggtitle('Training PR Take NB')


test_pr_plot = ggplot() +
  geom_line(data = test_curves[[2]], aes(x = recall, y = precision), color = '#DD0534', size = 1) +
  geom_hline(
        lty = 2, alpha = 0.5,
        color = "gray50",
        size = 1,
        yintercept = noskill_pr_test
    ) + theme_bw() + ggtitle('Holdout PR Take NB')

plot_grid(train_roc_plot, train_pr_plot, test_roc_plot, test_pr_plot)


```

```{r shSHAPs}
## Bind the data back together and get SHAPs for the full dataset.
data_shap = sh_data_train %>% rbind(sh_data_test)

shap_values = kernelshap(sh_bfmodl, data_shap, bg_X = data_shap, type = "prob")
sv = shapviz(shap_values, which_class = 1)

sv_importance(sv, kind = "bee")

sv_force(sv, row_id = 1)
sv_waterfall(sv, row_id = 1)

```

# End
```{r}
save(list = ls(all.names = TRUE), file = 'rota-lgbm.RData')
sessionInfo()
```
